### 2019-07-05 paper.tex
Table of some model parameters.

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\title{Stochastic modeling of EGS using continuum-fracture approach and multilevel Monte Carlo}

\author{
  Jan Březina \\
  \And
  Pavel Exner \\
  \And
  Jan Stebel \\
  \And
  Martin Špetlík \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

% keywords can be removed
\keywords{First keyword \and Second keyword \and More}

\section{Introduction}

\section{Mathematical model}

\subsection{Geological properties}

Rock type: granite, gneiss or mica schist \cite{capova} % zula, rula, svor

\begin{tabular}{|l|c|c|c|}
\hline
 & rock & fractures \\\hline
Density & & \\
Hydraulic conductivity [$m.s^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ \\
Young modulus [GPa] & 50 \cite{ljunggren} & $50\times 10^{-9}$ \\
Poisson's ratio [-] & 0.25 \cite{ljunggren} & 0.25 \\
\hline
\end{tabular}

\subsection{Parameters of EGS site}

Well depth: approx. 5 km\\
Well radius: approx. 15 cm\\
Distance of injection and production well: \\
Pressure difference / water flux: 60 l/s

\begin{thebibliography}{9}

\bibitem{capova} L. Čápová. Specification of the geothermic model in the
environs of several selected boreholes. Diploma thesis, Charles University in Prague, 2013.

\bibitem{sperl-trckova} Jan ŠPERL and Jiřina TRČKOVÁ. PERMEABILITY AND POROSITY OF ROCKS AND THEIR RELATIONSHIP BASED ON LABORATORY TESTING. Acta Geodyn. Geomater., Vol. 5, No. 1 (149), 41-47, 2008.

\bibitem{ljunggren} Christer  Ljunggren et al. Mechanical properties of granitic rocks from Gide\aa, Sweden. Technical Report 85-06, SKB, 1985.

\end{thebibliography}

\end{document}

### 2019-07-05 paper.tex

@@
 \usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{amsfonts,amsmath} % blackboard math symbols
 \usepackage{nicefrac}       % compact symbols for 1/2, etc.
 \usepackage{microtype}      % microtypography
+\usepackage[bbgreekl]{mathbbol} % for blackboard \sigma
+
+\renewcommand{\div}{\operatorname{div}}
+\newcommand{\eq}[1]{\begin{equation}#1\end{equation}}
+\newcommand{\nn}{\vc n}
+\newcommand{\uu}{\vc u}
+\newcommand{\vc}[1]{\boldsymbol{#1}}
+\newcommand{\xx}{\vc x}
 
@@
 
+Liquid: water
+
 \begin{tabular}{|l|c|c|c|}
 \hline
- & rock & fractures \\\hline
-Density & & \\
-Hydraulic conductivity [$m.s^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ \\
-Young modulus [GPa] & 50 \cite{ljunggren} & $50\times 10^{-9}$ \\
-Poisson's ratio [-] & 0.25 \cite{ljunggren} & 0.25 \\
+ & rock & fractures & liquid\\\hline
+Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
+Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ & ---\\
+Porosity $\vartheta$ [--] & $5\times10^{-3}$ \cite{intera} & $5\times10^{-2}$ & --- \\
+Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ \cite{zisman} & $2\times10^{-11}$ & $5\times10^{-10}$ \\
+Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
+Cross-section $\delta$ [m] & --- & $10^{-6}$ & --- \\
+\hline
+Young modulus $G$ [Pa] & $50\times10^9$ \cite{ljunggren} & 50 & --- \\
+Poisson's ratio $\nu$ [--] & 0.25 \cite{ljunggren} & 0.25 & --- \\
+Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
+\hline
+Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
+Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
 \hline
@@
 
-\subsection{Parameters of EGS site}
+Storativity is calculated as follows \cite{brace-et-al}:
+\[ S = \varrho_l g(\beta_r + \vartheta\beta_l), \]
+where $g$ is the constant of gravity, which gives approximately the same values both for rock and fractures.
 
-Well depth: approx. 5 km\\
-Well radius: approx. 15 cm\\
-Distance of injection and production well: \\
-Pressure difference / water flux: 60 l/s
+\subsection{Parameters of EGS site and model geometry}
+
+Number of wells: 2 \\
+Depth: 5 km \\
+Distance: 200 m \\
+Radius: 15 cm \\
+Length of open part of wells: 60 m \\
+
+% Pressure difference / water flux: 60 l/s
+
+In our model we consider the cube $600 \times 600 \times 600$ m around the lower well ends, from which two cylinders of radius 10 m and height 60 m are removed.
+These represent the surrounding of the open well parts.
+The reason for using 10 m radius instead of the real well radius is to keep the number of necessary mesh elements reasonable.
+The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+
+On the surface of the artificial cylinders we shall extrapolate the value of piezometric head using the following idea: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
+\eq{\label{eq:H_log} H(\xx) = H_w\frac{\log d_w(|\xx|)}{\log r_w}. }
+Then, on the surface of the cylinder at radius $r_c$ we impose the Robin condition in the Darcy flow equation:
+\eq{\label{eq:H_robin} k\nabla H\cdot\nn = \sigma(H-H_w), }
+where $\nn$ stands for the unit outward normal vector.
+From \eqref{eq:H_log} and \eqref{eq:H_robin} we obtain the transition coefficient
+\[ \sigma = \frac{k}{r_c}\left|\frac{\log r_c}{\log r_w}\right| \]
+whose value depends on the hydraulic conductivity.
+
+\subsection{Hydro-mechanical model}
+\label{sc:hm_model}
+
+In the first model scenario the fluid is injected to both wells in order to open the preexisting fractures.
+The result is the modified cross-section of fractures.
+For simplicity we stimulate both wells simultaneously which produces very similar result in the change of fracture cross-section as if the wells were stimulated one by one.
+The liquid is injected for 1 day under the pressure 100 MPa (which corresponds to piezometric head $10^4$ m).
+
+Initial condition: zero piezometric head
+
+Boundary conditions:
+\begin{itemize}
+\item piezometric head $10^4$ m on surface of wells (including the intersections with fractures)
+\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
+\item no fluid flow through bottom and top part of the box and fracture tips
+\item zero displacement on bottom part of the box
+\item zero traction elsewhere
+\end{itemize}
+
+We use the Biot poroelasticity equations in the 3d continuum and 2d network of fractures.
+By dimension reduction of the Biot equations we obtain fluxes and forces which couple the rock with the lower dimensional fractures:
+\[ \partial_t(Sp + \alpha\div\uu) - k\Delta p = 0,\qquad -\mu\div(\nabla\uu+\nabla\uu^\top) - \lambda\nabla\div\uu + \alpha\varrho g\nabla H = 0. \]
+Here $p=H-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\mu$ and $\lambda$ are the Lamé parameters [Pa].
+...
+
+\subsection{Thermo-hydraulic model}
+
+The second scenario corresponds to operation of the EGS within 30 years after hydraulic stimulation performed according to Sec. \ref{sc:hm_model}.
+The fluid is continuously injected into one well and pumped out of the second one under the pressure difference 20 MPa and with inlet temperature $15\ ^\circ$C.
+
+Initial condition:
+temperature corresponding to $10\ ^\circ$C at the Earth surface and geothermal gradient $30\ ^\circ$C.km${}^{-1}$
+
+Boundary conditions:
+\begin{itemize}
+\item piezometric head $\pm10^3$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
+\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
+\item no flux on bottom and top part of the box and on fracture tips
+\item temperature $15\ ^\circ$C on surface of injection well (ZDE BY SE MELA UDELAT APROXIMACE DIRICHLETA ROBINEM PODOBNE JAKO V PROUDENI Z NEJAKEHO ANALYTICKEHO RESENI)
+\item initial temperature on lateral sides of the box where fluid flows into the box
+\item no heat flux through bottom and top part of the box and fracture tips
+\end{itemize}
+
+We use the following system of heat equation and steady-state Darcy flow with coupling terms connecting the rock to the fractures:
+\[ ... \]
 
@@
 
-\bibitem{capova} L. Čápová. Specification of the geothermic model in the
-environs of several selected boreholes. Diploma thesis, Charles University in Prague, 2013.
+\bibitem{brace-et-al} W. F. Brace,  J. B. Walsh, W. T. Grango.  Permeability of Granite Under High Pressure, J. Geophys. Res. 73(6):2225--2236, 1968.
 
-\bibitem{sperl-trckova} Jan ŠPERL and Jiřina TRČKOVÁ. PERMEABILITY AND POROSITY OF ROCKS AND THEIR RELATIONSHIP BASED ON LABORATORY TESTING. Acta Geodyn. Geomater., Vol. 5, No. 1 (149), 41-47, 2008.
+\bibitem{capova} L. Čápová. Specification of the geothermic model in the environs of several selected boreholes. Diploma thesis, Charles University in Prague, 2013.
 
-\bibitem{ljunggren} Christer  Ljunggren et al. Mechanical properties of granitic rocks from Gide\aa, Sweden. Technical Report 85-06, SKB, 1985.
+\bibitem{intera} INTERA Environmental Consultants, Inc. Porosity, Permeability,  and Their  Relationship  in  
+Granite,  Basalt, and Tuff. Accession  DE83-011519, NTIS, Springfield, Virginia, 1983.
+
+\bibitem{sperl-trckova} J. Šperl, J. Trčková. Permeability and porosity of rocks and their relationship based on laboratory testing. Acta Geodyn. Geomater. 5(149):41--47, 2008.
+
+\bibitem{ljunggren} C. Ljunggren, O. Stephansson, O. Alm, H. Hakami, U. Mattila. Mechanical properties of granitic rocks from Gide\aa, Sweden. Technical Report 85-06, SKB, 1985.
+
+\bibitem{zisman} W.
+A.
+Zisman.
+Compressibility and anisotropy of rocks at and near the Earth's surface.
+Proceedings of the National Academy of Sciences of the United States of America, 19(7):666--679, 1933.
 

### 2019-07-10 flow123d-data/parse_temp.py
Update text and simulation results.

@@ -1,0 +1,9 @@
+import yaml
+import sys
+
+dir = sys.argv[1]
+
+f = open(dir + '/heat_observe.yaml', 'r')
+d = yaml.safe_load(f)
+for i in d['data']:
+  print(i['time'],i['temperature'][0])

### 2019-07-10 flow123d-data/plot_power.sh

@@ -1,8 +1,16 @@
 #!/bin/bash
 
-#make plot of production power (sum of energy flux through well and fracture boundaries)
+dir=${1:-test_th}
+
+#make plot of production power
 echo "
 set terminal pdf color enhanced lw 2
-set output 'power.pdf'
+set output '${dir}_power.pdf'
 set xlabel 'time [years]'
-plot [1:] \"< sed -E '/\\\\.(left_fr_left_well|left_well|right_fr_right_well|right_well)\\\"/p' output_th/energy_balance.txt | awk 'BEGIN {t=0;v=0} {if (t==\$1) v+=\$4; else {print t, v; t=\$1;v=\$4}} END {print t,v}'\" u (\$1/365.25/86400):(-\$2/1e6) w lp t 'power [MW]'" | gnuplot
+set ylabel 'power [MW]'
+set y2label 'temperature [C]'
+set ytics nomirror
+set y2tics
+plot [1:] \"< sed -E '/\\\\.(left_fr_left_well|left_well|right_fr_right_well|right_well)\\\"/p' $dir/energy_balance.txt | awk 'BEGIN {t=0;v=0} {if (t==\$1) v+=\$4; else {print t, v; t=\$1;v=\$4}} END {print t,v}'\" u (\$1/365.25/86400):(-\$2/1e6) w l t 'power',\\
+          \"< python3 parse_temp.py $dir\" u (\$1/86400/365.25):(\$2-273.15) w l axes x1y2 t 'temperature'" | gnuplot
+

### 2019-07-10 paper.tex
Updates from Overleaf

@@
 Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
-Cross-section $\delta$ [m] & --- & $10^{-6}$ & --- \\
+Cross-section $\delta$ [m] & --- & $10^{-3}$ & --- \\
 \hline
@@
 \[ S = \varrho_l g(\beta_r + \vartheta\beta_l), \]
-where $g$ is the constant of gravity, which gives approximately the same values both for rock and fractures.
+where $g$ is the standard gravity [m.s${}^{-2}$], which gives approximately the same values both for rock and fractures.
 
@@
 Distance: 200 m \\
-Radius: 15 cm \\
+Well radius: 15 cm \\
 Length of open part of wells: 60 m \\
@@
 
-On the surface of the artificial cylinders we shall extrapolate the value of piezometric head using the following idea: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
+On the surface of the artificial cylinders we shall extrapolate the value of piezometric head as follows: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
 \eq{\label{eq:H_log} H(\xx) = H_w\frac{\log d_w(|\xx|)}{\log r_w}. }
@@
 For simplicity we stimulate both wells simultaneously which produces very similar result in the change of fracture cross-section as if the wells were stimulated one by one.
-The liquid is injected for 1 day under the pressure 100 MPa (which corresponds to piezometric head $10^4$ m).
+The liquid is injected for 1 day under the pressure 10 MPa (which corresponds to piezometric head $10^3$ m).
 
@@
 \begin{itemize}
-\item piezometric head $10^4$ m on surface of wells (including the intersections with fractures)
+\item piezometric head $10^3$ m on surface of wells (including the intersections with fractures)
 \item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
 \item no fluid flow through bottom and top part of the box and fracture tips
-\item zero displacement on bottom part of the box
+\item zero displacement on bottom part of the box, zero displacement in the direction orthogonal to box sides
 \item zero traction elsewhere
@@
 By dimension reduction of the Biot equations we obtain fluxes and forces which couple the rock with the lower dimensional fractures:
-\[ \partial_t(Sp + \alpha\div\uu) - k\Delta p = 0,\qquad -\mu\div(\nabla\uu+\nabla\uu^\top) - \lambda\nabla\div\uu + \alpha\varrho g\nabla H = 0. \]
+\[ \partial_t(Sp + \alpha\div\uu) - k\Delta p = 0,\qquad -\mu\div(\nabla\uu+\nabla\uu^\top) - \lambda\nabla\div\uu + \alpha\varrho g\nabla p = 0. \]
 Here $p=H-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\mu$ and $\lambda$ are the Lamé parameters [Pa].
@@
 The second scenario corresponds to operation of the EGS within 30 years after hydraulic stimulation performed according to Sec. \ref{sc:hm_model}.
-The fluid is continuously injected into one well and pumped out of the second one under the pressure difference 20 MPa and with inlet temperature $15\ ^\circ$C.
+The fluid is continuously injected into one well and pumped out of the second one under the pressure difference 2 MPa and with inlet temperature $15\ ^\circ$C.
 
@@
 \begin{itemize}
-\item piezometric head $\pm10^3$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
+\item piezometric head $\pm10^2$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
 \item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)

### 2019-07-17 random-frac/process.py
first domain with random fractures

@@ -1,0 +1,273 @@
+import sys
+sys.path.append('../dfn/src')
+sys.path.append('../MLMC/src')
+
+import os
+import shutil
+import subprocess
+import yaml
+import attr
+import numpy as np
+# import matplotlib.pyplot as plt
+
+from gmsh_api import gmsh
+from gmsh_api import options
+import fracture
+
+@attr.s(auto_attribs=True)
+class ValueDesctription:
+    time: float
+    position: str
+    quantity: str
+    unit: str
+
+
+def substitute_placeholders(file_in, file_out, params):
+    """
+    Substitute for placeholders of format '<name>' from the dict 'params'.
+    :param file_in: Template file.
+    :param file_out: Values substituted.
+    :param params: { 'name': value, ...}
+    """
+    used_params = []
+    with open(file_in, 'r') as src:
+        text = src.read()
+    for name, value in params.items():
+        placeholder = '<%s>' % name
+        n_repl = text.count(placeholder)
+        if n_repl > 0:
+            used_params.append(name)
+            text = text.replace(placeholder, str(value))
+    with open(file_out, 'w') as dst:
+        dst.write(text)
+    return used_params
+
+
+def prepare_mesh(config_dict):
+    geom = config_dict["geometry"]
+    dimensions = geom["box_dimensions"]
+    well_z0, well_z1 = geom["well_openning"]
+    well_r = geom["well_effective_radius"]
+    well_dist = geom["well_distance"]
+
+    # generate fracture set
+    volume = np.product(dimensions)
+    pop = fracture.Population(volume)
+    pop.initialize(geom["fracture_stats"])
+    pop.set_sample_range([1, well_dist], max_sample_size=geom["n_frac_limit"])
+    print("total mean size: ", pop.mean_size())
+    fractures = pop.sample()
+
+    factory = gmsh.GeometryOCC("three_frac_symmetric", verbose=True)
+    gopt = options.Geometry()
+    gopt.Tolerance = 1e-5
+    gopt.ToleranceBoolean = 1e-3
+    # gopt.MatchMeshTolerance = 1e-1
+
+    # Main box
+    box = factory.box(dimensions).set_region("box")
+    side_z = factory.rectangle([dimensions[0], dimensions[1]])
+    side_y = factory.rectangle([dimensions[0], dimensions[2]])
[…]

### 2019-07-18 random-frac/process.py
random-frac update

@@ -8,18 +8,22 @@ import subprocess
 import yaml
 import attr
 import numpy as np
+import collections
 # import matplotlib.pyplot as plt
 
-from gmsh_api import gmsh
-from gmsh_api import options
 import fracture
 
+# TODO:
+# - enforce creation of empty physical groups, or creation of empty regions in the flow input
+# - speedup mechanics
+
 @attr.s(auto_attribs=True)
-class ValueDesctription:
+class ValueDescription:
     time: float
     position: str
     quantity: str
     unit: str
+
 
 
 def substitute_placeholders(file_in, file_out, params):
@@ -42,23 +46,83 @@ def substitute_placeholders(file_in, file_out, params):
         dst.write(text)
     return used_params
 
+def to_polar(x, y, z):
+    rho = np.sqrt(x ** 2 + y ** 2)
+    phi = np.arctan2(y, x)
+    if z > 0:
+        phi += np.pi
+    return (phi, rho)
 
-def prepare_mesh(config_dict):
+def plot_fr_orientation(fractures):
+    family_dict = collections.defaultdict(list)
+    for fr in fractures:
+        x, y, z = fracture.FisherOrientation.rotate(np.array([0,0,1]), axis=fr.rotation_axis, angle=fr.rotation_angle)[0]
+        family_dict[fr.region].append([
+            to_polar(z, y, x),
+            to_polar(z, x, -y),
+            to_polar(y, x, z)
+            ])
+
+    import matplotlib.pyplot as plt
+    fig, axes = plt.subplots(1, 3, subplot_kw=dict(projection='polar'))
+    for name, data in family_dict.items():
+        # data shape = (N, 3, 2)
+        data = np.array(data)
+        for i, ax in enumerate(axes):
+            phi = data[:, i, 0]
+            r = data[:, i, 1]
+            c = ax.scatter(phi, r, cmap='hsv', alpha=0.75, label=name)
+    axes[0].set_title("X-view, Z-north")
+    axes[1].set_title("Y-view, Z-north")
+    axes[2].set_title("Z-view, Y-north")
+    for ax in axes:
+        ax.set_theta_zero_location("N")
+        ax.set_theta_direction(-1)
+        ax.set_ylim(0, 1)
+    fig.legend(loc = 1)
+    fig.savefig("fracture_orientation.pdf")
+    plt.close(fig)
+    #plt.show()
+
+def generate_fractures(config_dict):
     geom = config_dict["geometry"]
[…]

### 2019-07-18 paper.tex

@@
 Distance: 200 m \\
-Well radius: 15 cm \\
-Length of open part of wells: 60 m \\
+Well radius: $\rho = 15$ cm \\
+Length of open part of wells: $l=60$ m \\
 
@@
 
-In our model we consider the cube $600 \times 600 \times 600$ m around the lower well ends, from which two cylinders of radius 10 m and height 60 m are removed.
-These represent the surrounding of the open well parts.
-The reason for using 10 m radius instead of the real well radius is to keep the number of necessary mesh elements reasonable.
-The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+In our model, we consider pair of vertical wells with the radius $\rho=0.15$ m in the distance $D=200$ m.
+The open part of the length $L=60$ m is in the depth $5000$ m.
+The cube $600 \times 600 \times 600$ m around open part is considered as the computing domain.
 
+{\bf The test geometry:}The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+
+\subsection{Analytical models in the vicinity of the wells}
+In order to keep mesh size treatable, we replace the Dirichlet boundary condition for the pressure applied on the open part of the well by the Robin type condition:
+%
+\eq{\label{eq:H_robin} k\nabla H\cdot\nn = \sigma(H-H_w), }
+%
+applied on the cut-out cylinders of the radii $R=10$ m.
+We assume the constant conductivity $k$ in the vicinity of wells.
+The parameter $\sigma$ is determined from the analytical radially symmetric solution
+of the steady Darcy flow problem with a point source:
+\[
+    h(r) = -\alpha \log (r) + \beta
+\]
+where $r$ is the distance from the well and $\alpha$, $\beta$ are arbitrary constants.
+The flux through the cylinder of the radius $r$ and the length $L$ is the constant
+\[  
+    Q = -2\pi \rho L k h'(\rho) = 2\pi L k \alpha,
+\]
+where we have used the Darcy law with the conductivity $k$. Now we use the analytical solution to fit the Robin type boundary condition:
+\[
+    Q = 2\pi R l \sigma (h(\rho) - h(R)) = 2\pi R l \sigma \alpha \log(R/\rho),
+\]
+which combined with previous equality determines $\sigma$ as:
+\[ 
+    \sigma = \frac{k}{R\log(R/\rho)}.
+\]
+
+Since the heat transfer is dominated by the convection in the vicinity of the wells the temperature in the well is transported to the surface of the cut-off cylinders or vice verse in the time $t_0 = R/q$ which for the observed volume fluxes $Q\sim 1$ l/s is about $100$ days.
+This is negligible with respect to the total lifetime of the heat exchanger.
+
+{\bf Original JS idea:}
+This flux has to be equal to the flux through the surface of the extended well cylinders:
+\[
+    Q = 2\pi R l q,\quad q=\sigma(h_\rho - h)
+\]
 On the surface of the artificial cylinders we shall extrapolate the value of piezometric head as follows: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
@@
 Then, on the surface of the cylinder at radius $r_c$ we impose the Robin condition in the Darcy flow equation:
-\eq{\label{eq:H_robin} k\nabla H\cdot\nn = \sigma(H-H_w), }
 where $\nn$ stands for the unit outward normal vector.
@@
 \item piezometric head $10^3$ m on surface of wells (including the intersections with fractures)
-\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
+\item piezometric head 0 m on lateral sides of the box
 \item no fluid flow through bottom and top part of the box and fracture tips
@@
 \item piezometric head $\pm10^2$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
-\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
-\item no flux on bottom and top part of the box and on fracture tips
+\item no flux on the external boundary of the box and on fracture tips
 \item temperature $15\ ^\circ$C on surface of injection well (ZDE BY SE MELA UDELAT APROXIMACE DIRICHLETA ROBINEM PODOBNE JAKO V PROUDENI Z NEJAKEHO ANALYTICKEHO RESENI)

### 2019-07-18 paper.tex
Change notation of open part l to L.

@@
 Well radius: $\rho = 15$ cm \\
-Length of open part of wells: $l=60$ m \\
+Length of open part of wells: $L=60$ m \\
 
@@
 
-{\bf The test geometry:}The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+{\bf The test geometry:} The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
 
@@
 \[
-    Q = 2\pi R l \sigma (h(\rho) - h(R)) = 2\pi R l \sigma \alpha \log(R/\rho),
+    Q = 2\pi R L \sigma (h(\rho) - h(R)) = 2\pi R L \sigma \alpha \log(R/\rho),
 \]

### 2019-07-19 fixed-3-frac/process.py
Improve process script, change well sigma model.
- change well sigma model according to JB
- parametrize conductivity and cross section of fractures in config
- read mechanics output (cross section), compute conductivity and pass it to TH model

@@ -1,8 +1,15 @@
+import sys
+sys.path.append('../MLMC/src')
+
 import subprocess
 import yaml
 import attr
 import numpy as np
 import matplotlib.pyplot as plt
+import copy
+
+import MLMC.src.gmsh_io as gmsh_io
+
 
 @attr.s(auto_attribs=True)
 class ValueDesctription:
@@ -33,7 +40,6 @@ def substitute_placeholders(file_in, file_out, params):
     return used_params
 
 
-
 def compute_hm(config_dict):
     """
     :param config_dict: Parsed config.yaml. see key comments there.
@@ -44,12 +50,64 @@ def compute_hm(config_dict):
     print("Running: ", " ".join(arguments))
     subprocess.call(arguments)
 
+
 def prepare_th_input(config_dict):
     """
     Prepare FieldFE input file for the TH simulation.
     :param config_dict: Parsed config.yaml. see key comments there.
     """
-    pass
+    # pass
+    # we have to read region names from the input mesh
+    # input_mesh = gmsh_io.GmshIO(config_dict['hm_params']['mesh'])
+    #
+    # is_bc_region = {}
+    # for name, (id, _) in input_mesh.physical.items():
+    #     unquoted_name = name.strip("\"'")
+    #     is_bc_region[id] = (unquoted_name[0] == '.')
+
+    # read mesh and mechanichal output data
+    mesh = gmsh_io.GmshIO('output_hm/mechanics.msh')
+
+    n_bulk = len(mesh.elements)
+    ele_ids = np.zeros(n_bulk, dtype=int)
+    for i, id_bulk in zip(range(n_bulk), mesh.elements.items()):
+        ele_ids[i] = id_bulk[0]
+
+    init_fr_cs = float(config_dict['hm_params']['fr_cross_section'])
+    init_fr_K = float(config_dict['hm_params']['fr_conductivity'])
+    init_bulk_K = float(config_dict['hm_params']['bulk_conductivity'])
+
+    field_cs = mesh.element_data['cross_section_updated'][1]
+
+    K = np.zeros((n_bulk, 1), dtype=float)
+    cs = np.zeros((n_bulk, 1), dtype=float)
+    for i, valcs in zip(range(n_bulk), field_cs[1].values()):
+        cs_el = valcs[0]
+        cs[i, 0] = cs_el
+        if cs_el != 1.0:    # if cross_section == 1, i.e. 3d bulk
[…]

### 2019-07-20 random-frac/process.py
Both calculations running on random fracture mesh.
- several fixes
- mesh healing

@@ -101,6 +101,8 @@ def generate_fractures(config_dict):
     print("total mean size: ", pop.mean_size())
     pos_gen = fracture.UniformBoxPosition(fracture_box)
     fractures = pop.sample(pos_distr=pos_gen, keep_nonempty=True)
+    #fracture.fr_intersect(fractures)
+
     for fr in fractures:
         fr.region = "fr"
     used_families = set((f.region for f in fractures))
@@ -111,6 +113,43 @@ def generate_fractures(config_dict):
         model_dict["right_well_fracture_regions"] = [".{}_right_well".format(f) for f in used_families]
     return fractures
 
+
+def create_fractures_rectangles(gmsh_geom, fractures, base_shape: 'ObjectSet'):
+    # From given fracture date list 'fractures'.
+    # transform the base_shape to fracture objects
+    # fragment fractures by their intersections
+    # return dict: fracture.region -> GMSHobject with corresponding fracture fragments
+    shapes = []
+    for i, fr in enumerate(fractures):
+        shape = base_shape.copy()
+        print("fr: ", i, "tag: ", shape.dim_tags)
+        shape = shape.scale([fr.rx, fr.ry, 1]) \
+            .rotate(axis=fr.rotation_axis, angle=fr.rotation_angle) \
+            .translate(fr.centre) \
+            .set_region(fr.region)
+
+        shapes.append(shape)
+
+    fracture_fragments = gmsh_geom.fragment(*shapes)
+    return fracture_fragments
+
+
+def create_fractures_polygons(gmsh_geom, fractures):
+    # From given fracture date list 'fractures'.
+    # transform the base_shape to fracture objects
+    # fragment fractures by their intersections
+    # return dict: fracture.region -> GMSHobject with corresponding fracture fragments
+    frac_obj = fracture.Fractures(fractures)
+    frac_obj.snap_vertices_and_edges()
+    shapes = []
+    for fr, square in zip(fractures, frac_obj.squares):
+        shape = gmsh_geom.make_polygon(square).set_region(fr.region)
+        shapes.append(shape)
+
+    fracture_fragments = gmsh_geom.fragment(*shapes)
+    return fracture_fragments
+
+
 def prepare_mesh(config_dict, fractures):
[…]

### 2019-07-20 paper.tex

--- 2019-07-10 paper.tex
@@
 
+\newcommand{\CC}{\mathbb C}
 \renewcommand{\div}{\operatorname{div}}
 \newcommand{\eq}[1]{\begin{equation}#1\end{equation}}
+\newcommand{\II}{\mathbb I}
+\newcommand{\jmp}[1]{[\![#1]\!]}
 \newcommand{\nn}{\vc n}
@@
 
-Rock type: granite, gneiss or mica schist \cite{capova} % zula, rula, svor
+According to available geological data of the Litoměřice locality, the rock in the depth of several km is considered crystallinic, possibly consisting of granite, gneiss or mica schist \cite{capova}.
+% zula, rula, svor
+The mechanical properties of granitic rocks have been investigated e.g. in \cite{ljunggren}, where the measured Young modulus $E$ varies around 50 GPa and Poisson's ratio $\nu$ is around 0.25.
+The reported values of compressibility $\beta$ of several types of granite and orthogneiss are approx. $20\times 10^{-13}$ cm${}^2$.dyne${}^{-1}$ = $2\times10^{-11}$ Pa${}^{-1}$ \cite{zisman}.
+As far as porosity $\vartheta$ is concerned, \cite{intera} report values between 0.001 and 0.01 for metamorphic crystalline rocks, we shall consider the value 0.005.
+For water we use the density $\varrho_l=10^3$ kg.m${}^{-3}$, compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$.
+Here and in what follows, subscript ``l'' indicates quantities associated to the liquid.
+Similarly, subscript ``f'' will denote quantities for the fractures.
+Storativity $S$ is calculated from compressibility and porosity as follows (\cite{brace-et-al}, see also \cite{intera}):
+\[ S = \varrho_l g(\beta + \vartheta\beta_l). \]
+Here $g=9.81$ m.s${}^{-2}$ is the standard gravity.
+In our case we obtain $S\approx 2\times 10^{-7} \mbox{ m}^{-1}$.
 
-Liquid: water
+The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
+The hydraulic conductivity of fractures dominates than the one of intact rock by several orders of magnitude, we consider $k_f=10^{-3}$ m.s${}^{-1}$
+On the other hand, the elastic modulus of fracture is assumed much smaller, we take $E_f=50$ Pa.
+The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$.
+We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
 
+All values of physical parameters used in our model are listed in Table \ref{tab:params}.
+
+\begin{table}
+\centering
 \begin{tabular}{|l|c|c|c|}
@@
 Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ & ---\\
-Porosity $\vartheta$ [--] & $5\times10^{-3}$ \cite{intera} & $5\times10^{-2}$ & --- \\
-Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ \cite{zisman} & $2\times10^{-11}$ & $5\times10^{-10}$ \\
+Porosity $\vartheta$ [--] & $5\times10^{-3}$ & $5\times10^{-2}$ & --- \\
+Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ & $2\times10^{-11}$ & $5\times10^{-10}$ \\
 Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
@@
 \hline
-Young modulus $G$ [Pa] & $50\times10^9$ \cite{ljunggren} & 50 & --- \\
-Poisson's ratio $\nu$ [--] & 0.25 \cite{ljunggren} & 0.25 & --- \\
+Young modulus $E$ [Pa] & $50\times10^9$ & 50 & --- \\
+Poisson's ratio $\nu$ [--] & 0.25 & 0.25 & --- \\
 Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
@@
 \end{tabular}
+\caption{Physical parameters of rock, fractures and water, used in the model.}
+\label{tab:params}
+\end{table}
 
-Storativity is calculated as follows \cite{brace-et-al}:
-\[ S = \varrho_l g(\beta_r + \vartheta\beta_l), \]
-where $g$ is the standard gravity [m.s${}^{-2}$], which gives approximately the same values both for rock and fractures.
+Papers about effective properties of fractured porous media: \cite{Snow1969}, \cite{Bonnet2001}, \cite{Olson2003}
 
@@
 Distance: 200 m \\
-Well radius: 15 cm \\
-Length of open part of wells: 60 m \\
+Well radius: $\rho = 15$ cm \\
+Length of open part of wells: $L=60$ m \\
 
@@
 
-In our model we consider the cube $600 \times 600 \times 600$ m around the lower well ends, from which two cylinders of radius 10 m and height 60 m are removed.
-These represent the surrounding of the open well parts.
-The reason for using 10 m radius instead of the real well radius is to keep the number of necessary mesh elements reasonable.
-The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+In our model, we consider pair of vertical wells with the radius $\rho=0.15$ m in the distance $D=200$ m.
+The open part of the length $L=60$ m is in the depth $5000$ m.
+The cube $600 \times 600 \times 600$ m around the open parts is considered as the computational domain.
 
-On the surface of the artificial cylinders we shall extrapolate the value of piezometric head as follows: Given value $H_w$ of piezometric head at the well surface (with radius $r_w$) we assume that the piezometric head in the vicinity of the well is the function of the distance $d_w$ to the well axis:
-\eq{\label{eq:H_log} H(\xx) = H_w\frac{\log d_w(|\xx|)}{\log r_w}. }
-Then, on the surface of the cylinder at radius $r_c$ we impose the Robin condition in the Darcy flow equation:
-\eq{\label{eq:H_robin} k\nabla H\cdot\nn = \sigma(H-H_w), }
-where $\nn$ stands for the unit outward normal vector.
-From \eqref{eq:H_log} and \eqref{eq:H_robin} we obtain the transition coefficient
-\[ \sigma = \frac{k}{r_c}\left|\frac{\log r_c}{\log r_w}\right| \]
-whose value depends on the hydraulic conductivity.
+{\bf The test geometry:} The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+
+\subsection{Analytical models in the vicinity of the well}
+In order to keep the mesh size treatable, we cut out of the computational domain two cylinders around the open well parts with a larger radius $R=10$ m.
+On the surface of these cylinders we consider the Robin type boundary condition for the piezometric head:
+%
+\eq{\label{eq:H_robin} k\nabla h\cdot\nn = \sigma(H_w-h), }
+%
+where $\nn$ is the unit normal vector pointing to the centre of the well and $H_w$ is the piezometric head in the well.
+% We assume the constant conductivity $k$ in the vicinity of well.
+The parameter $\sigma$ is determined from the analytical radially symmetric solution
+of the steady Darcy flow problem with a point source:
+\[
+    \hat h(r) = -\alpha \log (r) + \beta,
+\]
+where $r$ is the distance from the well and $\alpha$, $\beta$ are arbitrary constants such that $\hat h(\rho)=H_w$:
+First, the flux from the well into the domain through the surface $C_\rho$ of the cylinder with radius $\rho$ and length $L$ is
+%  
+\eq{\label{eq:Q_rho}    Q_\rho := -\int_{C_\rho} k\nabla h\cdot\nn = -2\pi \rho L k \hat h'(\rho) = 2\pi L k \alpha. }
+%
+% where we have used the Darcy law with the conductivity $k$.
+%Now we use the analytical solution to fit the Robin type boundary condition:
+Next, the flux at the distance $R$ can be expressed from \eqref{eq:H_robin} as follows:
+%
+\eq{\label{eq:Q_R}   Q_R = 2\pi R L \sigma (\hat h(\rho) - \hat h(R)) = 2\pi R L \sigma \alpha \log(R/\rho).}
+%
+%which combined with previous equality determines $\sigma$ as:
+Finally, since $Q_\rho=Q_R$, we obtain from \eqref{eq:Q_rho} and \eqref{eq:Q_R}:
+\eq{ \label{eq:sigma} \sigma = \frac{k}{R\log(R/\rho)}. }
+In the simulations we shall use the hydraulic conductivity of the fractures in \eqref{eq:sigma} as the fractures have major influence on the flow in the vicinity of wells.
+
+Since the heat transfer is dominated by the convection in the vicinity of the wells, the temperature in the well is transported to the surface of the cut-off cylinders or vice verse in the time $t_0 = R/q$ which for the observed volume fluxes $Q\sim 1$ l/s is about $100$ days.
+This is negligible with respect to the total lifetime of the heat exchanger, hence in the simulation of heat we shall use Dirichlet boundary condition for the temperature.
 
@@
 In the first model scenario the fluid is injected to both wells in order to open the preexisting fractures.
-The result is the modified cross-section of fractures.
+This leads to increase in the cross-section of fractures, which is then used in the computation of long-term operation.
+As the HM model of 3d rock we use the Biot poroelasticity equations:
+\eq{\label{eq:biot} \partial_t(Sp + \alpha\div\uu) +\div\vc q = 0,\qquad -\div\bbsigma + \tilde\alpha\nabla p = 0. }
+Here $p=h-x_3$ is the pressure head [m], $\uu$ is the displacement [m] and $\tilde\alpha:=\alpha\varrho_l g$.
+The flux $\vc q$ [m.s${}^{-1}$] and the stress tensor $\bbsigma$ [Pa] are given by the Darcy and Hooke laws:
+\eq{ \vc q = - k\nabla p, \qquad \bbsigma = \CC(\nabla\uu) := \mu(\nabla\uu+\nabla\uu^\top) + \lambda(\II:\nabla\uu)\II, }
+where $\mu=E/(2(1+\nu))$ and $\lambda=E\nu/((1+\nu)(1-2\nu))$ are the Lamé parameters [Pa], $\II$ is the identity matrix and ``:'' denotes the scalar product of matrices.
+
+For the 2d fractures we apply a dimension reduction procedure, which was originally developed in \cite{martin-jaffre-roberts} for the Darcy flow.
+We introduce the unit normal vector $\vc\nu$ to the fracture plane with arbitrary but fixed orientation.
+For a quantity, say $p$, which is defined on both sides of the fracture, we define the jump $\jmp{p}:=p^+-p^-$, where $p^\pm$ is the trace of $p$ on the positive/negative side of the fracture, determined by the orientation of $\vc\nu$.
+Then, integrating \eqref{eq:biot} across fracture aperture results in the following system of equations:
+\eq{\label{eq:biot_f}\left.\begin{aligned}
+\delta\left(\partial_t\left(S_fp_f + \alpha_f\left(\div_\tau\uu_f+\tfrac1\delta\jmp{\uu}\cdot\vc\nu\right)\right) - k_f\Delta_\tau p_f\right) &= \jmp{F}, \\
+\delta\left(-\div_\tau\CC_f(\nabla_\tau\uu_f+\tfrac1\delta\vc\nu\otimes\jmp{\uu}) + \tilde\alpha_f\nabla_\tau p_f\right) &= \jmp{\vc Q}.
+\end{aligned}\right\} }
+Here $p_f$ and $\uu_f$ is the average pressure head and displacement in fractures, the subscript $\tau$ indicates tangential operators in the fracture plane, ``$\otimes$'' denotes the outer product and $\tilde\alpha_f:=\alpha_f\varrho_l g$.
+The expressions $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are sources and forces that arise from the approximation of derivatives in the direction orthogonal to fracture plane:
+\eq{ \left.\begin{aligned}
+F^\pm &:= \pm\frac2\delta k_f(p^\pm-p_f),\\
+\vc Q^\pm &:= \pm\frac2\delta\CC_f(\nabla_\tau\uu^\pm + \vc\nu\otimes(\uu^\pm-\uu_f))\vc\nu.
+%\pm\frac2\delta\left(\mu_f(\uu^\pm-\uu_f)+(\mu_f+\lambda_f)((\uu^\pm-\uu_f)\cdot\vc\nu)\vc\nu\right) +\mu_f(\nabla_\tau\uu^\pm)^\top\vc\nu + \lambda_f(\div_\tau\uu^\pm)\vc\nu \pm\tilde\alpha_f p_f\vc\nu.
+\end{aligned}\right\} }
+Systems \eqref{eq:biot} and \eqref{eq:biot_f} are complemented by the following interface conditions on the fractures:
+\eq{ k\nabla p^\pm\cdot\vc\nu = F^\pm,\qquad \left(\CC(\nabla\uu^\pm) - \tilde\alpha p^\pm\II\right)\vc\nu = \vc Q^\pm. }
+
 For simplicity we stimulate both wells simultaneously which produces very similar result in the change of fracture cross-section as if the wells were stimulated one by one.
@@
 \end{itemize}
-
-We use the Biot poroelasticity equations in the 3d continuum and 2d network of fractures.
-By dimension reduction of the Biot equations we obtain fluxes and forces which couple the rock with the lower dimensional fractures:
-\[ \partial_t(Sp + \alpha\div\uu) - k\Delta p = 0,\qquad -\mu\div(\nabla\uu+\nabla\uu^\top) - \lambda\nabla\div\uu + \alpha\varrho g\nabla p = 0. \]
-Here $p=H-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\mu$ and $\lambda$ are the Lamé parameters [Pa].
-...
 
@@
 \item piezometric head $\pm10^2$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
-\item piezometric head 0 m on lateral sides of the box (MOZNA BY MELO BYT I NAHORE A DOLE?)
-\item no flux on bottom and top part of the box and on fracture tips
-\item temperature $15\ ^\circ$C on surface of injection well (ZDE BY SE MELA UDELAT APROXIMACE DIRICHLETA ROBINEM PODOBNE JAKO V PROUDENI Z NEJAKEHO ANALYTICKEHO RESENI)
+\item no flux on the external boundary of the box and on fracture tips
+\item temperature $15\ ^\circ$C on surface of injection well
 \item initial temperature on lateral sides of the box where fluid flows into the box
@@
 \[ ... \]
+
+\section*{Acknowledgement}
+\begin{itemize}
+    \item[a)] RINGEN:
+    The research was supported by the Czech Ministry of Education, Youth and Sports under the project No. LM2015084.
+
+    \item[b)] RINGEN+:
+    The resrarch was supported by the project No. CZ.02.1.01/0.0/0.0/16\_013/0001792, co-funded by the EU Operational Programme ``Research, Development and Education''.
+\end{itemize}
 
@@
 Granite,  Basalt, and Tuff. Accession  DE83-011519, NTIS, Springfield, Virginia, 1983.
+
+\bibitem{martin-jaffre-roberts} V.
+Martin, J.
+Jaffré, J.
+E.
+Roberts.
+Modeling fractures and barriers as interfaces for flow in porous media.
+SIAM Journal on Scientific Computing 26(5):1667--1691, 2005.
 

### 2019-07-20 mlmc_fixed_frac/frac_simulation.py
mlmc fixed frac

@@ -1,0 +1,413 @@
+import os
+import sys
+import shutil
+import yaml
+import numpy as np
+import time as t
+import mlmc.sample as sample
+import copy
+import gmsh_io as gmsh_io
+from mlmc.simulation import Simulation
+
+
+def substitute_placeholders(file_in, file_out, params):
+    """
+    Substitute for placeholders of format '<name>' from the dict 'params'.
+    :param file_in: Template file.
+    :param file_out: Values substituted.
+    :param params: { 'name': value, ...}
+    """
+    used_params = []
+    with open(file_in, 'r') as src:
+        text = src.read()
+    for name, value in params.items():
+        placeholder = '<%s>' % name
+        n_repl = text.count(placeholder)
+        if n_repl > 0:
+            used_params.append(name)
+            text = text.replace(placeholder, str(value))
+    with open(file_out, 'w') as dst:
+        dst.write(text)
+
+    return used_params
+
+
+def force_mkdir(path, force=False):
+    """
+    Make directory 'path' with all parents,
+    remove the leaf dir recursively if it already exists.
+    :param path: path to directory
+    :param force: if dir already exists then remove it and create new one
+    :return: None
+    """
+    if force:
+        if os.path.isdir(path):
+            shutil.rmtree(path)
+    os.makedirs(path, mode=0o775, exist_ok=True)
+
+
+def load_config_dict():
+    with open("config.yaml", "r") as f:
+        return yaml.safe_load(f)
+
+
+class FracSimulation(Simulation):
+    HM_YAML_TEMPLATE = '01_hm_tmpl.yaml'
+    TH_YAML_TEMPLATE = '02_th_tmpl.yaml'
+
+    HM_YAML_FILE = "01_hm.yaml"
+    TH_YAML_FILE = "02_th.yaml"
+
+    def __init__(self, mesh_step, level_id=None, config=None, clean=False, parent_fine_sim=None):
+        """
+
+        :param config: configuration of the simulation, processed keys:
+            env - Environment object.
+            fields - FieldSet object
+            yaml_file: Template for main input file. Placeholders:
+                <mesh_file> - replaced by generated mesh
+                <FIELD> - for FIELD be name of any of `fields`, replaced by the FieldElementwise field with generated
+                 field input file and the field name for the component.
+                 (TODO: allow relative paths, not tested but should work)
+            geo_file: Path to the geometry file. (TODO: default is <yaml file base>.geo
[…]

### 2019-07-20 mlmc_fixed_frac/process.py

@@ -1,0 +1,220 @@
+import sys
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+src_path = os.path.dirname(os.path.abspath(__file__))
+sys.path.append('../MLMC/src')
+
+from mlmc import mlmc
+from mlmc import base_process
+from mlmc.estimate import Estimate, CompareLevels
+from mlmc.moments import Legendre, Monomial
+
+import pbs
+from frac_simulation import FracSimulation
+
+
+class Process(base_process.Process):
+    def run(self):
+        """
+        Run mlmc
+        :return: None
+        """
+        mc_samples = 2
+        os.makedirs(self.work_dir, mode=0o775, exist_ok=True)
+        self.n_moments = 10
+
+        mlmc_list = []
+        # Run one level Monte-Carlo method
+        for nl in [1]:  # , 2, 3, 4,5, 7, 9]:
+            mlmc = self.setup_config(nl, clean=True)
+            # self.n_sample_estimate(mlmc)
+            self.generate_jobs(mlmc, n_samples=[mc_samples])
+            mlmc_list.append(mlmc)
+
+        self.all_collect(mlmc_list)
+
+    def process(self):
+        """
+        Use collected data
+        :return: None
+        """
+        assert os.path.isdir(self.work_dir)
+        mlmc_est_list = []
+
+        for nl in [1]:  # high resolution fields
+            mlmc = self.setup_config(nl, clean=False)
+            # Use wrapper object for working with collected data
+            mlmc_est = Estimate(mlmc)
+            mlmc_est_list.append(mlmc_est)
+
+        cl = CompareLevels([mlmc],
+                           output_dir=src_path,
+                           quantity_name="Q [m/s]",
+                           moment_class=Monomial,
+                           log_scale=False,
+                           n_moments=3, )
+
+        self.plot_temp_power(mlmc_est)
+
+        # self.process_analysis(cl)
+
+    def set_environment_variables(self):
+        """
+        Set pbs config, flow123d, gmsh
+        :return: None
+        """
+        root_dir = os.path.abspath(self.work_dir)
+        while root_dir != '/':
+            root_dir, tail = os.path.split(root_dir)
+
+        self.pbs_config = dict(
+            job_weight=250000,  # max number of elements per job
+            n_cores=1,
+            n_nodes=1,
+            select_flags=['cgroups=cpuacct'],
+            mem='4gb',
+            queue='charon',
[…]

### 2019-07-20 mlmc_fixed_frac/submit.sh

@@ -1,0 +1,27 @@
+#!/bin/bash
+
+set -x
+
+py_script=`pwd`/$1
+pbs_script=`pwd`/$1.pbs
+script_path=${py_script%/*}
+
+output_prefix="mlmc"
+
+cat >$pbs_script <<EOF
+#!/bin/bash 
+#PBS -S /bin/bash
+#PBS -l select=1:ncpus=1:cgroups=cpuacct:mem=8GB -l walltime=48:00:00
+#PBS -q charon
+#PBS -N MLMC_frac
+#PBS -j oe
+
+cd ${script_path}
+module load python36-modules-gcc
+module load hdf5-1.10.0-gcc
+# module use /storage/praha1/home/jan-hybs/modules
+# module load flow123d
+python3.6 ${py_script} -r -k run /storage/liberec3-tul/home/martin_spetlik/WGC/mlmc_frac
+EOF
+
+qsub $pbs_script

### 2019-07-21 random-frac/process.py
Finer mesh to improve solution stability.

@@ -159,10 +159,10 @@ def prepare_mesh(config_dict, fractures):
     well_dist = geom["well_distance"]
     mesh_name = config_dict["mesh_name"]
     mesh_file = mesh_name + ".msh"
-    fracture_mesh_step = 20
+    fracture_mesh_step = 10
 
     if os.path.isfile(mesh_file):
-        return mesh_file
+        return mesh_name + "_healed.msh"
 
 
     from gmsh_api import gmsh
@@ -229,13 +229,13 @@ def prepare_mesh(config_dict, fractures):
     b_fr_right_well = b_fractures.select_by_intersect(b_right_well).modify_regions("{}_right_well")
     b_fractures = factory.group(b_fr_left_well, b_fr_right_well, b_fractures_box)
     mesh_groups = [*box_all, fractures_fr, b_fractures]
-    #fractures_fr.set_mesh_step(fracture_mesh_step)
+    fractures_fr.set_mesh_step(fracture_mesh_step)
 
     factory.keep_only(*mesh_groups)
     factory.remove_duplicate_entities()
     factory.write_brep()
 
-    min_el_size = fracture_mesh_step / 100
+    min_el_size = fracture_mesh_step / 10
     fracture_el_size = np.max(dimensions) / 20
     max_el_size = np.max(dimensions) / 8
 

### 2019-07-21 random-frac/process.py
Turn off removing of small shapes.

@@ -205,7 +205,7 @@ def prepare_mesh(config_dict, fractures):
     fractures = create_fractures_rectangles(factory, fractures, factory.rectangle())
     #fractures = create_fractures_polygons(factory, fractures)
     fractures_group = factory.group(*fractures)
-    fractures_group = fractures_group.remove_small_mass(fracture_mesh_step * fracture_mesh_step / 10)
+    #fractures_group = fractures_group.remove_small_mass(fracture_mesh_step * fracture_mesh_step / 10)
 
     # drilled box and its boundary
     box_drilled = box.cut(left_well, right_well)

### 2019-07-21 mlmc_fixed_frac/frac_simulation.py
extract result improvements + separate prepare_th_input

@@ -4,9 +4,10 @@ import shutil
 import yaml
 import numpy as np
 import time as t
-import mlmc.sample as sample
 import copy
+sys.path.append('../MLMC/src')
 import gmsh_io as gmsh_io
+import mlmc.sample as sample
 from mlmc.simulation import Simulation
 
 
@@ -105,6 +106,9 @@ class FracSimulation(Simulation):
         self.coarse_sim_set = False
         self.pbs_script = []
 
+        # Auxiliary param for extracting results
+        self.previous_length = 0
+
         self.config_dict = load_config_dict()
         super(Simulation, self).__init__()
 
@@ -141,7 +145,6 @@ class FracSimulation(Simulation):
         Store them separeted.
         :return:
         """
-        # @TODO: create random fractures and whole mesh here
         pass
 
     def compute_hm(self, sample_dir):
@@ -244,7 +247,7 @@ class FracSimulation(Simulation):
         """
         self.compute_hm(sample_dir)
         # All auxiliary methods must be run in pbs script
-        self.pbs_script.append("python3 frac_simulation.py {}".format(sample_dir)) #causes some errors
+        self.pbs_script.append("python3 {}/prepare_th_input.py {}".format(self.process_dir, sample_dir)) #causes some errors
         self.compute_th(sample_dir)
 
     def run_sim_sample(self, out_subdir):
@@ -268,7 +271,11 @@ class FracSimulation(Simulation):
         :param regions:
         :return: times list, list: for every region the array of value series
         """
-        data = yaml.safe_load(yaml_stream)['data']
+        content = yaml.safe_load(yaml_stream)
+        if content is None:
+            return None, None
+
+        data = content['data']
         times = set()
         reg_series = {reg: [] for reg in regions}
 
@@ -302,35 +309,50 @@ class FracSimulation(Simulation):
         if os.path.exists(os.path.join(sample_dir, "FINISHED")):
             if os.path.exists(water_balance_file) and os.path.exists(energy_balance_file) and os.path.exists(heat_region_stat):
                 # Sometimes content of files is not complete, sleep() seems to be workaround
-                t.sleep(60)
+                if self.previous_length == 0:
+                    t.sleep(60)
 
-                # extract the flux
-                bc_regions = ['.left_fr_left_well', '.left_well', '.right_fr_right_well', '.right_well']
-                out_regions = bc_regions[2:]
+                while True:
+                    # extract the flux
[…]

### 2019-07-21 mlmc_fixed_frac/prepare_th_input.py

@@ -1,0 +1,89 @@
+#!/usr/bin/env python3
+import sys
+import os
+
+WGC_DIR = "/storage/liberec3-tul/home/martin_spetlik/WGC"
+WORK_DIR = os.path.join(WGC_DIR, 'mlmc_frac')
+
+sys.path.append(os.path.join(WGC_DIR, 'MLMC/src'))
+sys.path.append(WORK_DIR)
+import yaml
+import numpy as np
+import gmsh_io
+
+
+def load_config_dict():
+    with open(os.path.join(WORK_DIR, "config.yaml"), "r") as f:
+        return yaml.safe_load(f)
+
+
+def prepare_th_input(sample_dir):
+    """
+    Prepare FieldFE input file for the TH simulation.
+    """
+
+    # #we have to read region names from the input mesh
+    # input_mesh = gmsh_io.GmshIO(config_dict['hm_params']['mesh'])
+    #
+    # is_bc_region = {}
+    # for name, (id, _) in input_mesh.physical.items():
+    #     unquoted_name = name.strip("\"'")
+    #     is_bc_region[id] = (unquoted_name[0] == '.')
+
+    config_dict = load_config_dict()
+
+    # read mesh and mechanical output data
+    mesh = gmsh_io.GmshIO(os.path.join(sample_dir, 'output_hm/mechanics.msh'))
+
+    n_bulk = len(mesh.elements)
+    ele_ids = np.zeros(n_bulk, dtype=int)
+    for i, id_bulk in zip(range(n_bulk), mesh.elements.items()):
+        ele_ids[i] = id_bulk[0]
+
+    init_fr_cs = float(config_dict['hm_params']['fr_cross_section'])
+    init_fr_K = float(config_dict['hm_params']['fr_conductivity'])
+    init_bulk_K = float(config_dict['hm_params']['bulk_conductivity'])
+
+    field_cs = mesh.element_data['cross_section_updated'][1]
+
+    K = np.zeros((n_bulk, 1), dtype=float)
+    cs = np.zeros((n_bulk, 1), dtype=float)
+    for i, valcs in zip(range(n_bulk), field_cs[1].values()):
+        cs_el = valcs[0]
+        cs[i, 0] = cs_el
+        if cs_el != 1.0:  # if cross_section == 1, i.e. 3d bulk
+            K[i, 0] = init_fr_K * (cs_el * cs_el) / (init_fr_cs * init_fr_cs)
+        else:
+            K[i, 0] = init_bulk_K
+
[…]

### 2019-07-21 mlmc_fixed_frac/process.py

@@ -48,12 +48,7 @@ class Process(base_process.Process):
             mlmc_est = Estimate(mlmc)
             mlmc_est_list.append(mlmc_est)
 
-        cl = CompareLevels([mlmc],
-                           output_dir=src_path,
-                           quantity_name="Q [m/s]",
-                           moment_class=Monomial,
-                           log_scale=False,
-                           n_moments=3, )
+        #self.plot_density(mlmc)
 
         self.plot_temp_power(mlmc_est)
 
@@ -149,9 +144,23 @@ class Process(base_process.Process):
             mlmc_obj.load_from_file()
         return mlmc_obj
 
-    def _plot_density(self, mlmc_est):
-        # @TODO: plot densities for temp a power at given time
-        pass
+    def plot_density(self, mlmc):
+        # for level in mlmc.levels:
+        #     for f_sample, c_sample in level.collected_samples:
+        #         print("f sample ", f_sample.result_data)
+        mlmc.select_values({"power_time": (0, "=")}, selected_param="power")
+        # for level in mlmc.levels:
+        #     for f_sample, c_sample in level.collected_samples:
+        #         print("f sample ", f_sample.result)
+
+        cl = CompareLevels([mlmc],
+                           output_dir=src_path,
+                           quantity_name="Q [m/s]",
+                           moment_class=Legendre,
+                           log_scale=False,
+                           n_moments=8, )
+        cl.construct_densities()
+        cl.plot_densities()
 
     def get_all_results_by_param(self, mlmc_est, param_name):
         """
@@ -161,10 +170,12 @@ class Process(base_process.Process):
         :return: moments means, moments vars -> two numpy arrays
         """
         n_moments = 3
+        mlmc_est.mlmc.clean_select()
         mlmc_est.mlmc.select_values({param_name: (0, ">=")})
         domain = Estimate.estimate_domain(mlmc_est.mlmc)
         moments_fn = Monomial(n_moments, domain, False, ref_domain=domain)
         means, vars = mlmc_est.estimate_moments(moments_fn)
+
         return means, vars
 
     def plot_temp_power(self, mlmc_est):
@@ -173,8 +184,6 @@ class Process(base_process.Process):
         :param mlmc_est: mlmc.Estimate instance
         :return: None
         """
-        #self._plot_density(mlmc_est)
-
         times_means, times_vars = self.get_all_results_by_param(mlmc_est, "value")
         temp_times = times_means[:, 1]
 

### 2019-07-22 random-frac/process.py
Improved random_frac
adopt model changes from fixed-3-frac:
- model for vicinity of wells
- creation of th_input.msh

@@ -9,6 +9,7 @@ import yaml
 import attr
 import numpy as np
 import collections
+import gmsh_io
 # import matplotlib.pyplot as plt
 
 import fracture
@@ -301,6 +302,21 @@ def smooth_grad_error_indicator_3d(nodes):
     return quality, edges[i_min_edge]
 
 
+def flat_indicator_3d(nodes):
+    edges = [[0, 1], [0, 2], [0, 3], [1, 2], [1, 3], [2, 3]]
+    max_edge = np.sum([np.linalg.norm(nodes[i]- nodes[j]) for i,j in edges])
+    quality = np.abs(tet_measure(nodes)) / (max_edge **3 / 6 / np.sqrt(2))
+    return quality
+
+
+def skew_line_dist(nodes):
+    a_vec = nodes[1] - nodes[0]
+    b_vec = nodes[3] - nodes[2]
+    normal = np.cross(a_vec, b_vec)
+    normal /= np.linalg.norm(normal)
+    dist = normal @ (nodes[0] - nodes[1])
+    return dist, normal
+
 def check_element(mesh, eid):
     """
     Check element, possibly contract its shortest edge.
@@ -320,6 +336,19 @@ def check_element(mesh, eid):
         if quality < quality_tol:
             loc_node_a, loc_node_b = min_edge_nodes
             merge_nodes = (node_ids[loc_node_a], node_ids[loc_node_b])
+
+        if flat_indicator_3d(nodes) < quality_tol:
+            edges = [[0, 1, 2, 3], [0, 2, 1, 3], [0, 3, 1, 2]]
+            dists = [skew_line_dist(nodes[e]) for e in edges]
+            i_min = np.argmin([np.abs(d) for d,norm in dists])
+            shift = dists[i_min][0] * dists[i_min][1]
+            edge = edges[i_min]
+            nodes[edge[0]] += shift
+            nodes[edge[1]] += shift
+            nn = nodes[edge]
+
+
+
     elif len(nodes) == 3:
         quality, min_edge_nodes = smooth_grad_error_indicator_2d(nodes)
         #print(eid, "2dq: ", quality)
@@ -425,7 +454,56 @@ def prepare_th_input(config_dict):
     Prepare FieldFE input file for the TH simulation.
     :param config_dict: Parsed config.yaml. see key comments there.
     """
-    pass
+    # pass
+    # we have to read region names from the input mesh
+    # input_mesh = gmsh_io.GmshIO(config_dict['hm_params']['mesh'])
+    #
[…]

### 2019-07-22 random-frac/process.py
Remove incomplete healing improvement

@@ -337,15 +337,15 @@ def check_element(mesh, eid):
             loc_node_a, loc_node_b = min_edge_nodes
             merge_nodes = (node_ids[loc_node_a], node_ids[loc_node_b])
 
-        if flat_indicator_3d(nodes) < quality_tol:
-            edges = [[0, 1, 2, 3], [0, 2, 1, 3], [0, 3, 1, 2]]
-            dists = [skew_line_dist(nodes[e]) for e in edges]
-            i_min = np.argmin([np.abs(d) for d,norm in dists])
-            shift = dists[i_min][0] * dists[i_min][1]
-            edge = edges[i_min]
-            nodes[edge[0]] += shift
-            nodes[edge[1]] += shift
-            nn = nodes[edge]
+        # if flat_indicator_3d(nodes) < quality_tol:
+        #     edges = [[0, 1, 2, 3], [0, 2, 1, 3], [0, 3, 1, 2]]
+        #     dists = [skew_line_dist(nodes[e]) for e in edges]
+        #     i_min = np.argmin([np.abs(d) for d,norm in dists])
+        #     shift = dists[i_min][0] * dists[i_min][1]
+        #     edge = edges[i_min]
+        #     nodes[edge[0]] += shift
+        #     nodes[edge[1]] += shift
+        #     nn = nodes[edge]
 
 
 

### 2019-07-22 mlmc_random_frac/mlmc_process.py
mlmc random frac

@@ -1,0 +1,229 @@
+import sys
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+src_path = os.path.dirname(os.path.abspath(__file__))
+sys.path.append('../MLMC/src')
+
+from mlmc import mlmc
+from mlmc import base_process
+from mlmc.estimate import Estimate, CompareLevels
+from mlmc.moments import Legendre, Monomial
+
+import pbs
+from rand_frac_simulation import RandomFracSimulation
+
+
+class Process(base_process.Process):
+    def run(self):
+        """
+        Run mlmc
+        :return: None
+        """
+        mc_samples = 2
+        os.makedirs(self.work_dir, mode=0o775, exist_ok=True)
+        self.n_moments = 10
+
+        mlmc_list = []
+        # Run one level Monte-Carlo method
+        for nl in [1]:  # , 2, 3, 4,5, 7, 9]:
+            mlmc = self.setup_config(nl, clean=True)
+            # self.n_sample_estimate(mlmc)
+            self.generate_jobs(mlmc, n_samples=[mc_samples])
+            mlmc_list.append(mlmc)
+
+        self.all_collect(mlmc_list)
+
+    def process(self):
+        """
+        Use collected data
+        :return: None
+        """
+        assert os.path.isdir(self.work_dir)
+        mlmc_est_list = []
+
+        for nl in [1]:  # high resolution fields
+            mlmc = self.setup_config(nl, clean=False)
+            # Use wrapper object for working with collected data
+            mlmc_est = Estimate(mlmc)
+            mlmc_est_list.append(mlmc_est)
+
+        #self.plot_density(mlmc)
+
+        self.plot_temp_power(mlmc_est)
+
+        # self.process_analysis(cl)
+
+    def set_environment_variables(self):
+        """
+        Set pbs config, flow123d, gmsh
+        :return: None
+        """
+        root_dir = os.path.abspath(self.work_dir)
+        while root_dir != '/':
+            root_dir, tail = os.path.split(root_dir)
+
+        self.pbs_config = dict(
+            job_weight=250000,  # max number of elements per job
+            n_cores=3,
+            n_nodes=1,
+            select_flags=['cgroups=cpuacct'],
+            mem='8gb',
+            queue='charon',
+            home_dir='/storage/liberec3-tul/home/martin_spetlik/')
+
+        if tail == 'storage':
+            # Metacentrum
+            self.sample_sleep = 30
+            self.init_sample_timeout = 600
+            self.sample_timeout = 0
[…]

### 2019-07-22 mlmc_random_frac/pbs.py

@@ -1,0 +1,37 @@
+import os
+import os.path
+import shutil
+import subprocess
+import mlmc.pbs as pbs
+
+
+class Pbs(pbs.Pbs):
+    def add_realization(self, weight, lines, **kwargs):
+        """
+        Append new flow123d realization to the existing script content
+        :param weight: current simulation steps
+        :param kwargs: dict with params
+        :return: None
+        """
+        if self._number_of_realizations == 0:
+            self.clean_script()
+
+        assert self.pbs_script is not None
+
+        # lines = [
+        #     'cd {work_dir}',
+        #     'date +%y.%m.%d_%H:%M:%S',
+        #     'time -p {flow123d} --yaml_balance -i {output_subdir} -s {work_dir}/flow_input.yaml  -o {output_subdir} >{work_dir}/{output_subdir}/flow.out',
+        #     'date +%y.%m.%d_%H:%M:%S',
+        #     'touch {output_subdir}/FINISHED',
+        #     'echo \\"Finished simulation:\\" \\"{flow123d}\\" \\"{work_dir}\\" \\"{output_subdir}\\"',
+        #     '']
+        lines = [line.format(**kwargs) for line in lines]
+        self.pbs_script.extend(lines)
+
+        self._number_of_realizations += 1
+        self._current_job_weight += weight
+        if self._current_job_weight > self.job_weight or self._number_of_realizations > self.max_realizations:
+            self.execute()
+
+        return self._pbs_config['job_name']

### 2019-07-22 mlmc_random_frac/process.py

@@ -1,0 +1,637 @@
+import sys
+import os
+WGC_DIR = "/home/martin/Documents/WGC"
+WORK_DIR = os.path.join(WGC_DIR, 'mlmc_random_frac')
+
+sys.path.append(os.path.join(WGC_DIR, 'MLMC/src'))
+sys.path.append(os.path.join(WGC_DIR, 'dfn/src'))
+
+import shutil
+import subprocess
+import yaml
+import attr
+import numpy as np
+import collections
+import gmsh_io
+# import matplotlib.pyplot as plt
+
+import fracture
+
+from gmsh_api import gmsh
+from gmsh_api import options
+
+# TODO:
+# - enforce creation of empty physical groups, or creation of empty regions in the flow input
+# - speedup mechanics
+
+
+def substitute_placeholders(file_in, file_out, params):
+    """
+    Substitute for placeholders of format '<name>' from the dict 'params'.
+    :param file_in: Template file.
+    :param file_out: Values substituted.
+    :param params: { 'name': value, ...}
+    """
+    used_params = []
+    with open(file_in, 'r') as src:
+        text = src.read()
+    for name, value in params.items():
+        placeholder = '<%s>' % name
+        n_repl = text.count(placeholder)
+        if n_repl > 0:
+            used_params.append(name)
+            text = text.replace(placeholder, str(value))
+    with open(file_out, 'w') as dst:
+        dst.write(text)
+    return used_params
+
+
+def to_polar(x, y, z):
+    rho = np.sqrt(x ** 2 + y ** 2)
+    phi = np.arctan2(y, x)
+    if z > 0:
+        phi += np.pi
+    return (phi, rho)
+
+
+def plot_fr_orientation(fractures):
+    family_dict = collections.defaultdict(list)
+    for fr in fractures:
+        x, y, z = fracture.FisherOrientation.rotate(np.array([0,0,1]), axis=fr.rotation_axis, angle=fr.rotation_angle)[0]
+        family_dict[fr.region].append([
+            to_polar(z, y, x),
+            to_polar(z, x, -y),
+            to_polar(y, x, z)
+            ])
+
+    import matplotlib.pyplot as plt
+    fig, axes = plt.subplots(1, 3, subplot_kw=dict(projection='polar'))
+    for name, data in family_dict.items():
[…]

### 2019-07-22 mlmc_random_frac/rand_frac_simulation.py

@@ -1,0 +1,261 @@
+import os
+import sys
+import shutil
+import yaml
+import numpy as np
+import time as t
+
+sys.path.append('../MLMC/src')
+import mlmc.sample as sample
+from mlmc.simulation import Simulation
+
+
+def force_mkdir(path, force=False):
+    """
+    Make directory 'path' with all parents,
+    remove the leaf dir recursively if it already exists.
+    :param path: path to directory
+    :param force: if dir already exists then remove it and create new one
+    :return: None
+    """
+    if force:
+        if os.path.isdir(path):
+            shutil.rmtree(path)
+    os.makedirs(path, mode=0o775, exist_ok=True)
+
+
+def load_config_dict():
+    with open("config.yaml", "r") as f:
+        return yaml.safe_load(f)
+
+
+class RandomFracSimulation(Simulation):
+
+    def __init__(self, mesh_step, level_id=None, config=None, clean=False, parent_fine_sim=None):
+        """
+
+        :param config: configuration of the simulation, processed keys:
+            env - Environment object.
+            fields - FieldSet object
+            yaml_file: Template for main input file. Placeholders:
+                <mesh_file> - replaced by generated mesh
+                <FIELD> - for FIELD be name of any of `fields`, replaced by the FieldElementwise field with generated
+                 field input file and the field name for the component.
+                 (TODO: allow relative paths, not tested but should work)
+            geo_file: Path to the geometry file. (TODO: default is <yaml file base>.geo
+        :param mesh_step: Mesh step, decrease with increasing MC Level.
+        :param parent_fine_sim: Allow to set the fine simulation on previous level (Sim_f_l) which corresponds
+        to 'self' (Sim_c_l+1) as a coarse simulation. Usually Sim_f_l and Sim_c_l+1 are same simulations, but
+        these need to be different for advanced generation of samples (zero-mean control and antithetic).
+        """
+        if level_id is not None:
+            self.sim_id = level_id
+        else:
+            self.sim_id = RandomFracSimulation.total_sim_id
+            RandomFracSimulation.total_sim_id += 1
+
+        self.env = config['env']
+        self.step = mesh_step
+        # Pbs script creater
+        self.pbs_creater = self.env["pbs"]
+
+        # @TODO: set n fine elements
+        self.n_fine_elements = 0
+
+        # Prepare base workdir for this mesh_step
+        output_dir = config['output_dir']
+
[…]

### 2019-07-24 mlmc_random_frac/process.py
mlmc random frac fix extract

@@ -1,6 +1,6 @@
 import sys
 import os
-WGC_DIR = "/home/martin/Documents/WGC"
+WGC_DIR = "/storage/liberec3-tul/home/martin_spetlik/WGC_rand"
 WORK_DIR = os.path.join(WGC_DIR, 'mlmc_random_frac')
 
 sys.path.append(os.path.join(WGC_DIR, 'MLMC/src'))
@@ -335,15 +335,15 @@ def check_element(mesh, eid):
             loc_node_a, loc_node_b = min_edge_nodes
             merge_nodes = (node_ids[loc_node_a], node_ids[loc_node_b])
 
-        if flat_indicator_3d(nodes) < quality_tol:
-            edges = [[0, 1, 2, 3], [0, 2, 1, 3], [0, 3, 1, 2]]
-            dists = [skew_line_dist(nodes[e]) for e in edges]
-            i_min = np.argmin([np.abs(d) for d,norm in dists])
-            shift = dists[i_min][0] * dists[i_min][1]
-            edge = edges[i_min]
-            nodes[edge[0]] += shift
-            nodes[edge[1]] += shift
-            nn = nodes[edge]
+        # if flat_indicator_3d(nodes) < quality_tol:
+        #     edges = [[0, 1, 2, 3], [0, 2, 1, 3], [0, 3, 1, 2]]
+        #     dists = [skew_line_dist(nodes[e]) for e in edges]
+        #     i_min = np.argmin([np.abs(d) for d,norm in dists])
+        #     shift = dists[i_min][0] * dists[i_min][1]
+        #     edge = edges[i_min]
+        #     nodes[edge[0]] += shift
+        #     nodes[edge[1]] += shift
+        #     nn = nodes[edge]
 
     elif len(nodes) == 3:
         quality, min_edge_nodes = smooth_grad_error_indicator_2d(nodes)
@@ -447,7 +447,7 @@ def call_flow(config_dict, param_key, sample_dir, result_files):
     return status
 
 
-def prepare_th_input(config_dict):
+def prepare_th_input(config_dict, sample_dir):
     """
     Prepare FieldFE input file for the TH simulation.
     :param config_dict: Parsed config.yaml. see key comments there.
@@ -485,7 +485,7 @@ def prepare_th_input(config_dict):
 
     # mesh.write_fields('output_hm/th_input.msh', ele_ids, {'conductivity': K})
     th_input_file = 'th_input.msh'
-    with open(th_input_file, "w") as fout:
[…]

### 2019-07-24 mlmc_random_frac/rand_frac_simulation.py

@@ -145,7 +145,8 @@ class RandomFracSimulation(Simulation):
         :return: None
         """
         # All auxiliary methods must be run in pbs script
-        self.pbs_script.append("python3.6 {}/process.py {}".format(self.process_dir, sample_dir))
+        self.pbs_script.append("pip3 install --user gmsh-sdk")
+        self.pbs_script.append("python3 {}/process.py {}".format(self.process_dir, sample_dir))
         self.pbs_script.append("touch {}/FINISHED".format(sample_dir))
 
     def run_sim_sample(self, out_subdir):
@@ -212,7 +213,7 @@ class RandomFracSimulation(Simulation):
 
                 while True:
                     # extract the flux
-                    bc_regions = ['.left_fr_left_well', '.left_well', '.right_fr_right_well', '.right_well']
+                    bc_regions = ['.fr_left_well', '.left_well', '.fr_right_well', '.right_well']
                     out_regions = bc_regions[2:]
 
                     with open(energy_balance_file, "r") as f:

### 2019-07-24 mlmc_random_frac/process.py
Fix path settings in mlmc random frac.

@@ -261,8 +261,8 @@ def prepare_mesh(config_dict, fractures, sample_dir):
 
     #factory.make_mesh(mesh_groups, dim=2)
     factory.make_mesh(mesh_groups)
-    factory.write_mesh(filename=os.path.join(sample_dir, mesh_name + "." + gmsh.MeshFormat.msh2.name), format=gmsh.MeshFormat.msh2)
-    os.rename(os.path.join(sample_dir, mesh_name + ".msh2"), mesh_file)
+    factory.write_mesh(filename=mesh_name + "." + gmsh.MeshFormat.msh2.name, format=gmsh.MeshFormat.msh2)
+    os.rename(mesh_name + ".msh2", mesh_file)
 
     healed_mesh = heal_mesh(mesh_file)
     #factory.show()
@@ -431,7 +431,7 @@ def call_flow(config_dict, param_key, sample_dir, result_files):
     substitute_placeholders(os.path.join(sample_dir, fname + '_tmpl.yaml'), os.path.join(sample_dir, fname + '.yaml'), params)
     arguments = config_dict["_aux_flow_path"].copy()
 
-    output_dir = os.path.join(sample_dir, "output_" + fname)
+    output_dir = "output_" + fname
     config_dict[param_key]["output_dir"] = output_dir
 
     if all([os.path.isfile(os.path.join(output_dir, f)) for f in result_files]):
@@ -462,7 +462,7 @@ def prepare_th_input(config_dict, sample_dir):
     #     is_bc_region[id] = (unquoted_name[0] == '.')
 
     # read mesh and mechanichal output data
-    mechanics_output = os.path.join(config_dict['hm_params']["output_dir"], 'mechanics.msh')
+    mechanics_output = os.path.join(sample_dir,config_dict['hm_params']["output_dir"], 'mechanics.msh')
     mesh = gmsh_io.GmshIO(mechanics_output)
 
     n_bulk = len(mesh.elements)
@@ -606,9 +606,11 @@ def sample(sample_dir, config_dict):
 
     fractures = generate_fractures(config_dict)
     # plot_fr_orientation(fractures)
+    
     healed_mesh = prepare_mesh(config_dict, fractures, sample_dir)
-    config_dict["hm_params"]["mesh"] = healed_mesh
-    config_dict["th_params"]["mesh"] = healed_mesh
+    healed_mesh_bn = os.path.basename(healed_mesh)
[…]

### 2019-07-24 mlmc_random_frac/rand_frac_simulation.py
Improve pbs python settings, fix some dtypes. - add script for setting the environment - use the environment in the pbs script - fix dtypes in extract results

@@ -145,8 +145,9 @@ class RandomFracSimulation(Simulation):
         :return: None
         """
         # All auxiliary methods must be run in pbs script
-        self.pbs_script.append("pip3 install --user gmsh-sdk")
-        self.pbs_script.append("python3 {}/process.py {}".format(self.process_dir, sample_dir))
+        #self.pbs_script.append("pip3 install --user gmsh-sdk")
+        self.pbs_script.append("source {}/env/bin/activate".format(self.process_dir))
+        self.pbs_script.append("python {}/process.py {}".format(self.process_dir, sample_dir))
         self.pbs_script.append("touch {}/FINISHED".format(sample_dir))
 
     def run_sim_sample(self, out_subdir):
@@ -186,8 +187,8 @@ class RandomFracSimulation(Simulation):
                 reg_series[region].append(power_in_time)
         times = list(times)
         times.sort()
-        series = [np.array(region_series) for region_series in reg_series.values()]
-        return np.array(times), series
+        series = [np.array(region_series, dtype=float) for region_series in reg_series.values()]
+        return np.array(times, dtype=float), series
 
     def _extract_result(self, sample):
         """

### 2019-07-24 mlmc_random_frac/setup_python_environment.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+echo "Make sure you called 'source load_modules.sh' before!"
+echo "Creating python environment.."
+python3.6 -m venv env
+
+source env/bin/activate
+python --version
+which python
+which pip
+pip install --upgrade pip
+pip -V
+
+pip install gmsh-sdk
+
+#pip freeze
+deactivate
+

### 2019-07-24 mlmc_random_frac/process.py
remove sample_dir from process.py

@@ -1,10 +1,10 @@
 import sys
 import os
-WGC_DIR = "/storage/liberec3-tul/home/martin_spetlik/WGC_rand"
-WORK_DIR = os.path.join(WGC_DIR, 'mlmc_random_frac')
-
-sys.path.append(os.path.join(WGC_DIR, 'MLMC/src'))
-sys.path.append(os.path.join(WGC_DIR, 'dfn/src'))
+#WGC_DIR = "/storage/liberec3-tul/home/martin_spetlik/WGC_rand"
+#WORK_DIR = os.path.join(WGC_DIR, 'mlmc_random_frac')
+script_dir = os.path.realpath(__file__)
+sys.path.append(os.path.join(script_dir, '../MLMC/src'))
+sys.path.append(os.path.join(script_dir, '../dfn/src'))
 
 import shutil
 import subprocess
@@ -152,7 +152,7 @@ def create_fractures_polygons(gmsh_geom, fractures):
     return fracture_fragments
 
 
-def prepare_mesh(config_dict, fractures, sample_dir):
+def prepare_mesh(config_dict, fractures):
     geom = config_dict["geometry"]
     dimensions = geom["box_dimensions"]
     well_z0, well_z1 = geom["well_openning"]
@@ -160,11 +160,11 @@ def prepare_mesh(config_dict, fractures, sample_dir):
     well_r = geom["well_effective_radius"]
     well_dist = geom["well_distance"]
     mesh_name = config_dict["mesh_name"]
-    mesh_file = os.path.join(sample_dir, mesh_name + ".msh")
+    mesh_file = mesh_name + ".msh"
     fracture_mesh_step = 10
 
     if os.path.isfile(mesh_file):
-        return os.path.join(sample_dir, mesh_name + "_healed.msh")
+        return mesh_name + "_healed.msh"
 
     factory = gmsh.GeometryOCC(mesh_name, verbose=True)
     gopt = options.Geometry()
@@ -420,7 +420,7 @@ def heal_mesh(mesh_file):
     return healed_name
 
 
-def call_flow(config_dict, param_key, sample_dir, result_files):
+def call_flow(config_dict, param_key, result_files):
     """
     Redirect sstdout and sterr, return true on succesfull run.
     :param arguments:
@@ -428,7 +428,7 @@ def call_flow(config_dict, param_key, sample_dir, result_files):
     """
     params = config_dict[param_key]
     fname = params["in_file"]
[…]

### 2019-07-24 random-frac/process.py
random-frac heal mesh module
- separate module
- healing by edge contraction
- mesh statistics

@@ -151,20 +151,14 @@ def create_fractures_polygons(gmsh_geom, fractures):
     return fracture_fragments
 
 
-def prepare_mesh(config_dict, fractures):
+def make_mesh(config_dict, fractures, mesh_name, mesh_file):
+    fracture_mesh_step = 10
     geom = config_dict["geometry"]
     dimensions = geom["box_dimensions"]
     well_z0, well_z1 = geom["well_openning"]
     well_length = well_z1 - well_z0
     well_r = geom["well_effective_radius"]
     well_dist = geom["well_distance"]
-    mesh_name = config_dict["mesh_name"]
-    mesh_file = mesh_name + ".msh"
-    fracture_mesh_step = 10
-
-    if os.path.isfile(mesh_file):
-        return mesh_name + "_healed.msh"
-
 
     from gmsh_api import gmsh
     from gmsh_api import options
@@ -267,160 +261,23 @@ def prepare_mesh(config_dict, fractures):
     factory.write_mesh(format=gmsh.MeshFormat.msh2)
     os.rename(mesh_name + ".msh2", mesh_file)
 
-    healed_mesh = heal_mesh(mesh_file)
-    #factory.show()
-    return healed_mesh
 
+def prepare_mesh(config_dict, fractures):
+    mesh_name = config_dict["mesh_name"]
+    mesh_file = mesh_name + ".msh"
+    if not os.path.isfile(mesh_file):
+        make_mesh(config_dict, fractures, mesh_name, mesh_file)
 
-def tri_measure(nodes):
-    return np.linalg.norm(np.cross(nodes[2] - nodes[0], nodes[1] - nodes[0]))/2
+    mesh_healed = mesh_name + "_healed.msh"
+    if True: #not os.path.isfile(mesh_healed):
+        import heal_mesh
+        hm = heal_mesh.HealMesh(mesh_file)
+        hm.heal_small_edges(0.001)
+        hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
+        hm.write()
+        assert hm.healed_mesh_name == mesh_healed
+    return mesh_healed
 
-
-def smooth_grad_error_indicator_2d(nodes):
-    edges = [(0,1), (1,2), (2,0)]
-    e_lens = [np.linalg.norm(nodes[i]- nodes[j]) for i,j in edges]
-    i_min_edge = np.argmin(e_lens)
-    prod = max(1e-300, (np.prod(e_lens)) ** (2.0 / 3.0))
[…]

### 2019-07-24 mlmc_random_frac/mlmc_process.py
move mesh healing into dfn package/submodule

@@ -3,7 +3,8 @@ import os
 import numpy as np
 import matplotlib.pyplot as plt
 src_path = os.path.dirname(os.path.abspath(__file__))
-sys.path.append('../MLMC/src')
+sys.path.append(os.path.join(src_path, '../MLMC/src'))
+sys.path.append(os.path.join(src_path, '../dfn/src'))
 
 from mlmc import mlmc
 from mlmc import base_process

### 2019-07-24 random-frac/heal_mesh.py

@@ -1,374 +1,0 @@
-import os
-import collections
-import numpy as np
-
-
-
-
-class Tetrahedron:
-    vtxs_faces = [[0, 1, 2], [0, 1, 3], [0, 2, 3], [1, 2, 3]]
-    edges = [[0, 1], [0, 2], [0, 3], [2, 3], [1, 3], [1, 2]]
-
-    def __init__(self, nodes):
-        self.dim = 3
-        self.nodes = nodes
-        self._measure = None
-        self._face_areas = None
-        self._edge_lens = None
-
-    @property
-    def measure(self):
-        if not self._measure:
-            self._measure = np.linalg.det(self.nodes[1:, :] - self.nodes[0, :]) / 6
-        return self._measure
-
-    @property
-    def face_areas(self):
-        if not self._face_areas:
-            self._face_areas = [Triangle(self.nodes[face_vtxs]).measure for face_vtxs in self.vtxs_faces]
-        return self._face_areas
-
-    @property
-    def edge_lens(self):
-        if not self._edge_lens:
-            self._edge_lens = [np.linalg.norm(self.nodes[i] - self.nodes[j]) for i,j in self.edges]
-        return self._edge_lens
-
-    def smooth_grad_error_indicator(self):
-        faces = self.face_areas
-        e_lens = self.edge_lens
-        e_faces = [[0, 1], [0, 2], [1, 2], [0, 3], [1, 3], [2, 3]] # faces joining the edge
-        sum_pairs = max(1e-300, np.sum([faces[i] * faces[j] * elen ** 2 for (i,j),elen in zip(e_faces, e_lens)]))
-        regular = (2.0 * np.sqrt(2.0 / 3.0) / 9.0)
-        quality = np.abs(self.measure) * (np.sum(faces) / sum_pairs) ** (3.0/4.0) / regular
-        return quality
-
-    def gmsh_gamma(self):
-        faces = self.face_areas
-        V = np.abs(self.measure)
-        # inradius
-        r = 3*V / sum(faces)
-        # circum radius
-        a,b,c,A,B,C = self.edge_lens
-        x_area = (a*A + b*B + c*C)*(a*A + b*B - c*C)*(a*A - b*B + c*C)*(-a*A + b*B + c*C)
-        assert x_area > 0, x_area
-        R = np.sqrt(x_area) / 24 / V
-        return 3 * r/max(1e-300, R)
-
-
-    def flat_indicator(self):
[…]

### 2019-07-24 random-frac/process.py

@@ -1,6 +1,8 @@
 import sys
-sys.path.append('../dfn/src')
-sys.path.append('../MLMC/src')
+import os
+src_path = os.path.dirname(os.path.abspath(__file__))
+sys.path.append(os.path.join(src_path, '../MLMC/src'))
+sys.path.append(os.path.join(src_path, '../dfn/src'))
 
 import os
 import shutil

### 2019-07-24 random-frac/test_heal_mesh.py

@@ -1,20 +1,0 @@
-import heal_mesh
-import numpy as np
-
-def test_check_flat_element():
-    h = 0.01
-    nodes = np.array([[0,0,0], [1,1,0], [0,1,h], [1,0,h]], dtype=float)
-    ref_nodes = np.array([[0,0,h/2], [1,1,h/2], [0,1, h/2], [1,0,h/2]], dtype=float)
-    result = heal_mesh.check_flat_element(nodes, 0.05)
-    edge, new_nodes, isec = result
-    assert edge == [0,1,2,3]
-    assert np.allclose(new_nodes, ref_nodes)
-    assert np.allclose(isec, np.array([0.5, 0.5, h/2]))
-
-    nodes = np.array([[0,0,0], [0,1,h], [1,1,0], [1,0,h]], dtype=float)
-    ref_nodes = np.array([[0,0,h/2], [0,1, h/2], [1,1,h/2], [1,0,h/2]], dtype=float)
-    result = heal_mesh.check_flat_element(nodes, 0.05)
-    edge, new_nodes, isec = result
-    assert edge == [0,2,1,3]
-    assert np.allclose(new_nodes, ref_nodes)
-    assert np.allclose(isec, np.array([0.5, 0.5, h/2]))

### 2019-07-24 mlmc_random_frac/process.py
Collect number of bad elements.

@@ -152,20 +152,17 @@ def create_fractures_polygons(gmsh_geom, fractures):
     return fracture_fragments
 
 
-def prepare_mesh(config_dict, fractures):
+def make_mesh(config_dict, fractures):
+    fracture_mesh_step = 10
     geom = config_dict["geometry"]
     dimensions = geom["box_dimensions"]
     well_z0, well_z1 = geom["well_openning"]
     well_length = well_z1 - well_z0
     well_r = geom["well_effective_radius"]
     well_dist = geom["well_distance"]
-    mesh_name = config_dict["mesh_name"]
-    mesh_file = mesh_name + ".msh"
-    fracture_mesh_step = 10
 
-    if os.path.isfile(mesh_file):
-        return mesh_name + "_healed.msh"
-
+    from gmsh_api import gmsh
+    from gmsh_api import options
     factory = gmsh.GeometryOCC(mesh_name, verbose=True)
     gopt = options.Geometry()
     gopt.Tolerance = 0.0001
@@ -269,155 +266,22 @@ def prepare_mesh(config_dict, fractures):
     return healed_mesh
 
 
-def tri_measure(nodes):
-    return np.linalg.norm(np.cross(nodes[2] - nodes[0], nodes[1] - nodes[0]))/2
+def prepare_mesh(config_dict, fractures):
+    mesh_name = config_dict["mesh_name"]
+    mesh_file = mesh_name + ".msh"
+    if not os.path.isfile(mesh_file):
+        make_mesh(config_dict, fractures, mesh_name, mesh_file)
 
+    mesh_healed = mesh_name + "_healed.msh"
+    if True: #not os.path.isfile(mesh_healed):
+        import heal_mesh
+        hm = heal_mesh.HealMesh(mesh_file)
+        hm.heal_small_edges(0.001)
+        hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
+        hm.write()
+        assert hm.healed_mesh_name == mesh_healed
+    return mesh_healed
 
-def smooth_grad_error_indicator_2d(nodes):
-    edges = [(0,1), (1,2), (2,0)]
-    e_lens = [np.linalg.norm(nodes[i]- nodes[j]) for i,j in edges]
-    i_min_edge = np.argmin(e_lens)
-    prod = max(1e-300, (np.prod(e_lens)) ** (2.0 / 3.0))
-    quality = 4 / np.sqrt(3) * np.abs(tri_measure(nodes)) / prod
-    return quality, edges[i_min_edge]
-
-
[…]

### 2019-07-24 mlmc_random_frac/rand_frac_simulation.py

@@ -198,16 +198,17 @@ class RandomFracSimulation(Simulation):
         abs_zero_temp = 273.15
         year_sec = 60 * 60 * 24 * 365
         # Sample result structure -> you can enlarge it and other stuff will be done automatically
-        self.result_struct = [["value", "power", "temp", "power_time"], ["f8", "f8", "f8", "f8"]]
+        self.result_struct = [["value", "power", "temp", "power_time", "n_bad_els"], ["f8", "f8", "f8", "f8", "u4"]]
 
         sample_dir = sample.directory
 
+        heal_stat_file = os.path.join(sample_dir, "random_fractures_heal_stats.yaml")
         water_balance_file = os.path.join(sample_dir, "output_02_th/water_balance.yaml")
         energy_balance_file = os.path.join(sample_dir, "output_02_th/energy_balance.yaml")
         heat_region_stat = os.path.join(sample_dir, "output_02_th/Heat_AdvectionDiffusion_region_stat.yaml")
 
         if os.path.exists(os.path.join(sample_dir, "FINISHED")):
-            if os.path.exists(water_balance_file) and os.path.exists(energy_balance_file) and os.path.exists(heat_region_stat):
+            if os.path.exists(heal_stat_file) and os.path.exists(water_balance_file) and os.path.exists(energy_balance_file) and os.path.exists(heat_region_stat):
                 # Sometimes content of files is not complete, sleep() seems to be workaround
                 if self.previous_length == 0:
                     t.sleep(60)
@@ -217,10 +218,13 @@ class RandomFracSimulation(Simulation):
                     bc_regions = ['.fr_left_well', '.left_well', '.fr_right_well', '.right_well']
                     out_regions = bc_regions[2:]
 
+                    with open(heal_stat_file, "r") as f:
+                        stat_doc = yaml.safe_load(f)
+                        n_bad_els = len(stat_doc["flow_stats"]["bad_elements"]) + len(stat_doc["gamma_stats"]["bad_elements"])
+
                     with open(energy_balance_file, "r") as f:
                         power_times, reg_powers = self._extract_time_series(f, bc_regions,
                                                                             extract=lambda frame: frame['data'][0])
-
                     with open(heat_region_stat, "r") as f:
                         temp_times, reg_temps = self._extract_time_series(f, out_regions,
                                                                           extract=lambda frame: frame['average'][0])
@@ -253,7 +257,7 @@ class RandomFracSimulation(Simulation):
 
                 result_values = []
[…]

### 2019-07-24 mlmc_random_frac/rand_frac_simulation.py
Add min and max temperature to extract results.

@@ -198,7 +198,7 @@ class RandomFracSimulation(Simulation):
         abs_zero_temp = 273.15
         year_sec = 60 * 60 * 24 * 365
         # Sample result structure -> you can enlarge it and other stuff will be done automatically
-        self.result_struct = [["value", "power", "temp", "power_time", "n_bad_els"], ["f8", "f8", "f8", "f8", "u4"]]
+        self.result_struct = [["value", "power", "temp", "power_time", "temp_min", "temp_max", "n_bad_els"], ["f8", "f8", "f8", "f8",  "f8", "f8", "u4"]]
 
         sample_dir = sample.directory
 
@@ -216,6 +216,7 @@ class RandomFracSimulation(Simulation):
                 while True:
                     # extract the flux
                     bc_regions = ['.fr_left_well', '.left_well', '.fr_right_well', '.right_well']
+                    regions = ['fr', 'box']
                     out_regions = bc_regions[2:]
 
                     with open(heal_stat_file, "r") as f:
@@ -228,6 +229,12 @@ class RandomFracSimulation(Simulation):
                     with open(heat_region_stat, "r") as f:
                         temp_times, reg_temps = self._extract_time_series(f, out_regions,
                                                                           extract=lambda frame: frame['average'][0])
+                    with open(heat_region_stat, "r") as f:
+                        temp_min_times, reg_temp_min = self._extract_time_series(f, regions,
+                                                                                 extract=lambda frame: frame['min'][0])
+                    with open(heat_region_stat, "r") as f:
+                        temp_max_times, reg_temp_max = self._extract_time_series(f, regions,
+                                                                                 extract=lambda frame: frame['max'][0])
                     with open(water_balance_file, "r") as f:
                         flux_times, reg_fluxes = self._extract_time_series(f, out_regions,
                                                                            extract=lambda frame: frame['data'][0])
@@ -245,6 +252,14 @@ class RandomFracSimulation(Simulation):
                     power_times = power_times / year_sec
                     avg_temp = avg_temp - abs_zero_temp
 
+                    # compute min and max temperature over regions
+                    temp_min = np.full(len(power_times), 1e10)
+                    temp_max = np.full(len(power_times), -1e10)
+                    for j in range(0, len(power_times)):
+                        for i in range(0, len(regions)):
+                            temp_min[j] = min([temp_min[j], reg_temp_min[i][j]])
+                            temp_max[j] = max([temp_max[j], reg_temp_max[i][j]])
[…]

### 2019-07-24 mlmc_random_frac/mlmc_process.py
process result text output

@@ -49,6 +49,7 @@ class Process(base_process.Process):
             mlmc_est = Estimate(mlmc)
             mlmc_est_list.append(mlmc_est)
 
+        self.result_text(mlmc)
         #self.plot_density(mlmc)
 
         self.plot_temp_power(mlmc_est)
@@ -178,6 +179,16 @@ class Process(base_process.Process):
         means, vars = mlmc_est.estimate_moments(moments_fn)
 
         return means, vars
+
+    def result_text(self, mlmc):
+        for level in mlmc.levels:
+            for f_sample, c_sample in level.collected_samples:
+                print("Sample result data ", f_sample.result_data)
+                print("Sample temp result", f_sample.result_data["temp"])
+
+        mlmc.select_values(None, selected_param="temp")
+        for l in mlmc.levels:
+            print("Sample values ", l.sample_values)
 
     def plot_temp_power(self, mlmc_est):
         """

### 2019-07-24 mlmc_random_frac/process.py
Fix errors in process.

@@ -152,7 +152,7 @@ def create_fractures_polygons(gmsh_geom, fractures):
     return fracture_fragments
 
 
-def make_mesh(config_dict, fractures):
+def make_mesh(config_dict, fractures, mesh_name, mesh_file):
     fracture_mesh_step = 10
     geom = config_dict["geometry"]
     dimensions = geom["box_dimensions"]
@@ -260,10 +260,6 @@ def make_mesh(config_dict, fractures):
     factory.make_mesh(mesh_groups)
     factory.write_mesh(filename=mesh_name + "." + gmsh.MeshFormat.msh2.name, format=gmsh.MeshFormat.msh2)
     os.rename(mesh_name + ".msh2", mesh_file)
-
-    healed_mesh = heal_mesh(mesh_file)
-    #factory.show()
-    return healed_mesh
 
 
 def prepare_mesh(config_dict, fractures):

### 2019-07-24 mlmc_random_frac/mlmc_process.py
Move some mlmc process settings to config.

@@ -1,5 +1,6 @@
 import sys
 import os
+import yaml
 import numpy as np
 import matplotlib.pyplot as plt
 src_path = os.path.dirname(os.path.abspath(__file__))
@@ -15,13 +16,17 @@ import pbs
 from rand_frac_simulation import RandomFracSimulation
 
 
+def load_config_dict():
+    with open("config.yaml", "r") as f:
+        return yaml.safe_load(f)
+
+
 class Process(base_process.Process):
     def run(self):
         """
         Run mlmc
         :return: None
         """
-        mc_samples = 2
         os.makedirs(self.work_dir, mode=0o775, exist_ok=True)
         self.n_moments = 10
 
@@ -30,7 +35,7 @@ class Process(base_process.Process):
         for nl in [1]:  # , 2, 3, 4,5, 7, 9]:
             mlmc = self.setup_config(nl, clean=True)
             # self.n_sample_estimate(mlmc)
-            self.generate_jobs(mlmc, n_samples=[mc_samples])
+            self.generate_jobs(mlmc, n_samples=[self.mc_samples])
             mlmc_list.append(mlmc)
 
         self.all_collect(mlmc_list)
@@ -65,31 +70,26 @@ class Process(base_process.Process):
         while root_dir != '/':
             root_dir, tail = os.path.split(root_dir)
 
-        self.pbs_config = dict(
-            job_weight=250000,  # max number of elements per job
-            n_cores=3,
-            n_nodes=1,
-            select_flags=['cgroups=cpuacct'],
-            mem='8gb',
-            queue='charon',
-            home_dir='/storage/liberec3-tul/home/martin_spetlik/')
-
-        if tail == 'storage':
-            # Metacentrum
+        self.flow123d = self.config_dict["flow_executable"]
+        if(self.config_dict["metacentrum"]):
+            self.pbs_config = dict(
+                job_weight=250000,  # max number of elements per job
+                n_cores=3,
+                n_nodes=1,
+                select_flags=['cgroups=cpuacct'],
+                mem='8gb',
+                queue='charon')
             self.sample_sleep = 30
             self.init_sample_timeout = 600
             self.sample_timeout = 0
             self.pbs_config['qsub'] = '/usr/bin/qsub'
-            self.flow123d = 'flow123d'  # "/storage/praha1/home/jan_brezina/local/flow123d_2.2.0/flow123d"
[…]

### 2019-07-24 mlmc_random_frac/mlmc_process.py
Fix running on metacentrum.
- fix abs path for pbs script
- update config
- remove gmsh settings

@@ -128,7 +128,7 @@ class Process(base_process.Process):
         self.create_pbs_object(output_dir, clean)
 
         simulation_config = {
-            'env': dict(flow123d=self.flow123d, gmsh=self.gmsh, pbs=self.pbs_obj),  # The Environment.
+            'env': dict(flow123d=self.flow123d, pbs=self.pbs_obj),  # The Environment.
             'output_dir': output_dir,
             'sim_param_range': self.step_range,  # Range of MLMC simulation parametr. Here the mesh step.
         }

### 2019-07-24 mlmc_random_frac/rand_frac_simulation.py

@@ -146,6 +146,7 @@ class RandomFracSimulation(Simulation):
         """
         # All auxiliary methods must be run in pbs script
         #self.pbs_script.append("pip3 install --user gmsh-sdk")
+        self.pbs_script.append("cd {}".format(os.path.abspath(self.process_dir)))
         self.pbs_script.append("source {}/env/bin/activate".format(self.process_dir))
         self.pbs_script.append("python {}/process.py {}".format(self.process_dir, sample_dir))
         self.pbs_script.append("touch {}/FINISHED".format(sample_dir))

### 2019-07-25 mlmc_random_frac/mlmc_process.py
Enable running mlmc_process without pbs.

@@ -87,7 +87,6 @@ class Process(base_process.Process):
             self.sample_sleep = 1
             self.init_sample_timeout = 60
             self.sample_timeout = 60
-            self.pbs_config['qsub'] = None
 
         self.mc_samples = self.config_dict["mc_samples"]
 
@@ -125,7 +124,10 @@ class Process(base_process.Process):
             self.rm_files(output_dir)
 
         # Init pbs object
-        self.create_pbs_object(output_dir, clean)
+        if (self.config_dict["metacentrum"]):
+            self.create_pbs_object(output_dir, clean)
+        else:
+            self.pbs_obj = None
 
         simulation_config = {
             'env': dict(flow123d=self.flow123d, pbs=self.pbs_obj),  # The Environment.

### 2019-07-25 mlmc_random_frac/rand_frac_simulation.py
Fix order of result values.

@@ -273,8 +273,8 @@ class RandomFracSimulation(Simulation):
 
                 result_values = []
                 for i in range(len(power_times)):
-                    result_values.append((i, power_series[i], avg_temp[i], power_times[i], n_bad_els,
-                                          temp_min[i], temp_max[i]))
+                    result_values.append((i, power_series[i], avg_temp[i], power_times[i],
+                                          temp_min[i], temp_max[i], n_bad_els))
 
                 return result_values
             else:

### 2019-07-25 mlmc_random_frac/rand_frac_simulation.py
Change n_bad_els to float.

@@ -199,7 +199,7 @@ class RandomFracSimulation(Simulation):
         abs_zero_temp = 273.15
         year_sec = 60 * 60 * 24 * 365
         # Sample result structure -> you can enlarge it and other stuff will be done automatically
-        self.result_struct = [["value", "power", "temp", "power_time", "temp_min", "temp_max", "n_bad_els"], ["f8", "f8", "f8", "f8",  "f8", "f8", "u4"]]
+        self.result_struct = [["value", "power", "temp", "power_time", "temp_min", "temp_max", "n_bad_els"], ["f8", "f8", "f8", "f8",  "f8", "f8", "f8"]]
 
         sample_dir = sample.directory
 

### 2019-07-25 random-frac/process.py
Update DFN healing

@@ -273,9 +273,10 @@ def prepare_mesh(config_dict, fractures):
     mesh_healed = mesh_name + "_healed.msh"
     if True: #not os.path.isfile(mesh_healed):
         import heal_mesh
-        hm = heal_mesh.HealMesh(mesh_file)
-        hm.heal_small_edges(0.001)
-        hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
+        hm = heal_mesh.HealMesh.read_mesh(mesh_file)
+        hm.heal_mesh(0.001)
+        #hm.heal_flat_elements(0.001)
+        #hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
         hm.write()
         assert hm.healed_mesh_name == mesh_healed
     return mesh_healed

### 2019-07-25 mlmc_random_frac/mlmc_process.py
Process temperature and bad elements results in mlmc samples.

@@ -54,10 +54,93 @@ class Process(base_process.Process):
             mlmc_est = Estimate(mlmc)
             mlmc_est_list.append(mlmc_est)
 
-        self.result_text(mlmc)
-        #self.plot_density(mlmc)
+        # self.result_text(mlmc)
+        # self.plot_density(mlmc)
 
-        self.plot_temp_power(mlmc_est)
+        # self.plot_temp_power(mlmc_est)
+
+        mlmc_est.mlmc.clean_select()
+        mlmc_est.mlmc.select_values({"n_bad_els": (-10, ">=")}, selected_param="n_bad_els")
+        n_bad_els = mlmc_est.mlmc.levels[0].sample_values
+
+        mlmc_est.mlmc.clean_select()
+        # mlmc_est.mlmc.select_values(None, selected_param="temp_min")
+        # mlmc_est.mlmc.select_values(None, selected_param="temp")
+        mlmc_est.mlmc.select_values({"temp_min": (100000, "<=")}, selected_param="temp_min")
+        temp_min = mlmc_est.mlmc.levels[0].sample_values
+
+        mlmc_est.mlmc.clean_select()
+        mlmc_est.mlmc.select_values({"temp_max": (-100000, ">=")}, selected_param="temp_max")
+        temp_max = mlmc_est.mlmc.levels[0].sample_values
+
+        n_samples = int(mlmc_est.mlmc.n_samples)
+        print("N samples: ", n_samples)
+        temp_v_ele = np.zeros((2 * n_samples, 2 * n_samples), dtype=int)
+        MIN_T = 250
+        MAX_T = 470
+        print("goot temperature interval: <{},{}>".format(MIN_T, MAX_T))
+        bad_temp_flag = np.zeros((len(n_bad_els),))
+        bad_ele_flag = np.zeros((len(n_bad_els),))
+        for i in range(0, n_samples):
+            bad_temp_flag[i] = min(temp_min[i][0]) < MIN_T or max(temp_max[i][0]) > MAX_T
+            bad_ele_flag[i] = n_bad_els[i][0][0] > 0
+            ty = i if bad_temp_flag[i] else n_samples + i
+            tx = i if bad_ele_flag[i] else n_samples + i
+            temp_v_ele[ty, tx] = 1
+
+        # print("temp_v_ele:\n", temp_v_ele)
+        temp_v_ele_sum = [
+            [np.sum(temp_v_ele[:n_samples, :n_samples]),
[…]

### 2019-07-25 random-frac/process.py
full rand_frac/process.py

@@ -271,12 +271,11 @@ def prepare_mesh(config_dict, fractures):
         make_mesh(config_dict, fractures, mesh_name, mesh_file)
 
     mesh_healed = mesh_name + "_healed.msh"
-    if True: #not os.path.isfile(mesh_healed):
+    if not os.path.isfile(mesh_healed):
         import heal_mesh
-        hm = heal_mesh.HealMesh.read_mesh(mesh_file)
-        hm.heal_mesh(0.001)
-        #hm.heal_flat_elements(0.001)
-        #hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
+        hm = heal_mesh.HealMesh.read_mesh(mesh_file, node_tol=1e-4)
+        hm.heal_mesh(tol_edge_ratio=0.01, tol_flat_ratio=0.01)
+        hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
         hm.write()
         assert hm.healed_mesh_name == mesh_healed
     return mesh_healed
@@ -472,14 +471,14 @@ def sample(tag, config_dict):
     config_dict["hm_params"]["mesh"] = healed_mesh
     config_dict["th_params"]["mesh"] = healed_mesh
 
-    # hm_succeed = call_flow(config_dict, 'hm_params', result_files=["mechanics.msh"])
-    # th_succeed = False
-    # if hm_succeed:
-    #     prepare_th_input(config_dict)
-    #     th_succeed = call_flow(config_dict, 'th_params', result_files=["energy_balance.yaml"])
-    #     if th_succeed:
-    #         series = extract_results(config_dict)
-    #         plot_exchanger_evolution(*series)
+    hm_succeed = call_flow(config_dict, 'hm_params', result_files=["mechanics.msh"])
+    th_succeed = False
+    if hm_succeed:
+        prepare_th_input(config_dict)
+        th_succeed = call_flow(config_dict, 'th_params', result_files=["energy_balance.yaml"])
+        if th_succeed:
+            series = extract_results(config_dict)
+            plot_exchanger_evolution(*series)
     os.chdir(root_dir)
 
 

### 2019-07-25 mlmc_random_frac/process.py
use updated mesh healing

@@ -269,10 +269,10 @@ def prepare_mesh(config_dict, fractures):
         make_mesh(config_dict, fractures, mesh_name, mesh_file)
 
     mesh_healed = mesh_name + "_healed.msh"
-    if True: #not os.path.isfile(mesh_healed):
+    if not os.path.isfile(mesh_healed):
         import heal_mesh
-        hm = heal_mesh.HealMesh(mesh_file)
-        hm.heal_small_edges(0.001)
+        hm = heal_mesh.HealMesh.read_mesh(mesh_file, node_tol=1e-4)
+        hm.heal_mesh(tol_edge_ratio=0.01, tol_flat_ratio=0.01)
         hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
         hm.write()
         assert hm.healed_mesh_name == mesh_healed

### 2019-07-25 mlmc_fixed_frac/frac_simulation.py
Updates from Overleaf

@@ -1,0 +1,363 @@
+import os
+import sys
+import shutil
+import yaml
+import numpy as np
+import time as t
+import copy
+sys.path.append('../MLMC/src')
+import gmsh_io as gmsh_io
+import mlmc.sample as sample
+from mlmc.simulation import Simulation
+
+
+def substitute_placeholders(file_in, file_out, params):
+    """
+    Substitute for placeholders of format '<name>' from the dict 'params'.
+    :param file_in: Template file.
+    :param file_out: Values substituted.
+    :param params: { 'name': value, ...}
+    """
+    used_params = []
+    with open(file_in, 'r') as src:
+        text = src.read()
+    for name, value in params.items():
+        placeholder = '<%s>' % name
+        n_repl = text.count(placeholder)
+        if n_repl > 0:
+            used_params.append(name)
+            text = text.replace(placeholder, str(value))
+    with open(file_out, 'w') as dst:
+        dst.write(text)
+
+    return used_params
+
+
+def force_mkdir(path, force=False):
+    """
+    Make directory 'path' with all parents,
+    remove the leaf dir recursively if it already exists.
+    :param path: path to directory
+    :param force: if dir already exists then remove it and create new one
+    :return: None
+    """
+    if force:
+        if os.path.isdir(path):
+            shutil.rmtree(path)
+    os.makedirs(path, mode=0o775, exist_ok=True)
+
+
+def load_config_dict():
+    with open("config.yaml", "r") as f:
+        return yaml.safe_load(f)
+
+
+class FracSimulation(Simulation):
+    HM_YAML_TEMPLATE = '01_hm_tmpl.yaml'
+    TH_YAML_TEMPLATE = '02_th_tmpl.yaml'
+
+    HM_YAML_FILE = "01_hm.yaml"
+    TH_YAML_FILE = "02_th.yaml"
+
+    def __init__(self, mesh_step, level_id=None, config=None, clean=False, parent_fine_sim=None):
+        """
+
+        :param config: configuration of the simulation, processed keys:
+            env - Environment object.
+            fields - FieldSet object
+            yaml_file: Template for main input file. Placeholders:
+                <mesh_file> - replaced by generated mesh
+                <FIELD> - for FIELD be name of any of `fields`, replaced by the FieldElementwise field with generated
+                 field input file and the field name for the component.
+                 (TODO: allow relative paths, not tested but should work)
+            geo_file: Path to the geometry file. (TODO: default is <yaml file base>.geo
[…]

### 2019-07-25 mlmc_fixed_frac/process.py

@@ -1,0 +1,229 @@
+import sys
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+src_path = os.path.dirname(os.path.abspath(__file__))
+sys.path.append('../MLMC/src')
+
+from mlmc import mlmc
+from mlmc import base_process
+from mlmc.estimate import Estimate, CompareLevels
+from mlmc.moments import Legendre, Monomial
+
+import pbs
+from frac_simulation import FracSimulation
+
+
+class Process(base_process.Process):
+    def run(self):
+        """
+        Run mlmc
+        :return: None
+        """
+        mc_samples = 2
+        os.makedirs(self.work_dir, mode=0o775, exist_ok=True)
+        self.n_moments = 10
+
+        mlmc_list = []
+        # Run one level Monte-Carlo method
+        for nl in [1]:  # , 2, 3, 4,5, 7, 9]:
+            mlmc = self.setup_config(nl, clean=True)
+            # self.n_sample_estimate(mlmc)
+            self.generate_jobs(mlmc, n_samples=[mc_samples])
+            mlmc_list.append(mlmc)
+
+        self.all_collect(mlmc_list)
+
+    def process(self):
+        """
+        Use collected data
+        :return: None
+        """
+        assert os.path.isdir(self.work_dir)
+        mlmc_est_list = []
+
+        for nl in [1]:  # high resolution fields
+            mlmc = self.setup_config(nl, clean=False)
+            # Use wrapper object for working with collected data
+            mlmc_est = Estimate(mlmc)
+            mlmc_est_list.append(mlmc_est)
+
+        #self.plot_density(mlmc)
+
+        self.plot_temp_power(mlmc_est)
+
+        # self.process_analysis(cl)
+
+    def set_environment_variables(self):
+        """
+        Set pbs config, flow123d, gmsh
+        :return: None
+        """
+        root_dir = os.path.abspath(self.work_dir)
+        while root_dir != '/':
+            root_dir, tail = os.path.split(root_dir)
+
+        self.pbs_config = dict(
+            job_weight=250000,  # max number of elements per job
+            n_cores=1,
+            n_nodes=1,
+            select_flags=['cgroups=cpuacct'],
+            mem='4gb',
+            queue='charon',
+            home_dir='/storage/liberec3-tul/home/martin_spetlik/')
+
+        if tail == 'storage':
+            # Metacentrum
+            self.sample_sleep = 30
+            self.init_sample_timeout = 600
+            self.sample_timeout = 0
[…]

### 2019-07-25 mlmc_random_frac/load_modules.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+echo "source this script: 'source load_modules.sh'"
+module -fs purge
+module load metabase
+module load cmake-3.6.1
+module load gcc-6.4.0
+module load boost-1.60-gcc
+module load mpich-3.0.2-gcc
+module rm python-2.7.6-gcc
+module load python-3.6.2-gcc
+module load python36-modules-gcc
+module list

### 2019-07-25 mlmc_random_frac/mlmc_process.py

@@ -1,0 +1,327 @@
+import sys
+import os
+import yaml
+import numpy as np
+import matplotlib.pyplot as plt
+src_path = os.path.dirname(os.path.abspath(__file__))
+sys.path.append(os.path.join(src_path, '../MLMC/src'))
+sys.path.append(os.path.join(src_path, '../dfn/src'))
+
+from mlmc import mlmc
+from mlmc import base_process
+from mlmc.estimate import Estimate, CompareLevels
+from mlmc.moments import Legendre, Monomial
+
+import pbs
+from rand_frac_simulation import RandomFracSimulation
+
+
+def load_config_dict():
+    with open("config.yaml", "r") as f:
+        return yaml.safe_load(f)
+
+
+class Process(base_process.Process):
+    def run(self):
+        """
+        Run mlmc
+        :return: None
+        """
+        os.makedirs(self.work_dir, mode=0o775, exist_ok=True)
+        self.n_moments = 10
+
+        mlmc_list = []
+        # Run one level Monte-Carlo method
+        for nl in [1]:  # , 2, 3, 4,5, 7, 9]:
+            mlmc = self.setup_config(nl, clean=True)
+            # self.n_sample_estimate(mlmc)
+            self.generate_jobs(mlmc, n_samples=[self.mc_samples])
+            mlmc_list.append(mlmc)
+
+        self.all_collect(mlmc_list)
+
+    def process(self):
+        """
+        Use collected data
+        :return: None
+        """
+        assert os.path.isdir(self.work_dir)
+        mlmc_est_list = []
+
+        for nl in [1]:  # high resolution fields
+            mlmc = self.setup_config(nl, clean=False)
+            # Use wrapper object for working with collected data
+            mlmc_est = Estimate(mlmc)
+            mlmc_est_list.append(mlmc_est)
+
+        # self.result_text(mlmc)
+        # self.plot_density(mlmc)
+
+        # self.plot_temp_power(mlmc_est)
+
+        mlmc_est.mlmc.clean_select()
+        mlmc_est.mlmc.select_values({"n_bad_els": (-10, ">=")}, selected_param="n_bad_els")
+        n_bad_els = mlmc_est.mlmc.levels[0].sample_values
+
+        mlmc_est.mlmc.clean_select()
+        # mlmc_est.mlmc.select_values(None, selected_param="temp_min")
+        # mlmc_est.mlmc.select_values(None, selected_param="temp")
[…]

### 2019-07-25 mlmc_random_frac/process.py

@@ -1,0 +1,492 @@
+import sys
+import os
+#WGC_DIR = "/storage/liberec3-tul/home/martin_spetlik/WGC_rand"
+#WORK_DIR = os.path.join(WGC_DIR, 'mlmc_random_frac')
+script_dir = os.path.dirname(os.path.realpath(__file__))
+sys.path.append(os.path.join(script_dir, '../MLMC/src'))
+sys.path.append(os.path.join(script_dir, '../dfn/src'))
+
+import shutil
+import subprocess
+import yaml
+import attr
+import numpy as np
+import collections
+import gmsh_io
+# import matplotlib.pyplot as plt
+
+import fracture
+
+from gmsh_api import gmsh
+from gmsh_api import options
+
+# TODO:
+# - enforce creation of empty physical groups, or creation of empty regions in the flow input
+# - speedup mechanics
+
+
+def substitute_placeholders(file_in, file_out, params):
+    """
+    Substitute for placeholders of format '<name>' from the dict 'params'.
+    :param file_in: Template file.
+    :param file_out: Values substituted.
+    :param params: { 'name': value, ...}
+    """
+    used_params = []
+    with open(file_in, 'r') as src:
+        text = src.read()
+    for name, value in params.items():
+        placeholder = '<%s>' % name
+        n_repl = text.count(placeholder)
+        if n_repl > 0:
+            used_params.append(name)
+            text = text.replace(placeholder, str(value))
+    with open(file_out, 'w') as dst:
+        dst.write(text)
+    return used_params
+
+
+def to_polar(x, y, z):
+    rho = np.sqrt(x ** 2 + y ** 2)
+    phi = np.arctan2(y, x)
+    if z > 0:
+        phi += np.pi
+    return (phi, rho)
+
+
+def plot_fr_orientation(fractures):
+    family_dict = collections.defaultdict(list)
+    for fr in fractures:
+        x, y, z = fracture.FisherOrientation.rotate(np.array([0,0,1]), axis=fr.rotation_axis, angle=fr.rotation_angle)[0]
+        family_dict[fr.region].append([
+            to_polar(z, y, x),
+            to_polar(z, x, -y),
+            to_polar(y, x, z)
+            ])
+
+    import matplotlib.pyplot as plt
+    fig, axes = plt.subplots(1, 3, subplot_kw=dict(projection='polar'))
+    for name, data in family_dict.items():
[…]

### 2019-07-25 mlmc_random_frac/rand_frac_simulation.py

@@ -1,0 +1,284 @@
+import os
+import sys
+import shutil
+import yaml
+import numpy as np
+import time as t
+
+sys.path.append('../MLMC/src')
+import mlmc.sample as sample
+from mlmc.simulation import Simulation
+
+
+def force_mkdir(path, force=False):
+    """
+    Make directory 'path' with all parents,
+    remove the leaf dir recursively if it already exists.
+    :param path: path to directory
+    :param force: if dir already exists then remove it and create new one
+    :return: None
+    """
+    if force:
+        if os.path.isdir(path):
+            shutil.rmtree(path)
+    os.makedirs(path, mode=0o775, exist_ok=True)
+
+
+def load_config_dict():
+    with open("config.yaml", "r") as f:
+        return yaml.safe_load(f)
+
+
+class RandomFracSimulation(Simulation):
+
+    def __init__(self, mesh_step, level_id=None, config=None, clean=False, parent_fine_sim=None):
+        """
+
+        :param config: configuration of the simulation, processed keys:
+            env - Environment object.
+            fields - FieldSet object
+            yaml_file: Template for main input file. Placeholders:
+                <mesh_file> - replaced by generated mesh
+                <FIELD> - for FIELD be name of any of `fields`, replaced by the FieldElementwise field with generated
+                 field input file and the field name for the component.
+                 (TODO: allow relative paths, not tested but should work)
+            geo_file: Path to the geometry file. (TODO: default is <yaml file base>.geo
+        :param mesh_step: Mesh step, decrease with increasing MC Level.
+        :param parent_fine_sim: Allow to set the fine simulation on previous level (Sim_f_l) which corresponds
+        to 'self' (Sim_c_l+1) as a coarse simulation. Usually Sim_f_l and Sim_c_l+1 are same simulations, but
+        these need to be different for advanced generation of samples (zero-mean control and antithetic).
+        """
+        if level_id is not None:
+            self.sim_id = level_id
+        else:
+            self.sim_id = RandomFracSimulation.total_sim_id
+            RandomFracSimulation.total_sim_id += 1
+
+        self.env = config['env']
+        self.step = mesh_step
+        # Pbs script creater
+        self.pbs_creater = self.env["pbs"]
+
+        # @TODO: set n fine elements
+        self.n_fine_elements = 0
+
+        # Prepare base workdir for this mesh_step
+        output_dir = config['output_dir']
+
[…]

### 2019-07-26 paper.tex

@@
 \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
+\usepackage[colorlinks]{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
@@
 \usepackage[bbgreekl]{mathbbol} % for blackboard \sigma
+\usepackage{color}
+\usepackage[square]{natbib}
 
@@
 \newcommand{\xx}{\vc x}
+\newcommand{\todo}[1]{{\color{red}#1}}
 
-\title{Stochastic modeling of EGS using continuum-fracture approach and multilevel Monte Carlo}
+\title{Stochastic modeling of EGS using continuum-fracture approach.}
 
@@
 \section{Introduction}
+\todo{
+- problem definition - uncertainty of EGS output due to  fracture network and other properties\\
+- goal - capturing these uncertainties by stochastic modelling\\
+- basic research (find some related work for dealing with modelling EGS with fracture opening and/or stochastic DFN\\
+- paper structure\\
+}
 
 \section{Mathematical model}
+According to available geological data of the Litoměřice locality, the rock in the depth of several km is considered crystallinic, possibly consisting of granite, gneiss or mica schist \cite{Capova2013}.
+% zula, rula, svor
+We also refer to the EGS study for this site in  contribution \cite{Ralek}.
 
-\subsection{Geological properties}
+\todo{short characterization of the locality, rock types,  temperature, reference to Ralek, identification of the model parameters postpone after model introduction.}
 
-According to available geological data of the Litoměřice locality, the rock in the depth of several km is considered crystallinic, possibly consisting of granite, gneiss or mica schist \cite{capova}.
-% zula, rula, svor
-The mechanical properties of granitic rocks have been investigated e.g. in \cite{ljunggren}, where the measured Young modulus $E$ varies around 50 GPa and Poisson's ratio $\nu$ is around 0.25.
-The reported values of compressibility $\beta$ of several types of granite and orthogneiss are approx. $20\times 10^{-13}$ cm${}^2$.dyne${}^{-1}$ = $2\times10^{-11}$ Pa${}^{-1}$ \cite{zisman}.
-As far as porosity $\vartheta$ is concerned, \cite{intera} report values between 0.001 and 0.01 for metamorphic crystalline rocks, we shall consider the value 0.005.
-For water we use the density $\varrho_l=10^3$ kg.m${}^{-3}$, compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$.
-Here and in what follows, subscript ``l'' indicates quantities associated to the liquid.
-Similarly, subscript ``f'' will denote quantities for the fractures.
-Storativity $S$ is calculated from compressibility and porosity as follows (\cite{brace-et-al}, see also \cite{intera}):
-\[ S = \varrho_l g(\beta + \vartheta\beta_l). \]
-Here $g=9.81$ m.s${}^{-2}$ is the standard gravity.
-In our case we obtain $S\approx 2\times 10^{-7} \mbox{ m}^{-1}$.
+\todo{Conceptual model:
 
-The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
-The hydraulic conductivity of fractures dominates than the one of intact rock by several orders of magnitude, we consider $k_f=10^{-3}$ m.s${}^{-1}$
-On the other hand, the elastic modulus of fracture is assumed much smaller, we take $E_f=50$ Pa.
-The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$.
-We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
-
-All values of physical parameters used in our model are listed in Table \ref{tab:params}.
-
-\begin{table}
-\centering
-\begin{tabular}{|l|c|c|c|}
-\hline
- & rock & fractures & liquid\\\hline
-Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
-Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{sperl-trckova} & $10^{-3}$ & ---\\
-Porosity $\vartheta$ [--] & $5\times10^{-3}$ & $5\times10^{-2}$ & --- \\
-Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ & $2\times10^{-11}$ & $5\times10^{-10}$ \\
-Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
-Cross-section $\delta$ [m] & --- & $10^{-3}$ & --- \\
-\hline
-Young modulus $E$ [Pa] & $50\times10^9$ & 50 & --- \\
-Poisson's ratio $\nu$ [--] & 0.25 & 0.25 & --- \\
-Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
-\hline
-Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
-Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
-\hline
-\end{tabular}
-\caption{Physical parameters of rock, fractures and water, used in the model.}
-\label{tab:params}
-\end{table}
-
-Papers about effective properties of fractured porous media: \cite{Snow1969}, \cite{Bonnet2001}, \cite{Olson2003}
-
-\subsection{Parameters of EGS site and model geometry}
-
-Number of wells: 2 \\
-Depth: 5 km \\
-Distance: 200 m \\
-Well radius: $\rho = 15$ cm \\
-Length of open part of wells: $L=60$ m \\
-
-% Pressure difference / water flux: 60 l/s
-
-In our model, we consider pair of vertical wells with the radius $\rho=0.15$ m in the distance $D=200$ m.
-The open part of the length $L=60$ m is in the depth $5000$ m.
-The cube $600 \times 600 \times 600$ m around the open parts is considered as the computational domain.
-
-{\bf The test geometry:} The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
-
-\subsection{Analytical models in the vicinity of the well}
-In order to keep the mesh size treatable, we cut out of the computational domain two cylinders around the open well parts with a larger radius $R=10$ m.
-On the surface of these cylinders we consider the Robin type boundary condition for the piezometric head:
-%
-\eq{\label{eq:H_robin} k\nabla h\cdot\nn = \sigma(H_w-h), }
-%
-where $\nn$ is the unit normal vector pointing to the centre of the well and $H_w$ is the piezometric head in the well.
-% We assume the constant conductivity $k$ in the vicinity of well.
-The parameter $\sigma$ is determined from the analytical radially symmetric solution
-of the steady Darcy flow problem with a point source:
-\[
-    \hat h(r) = -\alpha \log (r) + \beta,
-\]
-where $r$ is the distance from the well and $\alpha$, $\beta$ are arbitrary constants such that $\hat h(\rho)=H_w$:
-First, the flux from the well into the domain through the surface $C_\rho$ of the cylinder with radius $\rho$ and length $L$ is
-%  
-\eq{\label{eq:Q_rho}    Q_\rho := -\int_{C_\rho} k\nabla h\cdot\nn = -2\pi \rho L k \hat h'(\rho) = 2\pi L k \alpha. }
-%
-% where we have used the Darcy law with the conductivity $k$.
-%Now we use the analytical solution to fit the Robin type boundary condition:
-Next, the flux at the distance $R$ can be expressed from \eqref{eq:H_robin} as follows:
-%
-\eq{\label{eq:Q_R}   Q_R = 2\pi R L \sigma (\hat h(\rho) - \hat h(R)) = 2\pi R L \sigma \alpha \log(R/\rho).}
-%
-%which combined with previous equality determines $\sigma$ as:
-Finally, since $Q_\rho=Q_R$, we obtain from \eqref{eq:Q_rho} and \eqref{eq:Q_R}:
-\eq{ \label{eq:sigma} \sigma = \frac{k}{R\log(R/\rho)}. }
-In the simulations we shall use the hydraulic conductivity of the fractures in \eqref{eq:sigma} as the fractures have major influence on the flow in the vicinity of wells.
-
-Since the heat transfer is dominated by the convection in the vicinity of the wells, the temperature in the well is transported to the surface of the cut-off cylinders or vice verse in the time $t_0 = R/q$ which for the observed volume fluxes $Q\sim 1$ l/s is about $100$ days.
-This is negligible with respect to the total lifetime of the heat exchanger, hence in the simulation of heat we shall use Dirichlet boundary condition for the temperature.
+\begin{itemize}
+    \item modeling just small box around the heat exchanger
+    \item fracking considered as just opening existing fractures;simple HM model
+    \item deformation considered permanent
+    \item longterm heat transfer as separated problem;
+          TH model sequentialy coupled
+\end{itemize}
+}
 
@@
 \eq{\label{eq:biot} \partial_t(Sp + \alpha\div\uu) +\div\vc q = 0,\qquad -\div\bbsigma + \tilde\alpha\nabla p = 0. }
-Here $p=h-x_3$ is the pressure head [m], $\uu$ is the displacement [m] and $\tilde\alpha:=\alpha\varrho_l g$.
+Here $p=h-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\tilde\alpha:=\alpha\varrho_l g$ and $\alpha$ is the Biot's effective stress parameter \cite{Biot1941}.
 The flux $\vc q$ [m.s${}^{-1}$] and the stress tensor $\bbsigma$ [Pa] are given by the Darcy and Hooke laws:
@@
 
-For the 2d fractures we apply a dimension reduction procedure, which was originally developed in \cite{martin-jaffre-roberts} for the Darcy flow.
+For the 2d fractures we apply a dimension reduction procedure, which was originally developed in \cite{martin_modeling_2005} for the Darcy flow.
 We introduce the unit normal vector $\vc\nu$ to the fracture plane with arbitrary but fixed orientation.
@@
 
+\subsection{Conductivity dependent on displacement}
+We consider that after releasing the fracking pressure the displacement of larger fractures persists. So the change in new porosity is
+\[
+\phi = \phi_0 + \alpha\div \vc u
+\]
+where $\phi_0$ is the porosity before deformation, c.f. time term in \eqref{eq:biot}. 
+Using the Carman-Kozeny relation \cite{Kaviany1999}, 
+% \[
+%     k = \tilde{k}\frac{\phi^3}{(1-\phi)^2}
+% \]
+we can derive formula for the updated conductivity:
+\[
+    K = K_0\left(\frac{\phi}{\phi_0}\right)^3 \left(\frac{1-\phi_0}{1-\phi}\right)^2.
+\]
+
+\subsection{Analytical models in the vicinity of the well}
+In order to keep the mesh size treatable, we cut out of the computational domain two cylinders around the open well parts with a larger radius $R=10$ m.
+On the surface of these cylinders we consider the Robin type boundary condition for the piezometric head:
+%
+\eq{\label{eq:H_robin} k\nabla h\cdot\nn = \sigma(H_w-h), }
+%
+where $\nn$ is the unit normal vector pointing to the centre of the well and $H_w$ is the piezometric head in the well.
+% We assume the constant conductivity $k$ in the vicinity of well.
+The parameter $\sigma$ is determined from the analytical radially symmetric solution
+of the steady Darcy flow problem with a point source:
+\[
+    \hat h(r) = -\alpha \log (r) + \beta,
+\]
+where $r$ is the distance from the well and $\alpha$, $\beta$ are arbitrary constants such that $\hat h(\rho)=H_w$:
+First, the flux from the well into the domain through the surface $C_\rho$ of the cylinder with radius $\rho$ and length $L$ is
+%  
+\eq{\label{eq:Q_rho}    Q_\rho := -\int_{C_\rho} k\nabla h\cdot\nn = -2\pi \rho L k \hat h'(\rho) = 2\pi L k \alpha. }
+%
+% where we have used the Darcy law with the conductivity $k$.
+%Now we use the analytical solution to fit the Robin type boundary condition:
+Next, the flux at the distance $R$ can be expressed from \eqref{eq:H_robin} as follows:
+%
+\eq{\label{eq:Q_R}   Q_R = 2\pi R L \sigma (\hat h(\rho) - \hat h(R)) = 2\pi R L \sigma \alpha \log(R/\rho).}
+%
+%which combined with previous equality determines $\sigma$ as:
+Finally, since $Q_\rho=Q_R$, we obtain from \eqref{eq:Q_rho} and \eqref{eq:Q_R}:
+\eq{ \label{eq:sigma} \sigma = \frac{k}{R\log(R/\rho)}. }
+In the simulations we shall use the hydraulic conductivity of the fractures in \eqref{eq:sigma} as the fractures have major influence on the flow in the vicinity of wells.
+
+Since the heat transfer is dominated by the convection in the vicinity of the wells, the temperature in the well is transported to the surface of the cut-off cylinders or vice verse in the time $t_0 = R/q$ which for the observed volume fluxes $Q\sim 1$ l/s is about $100$ days.
+This is negligible with respect to the total lifetime of the heat exchanger, hence in the simulation of heat we shall use Dirichlet boundary condition for the temperature.
+
+\subsection{Geological properties}
+
+The mechanical properties of granitic rocks have been investigated e.g. in \cite{Ljunggren1985}, where the measured Young modulus $E$ varies around 50 GPa and Poisson's ratio $\nu$ is around 0.25.
+The reported values of the drained compressibility $\beta$ of several types of granite and orthogneiss are approx.
+$20\times 10^{-13}$ cm${}^2$.dyne${}^{-1}$ = $2\times10^{-11}$ Pa${}^{-1}$ \cite{Zisman1933}.
+As far as porosity $\vartheta$ is concerned, \cite{intera} report values between 0.001 and 0.01 for metamorphic crystalline rocks, we shall consider the value 0.005.
+For water we use the density $\varrho_l=10^3$ kg.m${}^{-3}$, compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$.
+Here and in what follows, subscript ``l'' indicates quantities associated to the liquid.
+Similarly, subscript ``f'' will denote quantities for the fractures.
+Storativity $S$ is calculated from compressibility and porosity as follows (\cite{Brace1968}, see also \cite{intera}):
+\[ S = \varrho_l g(\beta + \vartheta\beta_l). \]
+Here $g=9.81$ m.s${}^{-2}$ is the standard gravity.
+In our case we obtain $S\approx 2\times 10^{-7} \mbox{ m}^{-1}$.
+
+The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
+The hydraulic conductivity of fractures dominates than the one of intact rock by several orders of magnitude, we consider $k_f=10^{-3}$ m.s${}^{-1}$
+On the other hand, the elastic modulus of fracture is assumed much smaller, we take $E_f=50$ Pa.
+The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$.
+We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
+
+All values of physical parameters used in our model are listed in Table \ref{tab:params}.
+
+\begin{table}
+\centering
+\begin{tabular}{|l|c|c|c|}
+\hline
+ & rock & fractures & liquid\\\hline
+Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
+Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{Sperl2008} & $10^{-3}$ & ---\\
+Porosity $\vartheta$ [--] & $5\times10^{-3}$ & $5\times10^{-2}$ & --- \\
+Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ & $2\times10^{-11}$ & $5\times10^{-10}$ \\
+Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
+Cross-section $\delta$ [m] & --- & $10^{-3}$ & --- \\
+\hline
+Young modulus $E$ [Pa] & $50\times10^9$ & 50 & --- \\
+Poisson's ratio $\nu$ [--] & 0.25 & 0.25 & --- \\
+Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
+\hline
+Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
+Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
+\hline
+\end{tabular}
+\caption{Physical parameters of rock, fractures and water, used in the model.}
+\label{tab:params}
+\end{table}
+
+Papers about effective properties of fractured porous media: \cite{Snow1969}, \cite{Bonnet2001}, \cite{Olson2003}
+
+\subsection{Parameters of EGS site and model geometry}
+
+Number of wells: 2 \\
+Depth: 5 km \\
+Distance: 200 m \\
+Well radius: $\rho = 15$ cm \\
+Length of open part of wells: $L=60$ m \\
+
+% Pressure difference / water flux: 60 l/s
+
+In our model, we consider pair of vertical wells with the radius $\rho=0.15$ m in the distance $D=200$ m.
+The open part of the length $L=60$ m is in the depth $5000$ m.
+The cube $600 \times 600 \times 600$ m around the open parts is considered as the computational domain.
+
+{\bf The test geometry:} The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+
+\section{Fractured media}
+\subsection{Stochastic description of the fractures}
+\subsection{Effective properties of the fractured porous media}
+
 \section*{Acknowledgement}
@@
 
-\begin{thebibliography}{9}
+\bibliographystyle{dinat}
+\bibliography{references.bib}
 
-\bibitem{brace-et-al} W. F. Brace,  J. B. Walsh, W. T. Grango.  Permeability of Granite Under High Pressure, J. Geophys. Res. 73(6):2225--2236, 1968.
+% \begin{thebibliography}{9}
 
-\bibitem{capova} L. Čápová. Specification of the geothermic model in the environs of several selected boreholes. Diploma thesis, Charles University in Prague, 2013.
+% \bibitem{brace-et-al} W. F. Brace,  J. B. Walsh, W. T. Grango.  Permeability of Granite Under High Pressure, J. Geophys. Res. 73(6):2225--2236, 1968.
 
-\bibitem{intera} INTERA Environmental Consultants, Inc. Porosity, Permeability,  and Their  Relationship  in  
-Granite,  Basalt, and Tuff. Accession  DE83-011519, NTIS, Springfield, Virginia, 1983.
+% \bibitem{capova} L. Čápová. Specification of the geothermic model in the environs of several selected boreholes. Diploma thesis, Charles University in Prague, 2013.
 
-\bibitem{martin-jaffre-roberts} V.
+% \bibitem{intera} INTERA Environmental Consultants, Inc. Porosity, Permeability,  and Their  Relationship  in  
+% Granite,  Basalt, and Tuff. Accession  DE83-011519, NTIS, Springfield, Virginia, 1983.
+
+% \bibitem{martin-jaffre-roberts} V.
 Martin, J.
@@
 
-\bibitem{sperl-trckova} J. Šperl, J. Trčková. Permeability and porosity of rocks and their relationship based on laboratory testing. Acta Geodyn. Geomater. 5(149):41--47, 2008.
+% \bibitem{sperl-trckova} J. Šperl, J. Trčková. Permeability and porosity of rocks and their relationship based on laboratory testing. Acta Geodyn. Geomater. 5(149):41--47, 2008.
 
-\bibitem{ljunggren} C. Ljunggren, O. Stephansson, O. Alm, H. Hakami, U. Mattila. Mechanical properties of granitic rocks from Gide\aa, Sweden. Technical Report 85-06, SKB, 1985.
+% \bibitem{ljunggren} C. Ljunggren, O. Stephansson, O. Alm, H. Hakami, U. Mattila. Mechanical properties of granitic rocks from Gide\aa, Sweden. Technical Report 85-06, SKB, 1985.
 
-\bibitem{zisman} W.
+% \bibitem{zisman} W.
 A.
@@
 
-\end{thebibliography}
+% \end{thebibliography}
 

### 2019-07-26 mlmc_random_frac/process.py
Make charon options default in mlmc_random_frac dir

@@ -485,7 +485,7 @@ def sample(config_dict):
 if __name__ == "__main__":
     sample_dir = sys.argv[1]
     os.chdir(sample_dir)
-    np.random.seed(1)
+    np.random.seed()
 
     with open(os.path.join(script_dir, "config.yaml"), "r") as f:
         config_dict = yaml.safe_load(f)

### 2019-07-26 mlmc_random_frac/setup_python_environment.sh

@@ -3,6 +3,7 @@ echo "Make sure you called 'source load_modules.sh' before!"
 echo "Creating python environment.."
 python3.6 -m venv env
 
+source ./load_modules.sh
 source env/bin/activate
 python --version
 which python

### 2019-07-26 mlmc_random_frac/rand_frac_simulation.py
redirect stdou of process.py in PBS script

@@ -148,7 +148,7 @@ class RandomFracSimulation(Simulation):
         #self.pbs_script.append("pip3 install --user gmsh-sdk")
         self.pbs_script.append("cd {}".format(os.path.abspath(self.process_dir)))
         self.pbs_script.append("source {}/env/bin/activate".format(self.process_dir))
-        self.pbs_script.append("python {}/process.py {}".format(self.process_dir, sample_dir))
+        self.pbs_script.append("python {}/process.py {} >STDOUT 2>&1 ".format(self.process_dir, sample_dir))
         self.pbs_script.append("touch {}/FINISHED".format(sample_dir))
 
     def run_sim_sample(self, out_subdir):

### 2019-07-26 mlmc_random_frac/rand_frac_simulation.py
simplify pbs script creation and extraction

@@ -144,12 +144,16 @@ class RandomFracSimulation(Simulation):
         :param sample_dir: Sample dir abs path
         :return: None
         """
-        # All auxiliary methods must be run in pbs script
-        #self.pbs_script.append("pip3 install --user gmsh-sdk")
-        self.pbs_script.append("cd {}".format(os.path.abspath(self.process_dir)))
-        self.pbs_script.append("source {}/env/bin/activate".format(self.process_dir))
-        self.pbs_script.append("python {}/process.py {} >STDOUT 2>&1 ".format(self.process_dir, sample_dir))
-        self.pbs_script.append("touch {}/FINISHED".format(sample_dir))
+        self.pbs_script.append(
+            """
+            cd {proc_dir}
+            source {proc_dir}/env/bin/activate
+            python {proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
+            sleep 30
+            echo "done" >{sample_dir}/FINISHED
+            """
+            .format(proc_dir=self.process_dir, sample_dir=sample_dir))
+
 
     def run_sim_sample(self, out_subdir):
         """
@@ -207,12 +211,20 @@ class RandomFracSimulation(Simulation):
         water_balance_file = os.path.join(sample_dir, "output_02_th/water_balance.yaml")
         energy_balance_file = os.path.join(sample_dir, "output_02_th/energy_balance.yaml")
         heat_region_stat = os.path.join(sample_dir, "output_02_th/Heat_AdvectionDiffusion_region_stat.yaml")
+        finished_file = os.path.join(sample_dir, "FINISHED")
 
-        if os.path.exists(os.path.join(sample_dir, "FINISHED")):
-            if os.path.exists(heal_stat_file) and os.path.exists(water_balance_file) and os.path.exists(energy_balance_file) and os.path.exists(heat_region_stat):
+        finished = False
+        if os.path.exists(finished_file):
+            with open(finished_file, "r") as f:
+                content = f.read()
+            finished = content == "done"
+
+        if finished:
+            files_exist = all([os.path.exists(f) for f in [finished_file, heal_stat_file, water_balance_file, energy_balance_file, heat_region_stat]])
+            if files_exist:
                 # Sometimes content of files is not complete, sleep() seems to be workaround
-                if self.previous_length == 0:
-                    t.sleep(60)
+                # if self.previous_length == 0:
+                #     t.sleep(60)
 
                 while True:
[…]

### 2019-07-26 mlmc_random_frac/rand_frac_simulation.py
fix pbs script creation

@@ -146,13 +146,13 @@ class RandomFracSimulation(Simulation):
         """
         self.pbs_script.append(
             """
-            cd {proc_dir}
+            cd {abs_proc_dir}
             source {proc_dir}/env/bin/activate
             python {proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
             sleep 30
             echo "done" >{sample_dir}/FINISHED
             """
-            .format(proc_dir=self.process_dir, sample_dir=sample_dir))
+            .format(abs_proc_dir= os.path.abspath(self.process_dir), proc_dir=self.process_dir, sample_dir=sample_dir))
 
 
     def run_sim_sample(self, out_subdir):

### 2019-07-26 random-frac/process.py
Updated healing, imporved solvers, check conv.

@@ -274,12 +274,27 @@ def prepare_mesh(config_dict, fractures):
     if not os.path.isfile(mesh_healed):
         import heal_mesh
         hm = heal_mesh.HealMesh.read_mesh(mesh_file, node_tol=1e-4)
-        hm.heal_mesh(tol_edge_ratio=0.01, tol_flat_ratio=0.01)
+        hm.heal_mesh(tol_edge_ratio=0.02, tol_flat_ratio=0.02)
         hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
         hm.write()
         assert hm.healed_mesh_name == mesh_healed
     return mesh_healed
 
+def check_conv_reasons(log_fname):
+    with open(log_fname, "r") as f:
+        for line in f:
+            tokens = line.split(" ")
+            try:
+                i = tokens.index('convergence')
+                if tokens[i + 1] == 'reason':
+                    value = tokens[i + 2].rstrip(",")
+                    conv_reason = int(value)
+                    if conv_reason < 0:
+                        print("Failed to converge: ", conv_reason)
+                        return False
+            except ValueError:
+                continue
+    return True
 
 def call_flow(config_dict, param_key, result_files):
     """
@@ -302,9 +317,10 @@ def call_flow(config_dict, param_key, result_files):
         with open(fname + "_stdout", "w") as stdout:
             with open(fname + "_stderr", "w") as stderr:
                 completed = subprocess.run(arguments, stdout=stdout, stderr=stderr)
+        print("Exit status: ", completed.returncode)
         status = completed.returncode == 0
-    print("Exit status: ", status)
-    return status
+    conv_check = check_conv_reasons(os.path.join(output_dir, "flow123.0.log"))
+    return status # and conv_check
 
 
 
@@ -332,11 +348,13 @@ def prepare_th_input(config_dict):
     init_fr_cs = float(config_dict['hm_params']['fr_cross_section'])
     init_fr_K = float(config_dict['hm_params']['fr_conductivity'])
     init_bulk_K = float(config_dict['hm_params']['bulk_conductivity'])
+    min_fr_cross_section = float(config_dict['th_params']['min_fr_cross_section'])
 
     time_idx = 1
     time, field_cs = mesh.element_data['cross_section_updated'][time_idx]
 
-    cs = np.array([v[0] for v in field_cs.values()])
+
+    cs = np.maximum(np.array([v[0] for v in field_cs.values()]), min_fr_cross_section)
 
     K = np.where(
[…]

### 2019-07-26 mlmc_random_frac/process.py
Merge from rand-frac to mlmc

@@ -1,7 +1,5 @@
 import sys
 import os
-#WGC_DIR = "/storage/liberec3-tul/home/martin_spetlik/WGC_rand"
-#WORK_DIR = os.path.join(WGC_DIR, 'mlmc_random_frac')
 script_dir = os.path.dirname(os.path.realpath(__file__))
 sys.path.append(os.path.join(script_dir, '../MLMC/src'))
 sys.path.append(os.path.join(script_dir, '../dfn/src'))
@@ -17,12 +15,17 @@ import gmsh_io
 
 import fracture
 
-from gmsh_api import gmsh
-from gmsh_api import options
-
 # TODO:
 # - enforce creation of empty physical groups, or creation of empty regions in the flow input
 # - speedup mechanics
+
+@attr.s(auto_attribs=True)
+class ValueDescription:
+    time: float
+    position: str
+    quantity: str
+    unit: str
+
 
 
 def substitute_placeholders(file_in, file_out, params):
@@ -45,14 +48,12 @@ def substitute_placeholders(file_in, file_out, params):
         dst.write(text)
     return used_params
 
-
 def to_polar(x, y, z):
     rho = np.sqrt(x ** 2 + y ** 2)
     phi = np.arctan2(y, x)
     if z > 0:
         phi += np.pi
     return (phi, rho)
-
 
 def plot_fr_orientation(fractures):
     family_dict = collections.defaultdict(list)
@@ -84,7 +85,6 @@ def plot_fr_orientation(fractures):
     fig.savefig("fracture_orientation.pdf")
     plt.close(fig)
     #plt.show()
-
 
 def generate_fractures(config_dict):
     geom = config_dict["geometry"]
@@ -256,9 +256,10 @@ def make_mesh(config_dict, fractures, mesh_name, mesh_file):
     mesh.MinimumCirclePoints = 6
     mesh.MinimumCurvePoints = 2
 
+
     #factory.make_mesh(mesh_groups, dim=2)
     factory.make_mesh(mesh_groups)
-    factory.write_mesh(filename=mesh_name + "." + gmsh.MeshFormat.msh2.name, format=gmsh.MeshFormat.msh2)
+    factory.write_mesh(format=gmsh.MeshFormat.msh2)
     os.rename(mesh_name + ".msh2", mesh_file)
 
 
@@ -272,13 +273,27 @@ def prepare_mesh(config_dict, fractures):
[…]

### 2019-07-26 mlmc_random_frac/load_modules.sh
Fix to run on local

@@ -1,12 +1,16 @@
 #!/bin/bash
 echo "source this script: 'source load_modules.sh'"
-module -fs purge
-module load metabase
-module load cmake-3.6.1
-module load gcc-6.4.0
-module load boost-1.60-gcc
-module load mpich-3.0.2-gcc
-module rm python-2.7.6-gcc
-module load python-3.6.2-gcc
-module load python36-modules-gcc
-module list
+
+if which module
+then
+    module -fs purge
+    module load metabase
+    module load cmake-3.6.1
+    module load gcc-6.4.0
+    module load boost-1.60-gcc
+    module load mpich-3.0.2-gcc
+    module rm python-2.7.6-gcc
+    module load python-3.6.2-gcc
+    module load python36-modules-gcc
+    module list
+fi

### 2019-07-26 mlmc_random_frac/mlmc_process.py

@@ -154,19 +154,28 @@ class Process(base_process.Process):
             root_dir, tail = os.path.split(root_dir)
 
         self.flow123d = self.config_dict["flow_executable"]
+        self.pbs_config = dict(
+            job_weight=250000,  # max number of elements per job
+            n_cores=1,
+            n_nodes=1,
+            select_flags=[],
+            mem='8gb',
+            queue='charon',
+            qsub=None)
         if(self.config_dict["metacentrum"]):
-            self.pbs_config = dict(
-                job_weight=250000,  # max number of elements per job
-                n_cores=3,
-                n_nodes=1,
-                select_flags=['cgroups=cpuacct'],
-                mem='8gb',
-                queue='charon')
             self.sample_sleep = 30
             self.init_sample_timeout = 600
             self.sample_timeout = 0
-            self.pbs_config['qsub'] = '/usr/bin/qsub'
+            self.pbs_config.update(dict(
+                qsub='/usr/bin/qsub',
+                modules=[
+
+                ]
+            ))
         else:
+            # pbs_config is necessary for the local run but is not used
+            # as long as the pbs key is set to None
+            self.pbs_config['qsub'] = None
             self.sample_sleep = 1
             self.init_sample_timeout = 60
             self.sample_timeout = 60
@@ -189,7 +198,7 @@ class Process(base_process.Process):
                                job_count=num_jobs,
                                qsub=self.pbs_config['qsub'],
                                clean=clean)
-        self.pbs_obj.pbs_common_setting(flow_3=True, **self.pbs_config)
+        self.pbs_obj.pbs_common_setting(**self.pbs_config)
 
     def setup_config(self, n_levels, clean):
         """
@@ -207,10 +216,11 @@ class Process(base_process.Process):
             self.rm_files(output_dir)
 
         # Init pbs object
-        if (self.config_dict["metacentrum"]):
-            self.create_pbs_object(output_dir, clean)
-        else:
-            self.pbs_obj = None
+        self.create_pbs_object(output_dir, clean)
+        # if (self.config_dict["metacentrum"]):
+        #
+        # else:
+        #     self.pbs_obj = None
 
         simulation_config = {
             'env': dict(flow123d=self.flow123d, pbs=self.pbs_obj),  # The Environment.

### 2019-07-26 mlmc_random_frac/setup_python_environment.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 echo "Make sure you called 'source load_modules.sh' before!"
 echo "Creating python environment.."
-python3.6 -m venv env
+python3 -m venv env
 
 source ./load_modules.sh
 source env/bin/activate
@@ -11,7 +11,7 @@ which pip
 pip install --upgrade pip
 pip -V
 
-pip install gmsh-sdk
+pip install pyyaml attrs numpy gmsh-sdk
 
 #pip freeze
 deactivate

### 2019-07-27 mlmc_random_frac/process.py
update dfn, full heat period

@@ -273,7 +273,7 @@ def prepare_mesh(config_dict, fractures):
     if not os.path.isfile(mesh_healed):
         import heal_mesh
         hm = heal_mesh.HealMesh.read_mesh(mesh_file, node_tol=1e-4)
-        hm.heal_mesh(tol_edge_ratio=0.02, tol_flat_ratio=0.02)
+        hm.heal_mesh(gamma_tol=0.01)
         hm.stats_to_yaml(mesh_name + "_heal_stats.yaml")
         hm.write()
         assert hm.healed_mesh_name == mesh_healed
@@ -382,83 +382,83 @@ def prepare_th_input(config_dict):
 
 
 
-def get_result_description():
-    """
-    :return:
-    """
-    end_time = 30
-    values = [ [ValueDescription(time=t, position="extraction_well", quantity="power", unit="MW"),
-                ValueDescription(time=t, position="extraction_well", quantity="temperature", unit="Celsius deg.")
-                ] for t in np.linspace(0, end_time, 0.1)]
-    power_series, temp_series = zip(*values)
-    return power_series + temp_series
-
-
-def extract_time_series(yaml_stream, regions, extract):
-    """
-
-    :param yaml_stream:
-    :param regions:
-    :return: times list, list: for every region the array of value series
-    """
-    data = yaml.safe_load(yaml_stream)['data']
-    times = set()
-    reg_series = {reg: [] for reg in regions}
-
-    for time_data in data:
-        region = time_data['region']
-        if region in reg_series:
-            times.add(time_data['time'])
-            power_in_time = extract(time_data)
-            reg_series[region].append(power_in_time)
-    times = list(times)
-    times.sort()
-    series = [np.array(region_series, dtype=float) for region_series in reg_series.values()]
-    return np.array(times), series
-
-
-def extract_results(config_dict):
-    """
-    :param config_dict: Parsed config.yaml. see key comments there.
-    : return
-    """
-    bc_regions = ['.fr_left_well', '.left_well', '.fr_right_well', '.right_well']
-    out_regions = bc_regions[2:]
-    output_dir = config_dict["th_params"]["output_dir"]
-    with open(os.path.join(output_dir, "energy_balance.yaml"), "r") as f:
-        power_times, reg_powers = extract_time_series(f, bc_regions, extract=lambda frame: frame['data'][0])
[…]

### 2019-07-27 mlmc_random_frac/extract_meshes.py
Sampling from a mesh repository.

@@ -1,0 +1,41 @@
+import os
+from shutil import copyfile
+
+samples_dir = os.path.join("output_1", "sim_0_step_0.010000", "samples")
+mesh_repository_dir = os.path.join("mesh_repository")
+meshes = {}
+mesh_indices = []
+for mesh_file in os.listdir(os.fsencode(mesh_repository_dir)):
+    mesh_path = os.path.join(mesh_repository_dir, os.fsdecode(mesh_file))
+    info = os.stat(mesh_path)
+    meshes[info.st_size] = mesh_path
+    tags = os.fsdecode(mesh_file).split('_')
+    mesh_indices.append(int(tags[0]))
+mesh_indices.sort()
+if mesh_indices:
+  last_index = mesh_indices[-1]
+else:
+  last_index = 0
+
+print(samples_dir)
+for sdir in os.listdir(os.fsencode(samples_dir)):  
+    #print(sdir)
+    sdir = os.path.join(samples_dir,os.fsdecode(sdir))
+    if os.path.isdir(sdir):
+        finished_path = os.path.join(sdir, "FINISHED")
+        #print(finished_path)
+        if os.path.isfile(finished_path):
+            #print(sdir, " FINISHED")
+            mesh_file = "random_fractures_healed.msh"
+            mesh_path = os.path.join(sdir, mesh_file )
+            if os.path.isfile(mesh_path):             
+                info = os.stat(mesh_path)
+                if info.st_size not in meshes:
+                    last_index += 1
+                    copy_mesh_file = "{:04d}_{}".format(last_index, mesh_file)
+                    copyfile(mesh_path, os.path.join(mesh_repository_dir, copy_mesh_file))
+                    print("Copy {} to {}.".format(mesh_path, copy_mesh_file))
+                else:
+                    print("Skipping mesh {} same size as {}.".format(mesh_path, meshes[info.st_size]))
+
+    

### 2019-07-27 mlmc_random_frac/rand_frac_simulation.py

@@ -216,12 +216,15 @@ class RandomFracSimulation(Simulation):
         finished = False
         if os.path.exists(finished_file):
             with open(finished_file, "r") as f:
-                content = f.read()
-            finished = content == "done"
+                content = f.read().split()                
+            finished = len(content)==1 and content[0] == "done"
 
         if finished:
-            files_exist = all([os.path.exists(f) for f in [finished_file, heal_stat_file, water_balance_file, energy_balance_file, heat_region_stat]])
+            print(sample_dir, "Finished")
+            finished_map = {f:os.path.exists(f) for f in [finished_file, heal_stat_file, water_balance_file, energy_balance_file, heat_region_stat]}
+            files_exist = all(finished_map.values())
             if files_exist:
+                print(sample_dir, "Files exist")
                 # Sometimes content of files is not complete, sleep() seems to be workaround
                 # if self.previous_length == 0:
                 #     t.sleep(60)
@@ -288,9 +291,11 @@ class RandomFracSimulation(Simulation):
                 for i in range(len(power_times)):
                     result_values.append((i, power_series[i], avg_temp[i], power_times[i],
                                           temp_min[i], temp_max[i], n_bad_els))
-
+                
+                
                 return result_values
             else:
+                print(sample_dir, "missing files", finished_map)
                 return [np.inf, np.inf, np.inf, np.inf, np.inf, np.inf, np.inf]
 
         else:

### 2019-07-27 mlmc_random_frac/process.py
sampling from mesh repo, process.py

@@ -319,6 +319,7 @@ def call_flow(config_dict, param_key, result_files):
         print("Exit status: ", completed.returncode)
         status = completed.returncode == 0
     conv_check = check_conv_reasons(os.path.join(output_dir, "flow123.0.log"))
+    print("converged: ", conv_check)
     return status # and conv_check
 
 
@@ -460,6 +461,20 @@ def prepare_th_input(config_dict):
 #     fig.tight_layout()  # otherwise the right y-label is slightly clipped
 #     plt.show()
 
+def sample_mesh_repository(mesh_repository):
+    """
+    
+    """
+    mesh_repository_dir = os.path.join(script_dir, "mesh_repository")
+    mesh_file = np.random.choice(os.listdir(mesh_repository_dir))
+    healed_mesh = "random_fractures_healed.msh"
+    shutil.copyfile(os.path.join(mesh_repository_dir, mesh_file), healed_mesh)
+    heal_ref_report = { 'flow_stats': { 'bad_el_tol': 0.01, 'bad_elements': [], 'bins': [], 'hist': []},
+                        'gamma_stats': {'bad_el_tol': 0.01, 'bad_elements': [], 'bins': [], 'hist': []}}
+    with open("random_fractures_heal_stats.yaml", "w") as f:
+        yaml.dump(heal_ref_report, f)
+    return healed_mesh
+
 
 def setup_dir(config_dict, clean=False):
     for f in config_dict["copy_files"]:
@@ -471,11 +486,14 @@ def setup_dir(config_dict, clean=False):
 def sample(config_dict):
 
     setup_dir(config_dict, clean=True)
-    #setup_dir(tag, config_dict)
+    mesh_repo = config_dict.get('mesh_repository', None)
+    if mesh_repo:
+        healed_mesh = sample_mesh_repository(mesh_repo) 
+    else:
+        fractures = generate_fractures(config_dict)
+        # plot_fr_orientation(fractures)
+        healed_mesh = prepare_mesh(config_dict, fractures)
 
-    fractures = generate_fractures(config_dict)
-    # plot_fr_orientation(fractures)
-    healed_mesh = prepare_mesh(config_dict, fractures)
     healed_mesh_bn = os.path.basename(healed_mesh)
     config_dict["hm_params"]["mesh"] = healed_mesh_bn
     config_dict["th_params"]["mesh"] = healed_mesh_bn
@@ -488,7 +506,7 @@ def sample(config_dict):
         #if th_succeed:
[…]

### 2019-07-27 mlmc_random_frac/load_modules.sh
Fix some errors to run the simulation on charon.

@@ -1,8 +1,8 @@
 #!/bin/bash
 echo "source this script: 'source load_modules.sh'"
 
-if which module
-then
+# if which module
+# then
     module -fs purge
     module load metabase
     module load cmake-3.6.1
@@ -13,4 +13,4 @@ then
     module load python-3.6.2-gcc
     module load python36-modules-gcc
     module list
-fi
+# fi

### 2019-07-27 mlmc_random_frac/process.py

@@ -487,10 +487,10 @@ def sample(config_dict):
 
     setup_dir(config_dict, clean=True)
     mesh_repo = config_dict.get('mesh_repository', None)
+    fractures = generate_fractures(config_dict)
     if mesh_repo:
-        healed_mesh = sample_mesh_repository(mesh_repo) 
+        healed_mesh = sample_mesh_repository(mesh_repo)
     else:
-        fractures = generate_fractures(config_dict)
         # plot_fr_orientation(fractures)
         healed_mesh = prepare_mesh(config_dict, fractures)
 

### 2019-07-27 mlmc_random_frac/rand_frac_simulation.py

@@ -147,6 +147,7 @@ class RandomFracSimulation(Simulation):
         self.pbs_script.append(
             """
             cd {abs_proc_dir}
+            module load python36-modules-gcc
             source {proc_dir}/env/bin/activate
             python {proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
             sleep 30

### 2019-07-27 mlmc_random_frac/load_modules.sh
Fine fracture meshing

@@ -1,8 +1,9 @@
 #!/bin/bash
 echo "source this script: 'source load_modules.sh'"
 
-# if which module
-# then
+x=`declare -F | grep "declare -fx module"`
+if [ -n "$x" ]
+then
     module -fs purge
     module load metabase
     module load cmake-3.6.1
@@ -13,4 +14,4 @@ echo "source this script: 'source load_modules.sh'"
     module load python-3.6.2-gcc
     module load python36-modules-gcc
     module list
-# fi
+fi

### 2019-07-27 mlmc_random_frac/process.py

@@ -153,7 +153,7 @@ def create_fractures_polygons(gmsh_geom, fractures):
 
 
 def make_mesh(config_dict, fractures, mesh_name, mesh_file):
-    fracture_mesh_step = 10
+    fracture_mesh_step = 15
     geom = config_dict["geometry"]
     dimensions = geom["box_dimensions"]
     well_z0, well_z1 = geom["well_openning"]
@@ -163,6 +163,8 @@ def make_mesh(config_dict, fractures, mesh_name, mesh_file):
 
     from gmsh_api import gmsh
     from gmsh_api import options
+    from gmsh_api import field
+
     factory = gmsh.GeometryOCC(mesh_name, verbose=True)
     gopt = options.Geometry()
     gopt.Tolerance = 0.0001
@@ -225,7 +227,9 @@ def make_mesh(config_dict, fractures, mesh_name, mesh_file):
     b_fr_right_well = b_fractures.select_by_intersect(b_right_well).modify_regions("{}_right_well")
     b_fractures = factory.group(b_fr_left_well, b_fr_right_well, b_fractures_box)
     mesh_groups = [*box_all, fractures_fr, b_fractures]
-    fractures_fr.set_mesh_step(fracture_mesh_step)
+
+    print(fracture_mesh_step)
+    #fractures_fr.set_mesh_step(fracture_mesh_step)
 
     factory.keep_only(*mesh_groups)
     factory.remove_duplicate_entities()
@@ -236,21 +240,21 @@ def make_mesh(config_dict, fractures, mesh_name, mesh_file):
     max_el_size = np.max(dimensions) / 8
 
 
-    # fracture_el_size = field.constant(100, 10000)
-    # frac_el_size_only = field.restrict(fracture_el_size, fractures_fr, add_boundary=True)
-    # field.set_mesh_step_field(frac_el_size_only)
+    fracture_el_size = field.constant(fracture_mesh_step, 10000)
+    frac_el_size_only = field.restrict(fracture_el_size, fractures_fr, add_boundary=True)
+    field.set_mesh_step_field(frac_el_size_only)
 
     mesh = options.Mesh()
-    # mesh.Algorithm = options.Algorithm2d.MeshAdapt # produce some degenerated 2d elements on fracture boundaries ??
[…]

### 2019-07-27 mlmc_random_frac/rand_frac_simulation.py

@@ -147,7 +147,7 @@ class RandomFracSimulation(Simulation):
         self.pbs_script.append(
             """
             cd {abs_proc_dir}
-            module load python36-modules-gcc
+            source {proc_dir}/load_modules.sh
             source {proc_dir}/env/bin/activate
             python {proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
             sleep 30

### 2019-07-28 mlmc_random_frac/rand_frac_simulation.py
fix for run in subdirs

@@ -5,7 +5,10 @@ import yaml
 import numpy as np
 import time as t
 
-sys.path.append('../MLMC/src')
+src_path = os.path.dirname(os.path.abspath(__file__))
+sys.path.append(os.path.join(src_path, '../MLMC/src'))
+sys.path.append(os.path.join(src_path, '../dfn/src'))
+
 import mlmc.sample as sample
 from mlmc.simulation import Simulation
 
@@ -65,7 +68,7 @@ class RandomFracSimulation(Simulation):
         # Prepare base workdir for this mesh_step
         output_dir = config['output_dir']
 
-        self.process_dir = os.path.split(output_dir)[0]
+        self.process_dir = src_path
         self.work_dir = os.path.join(output_dir, 'sim_%d_step_%f' % (self.sim_id, self.step))
 
         force_mkdir(self.work_dir, clean)
@@ -147,13 +150,13 @@ class RandomFracSimulation(Simulation):
         self.pbs_script.append(
             """
             cd {abs_proc_dir}
-            source {proc_dir}/load_modules.sh
-            source {proc_dir}/env/bin/activate
-            python {proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
+            source {abs_proc_dir}/load_modules.sh
+            source {abs_proc_dir}/env/bin/activate
+            python {abs_proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
             sleep 30
             echo "done" >{sample_dir}/FINISHED
             """
-            .format(abs_proc_dir= os.path.abspath(self.process_dir), proc_dir=self.process_dir, sample_dir=sample_dir))
+            .format(abs_proc_dir= self.process_dir, sample_dir=sample_dir))
 
 
     def run_sim_sample(self, out_subdir):

### 2019-07-28 mlmc_random_frac/rand_frac_simulation.py
Support for sample reuse

@@ -126,13 +126,20 @@ class RandomFracSimulation(Simulation):
         # Get sample directory
         out_subdir = os.path.join("samples", str(sample_tag))
         sample_dir = os.path.join(self.work_dir, out_subdir)
-        force_mkdir(sample_dir, True)
+        reuse_samples = self.config_dict.get('reuse_samples', None)
+        if reuse_samples is None:
+            force_mkdir(sample_dir, True)
+        else:
+            force_mkdir(sample_dir, False)
 
         # Run part which can be run via pbs
         self.create_pbs_script(sample_dir)
 
         # Pbs package directory
         package_dir = self.run_sim_sample(out_subdir)
+        # if reuse_samples:
+        #     self.pbs_creater._number_of_realizations = 0
+        # else:
         self.pbs_creater.execute()
 
         return sample.Sample(directory=sample_dir, sample_id=sample_id,
@@ -144,16 +151,17 @@ class RandomFracSimulation(Simulation):
         :param sample_dir: Sample dir abs path
         :return: None
         """
+        finish_sleep = self.config_dict.get("finish_sleep", 30)
         self.pbs_script.append(
             """
             cd {abs_proc_dir}
             source {proc_dir}/load_modules.sh
             source {proc_dir}/env/bin/activate
             python {proc_dir}/process.py {sample_dir} >{sample_dir}/STDOUT 2>&1
-            sleep 30
+            sleep {finish_sleep}
             echo "done" >{sample_dir}/FINISHED
             """
-            .format(abs_proc_dir= os.path.abspath(self.process_dir), proc_dir=self.process_dir, sample_dir=sample_dir))
+            .format(abs_proc_dir= os.path.abspath(self.process_dir), proc_dir=self.process_dir, sample_dir=sample_dir, finish_sleep=finish_sleep))
 
 
     def run_sim_sample(self, out_subdir):

### 2019-07-28 mlmc_random_frac/process.py
Add reference TH model with constant conductivity and cross section.
- add third model
- extract results
- update config

@@ -108,7 +108,7 @@ def generate_fractures(config_dict):
     for fr in fractures:
         fr.region = "fr"
     used_families = set((f.region for f in fractures))
-    for model in ["hm_params", "th_params"]:
+    for model in ["hm_params", "th_params", "th_params_ref"]:
         model_dict = config_dict[model]
         model_dict["fracture_regions"] = list(used_families)
         model_dict["left_well_fracture_regions"] = [".{}_left_well".format(f) for f in used_families]
@@ -497,12 +497,15 @@ def sample(config_dict):
     healed_mesh_bn = os.path.basename(healed_mesh)
     config_dict["hm_params"]["mesh"] = healed_mesh_bn
     config_dict["th_params"]["mesh"] = healed_mesh_bn
+    config_dict["th_params_ref"]["mesh"] = healed_mesh_bn
 
     hm_succeed = call_flow(config_dict, 'hm_params', result_files=["mechanics.msh"])
+    th_succeed = call_flow(config_dict, 'th_params_ref', result_files=["energy_balance.yaml"])
     th_succeed = False
     if hm_succeed:
         prepare_th_input(config_dict)
         th_succeed = call_flow(config_dict, 'th_params', result_files=["energy_balance.yaml"])
+        
         #if th_succeed:
         #    series = extract_results(config_dict)
         #    plot_exchanger_evolution(*series)

### 2019-07-28 mlmc_random_frac/rand_frac_simulation.py

@@ -204,14 +204,17 @@ class RandomFracSimulation(Simulation):
         abs_zero_temp = 273.15
         year_sec = 60 * 60 * 24 * 365
         # Sample result structure -> you can enlarge it and other stuff will be done automatically
-        self.result_struct = [["value", "power", "temp", "power_time", "temp_min", "temp_max", "n_bad_els"], ["f8", "f8", "f8", "f8",  "f8", "f8", "f8"]]
+        self.result_struct = [["value", "power_time",
+                               "power", "temp", "fluxes", "temp_min", "temp_max",
+                               "power_ref", "temp_ref", "fluxes_ref", "temp_min_ref", "temp_max_ref",
+                               "n_bad_els"],
+                              ["f8", "f8",
+                               "f8", "f8", "f8", "f8",  "f8",
+                               "f8", "f8", "f8", "f8", "f8",
+                               "f8"]]
 
         sample_dir = sample.directory
-
         heal_stat_file = os.path.join(sample_dir, "random_fractures_heal_stats.yaml")
-        water_balance_file = os.path.join(sample_dir, "output_02_th/water_balance.yaml")
-        energy_balance_file = os.path.join(sample_dir, "output_02_th/energy_balance.yaml")
-        heat_region_stat = os.path.join(sample_dir, "output_02_th/Heat_AdvectionDiffusion_region_stat.yaml")
         finished_file = os.path.join(sample_dir, "FINISHED")
 
         finished = False
@@ -222,7 +225,7 @@ class RandomFracSimulation(Simulation):
 
         if finished:
             print(sample_dir, "Finished")
-            finished_map = {f:os.path.exists(f) for f in [finished_file, heal_stat_file, water_balance_file, energy_balance_file, heat_region_stat]}
+            finished_map = {f:os.path.exists(f) for f in [finished_file, heal_stat_file]}
             files_exist = all(finished_map.values())
             if files_exist:
                 print(sample_dir, "Files exist")
@@ -230,69 +233,19 @@ class RandomFracSimulation(Simulation):
                 # if self.previous_length == 0:
                 #     t.sleep(60)
 
-                while True:
-                    # extract the flux
-                    bc_regions = ['.fr_left_well', '.left_well', '.fr_right_well', '.right_well']
-                    regions = ['fr', 'box']
-                    out_regions = bc_regions[2:]
+                with open(heal_stat_file, "r") as f:
[…]

### 2019-07-28 mlmc_random_frac/rand_frac_simulation.py
Try to fix extract results when failed.

@@ -251,6 +251,9 @@ class RandomFracSimulation(Simulation):
                 power_times, th_result_values = self._extract_result_th(os.path.join(sample_dir, "output_02_th"))
                 power_times, th_ref_result_values = self._extract_result_th(os.path.join(sample_dir, "output_03_th"))
 
+                if power_times[0] == np.inf:
+                    return [np.inf, np.inf, np.inf, np.inf, np.inf, np.inf, np.inf]
+
                 if self.previous_length == 0:
                     self.previous_length = len(power_times)
 
@@ -325,4 +328,4 @@ class RandomFracSimulation(Simulation):
             return power_times, th_result_values
         else:
             print(th_outdir, "missing files", finished_map)
-            return None, [np.inf, np.inf, np.inf, np.inf, np.inf]
+            return [np.inf], [np.inf, np.inf, np.inf, np.inf, np.inf]

### 2019-07-28 mlmc_random_frac/mlmc_process.py
select samples by temperature mask

@@ -41,6 +41,36 @@ class Process(base_process.Process):
 
         self.all_collect(mlmc_list)
 
+
+    def get_samples(self, mlmc_est, quantity):
+        mlmc_est.mlmc.select_values(None, selected_param=quantity)
+        q_array = mlmc_est.mlmc.levels[0].sample_values
+        mlmc_est.mlmc.clean_select()
+        return q_array[:,0,:]
+
+
+    def select_samples_with_good_values(self, mlmc_est):
+        # determine samples with correct temperature
+        n_bad_els = self.get_samples(mlmc_est, 'n_bad_els')
+        temp_min = self.get_samples(mlmc_est, 'temp_min')
+        temp_max = self.get_samples(mlmc_est, 'temp_max')
+        temp = self.get_samples(mlmc_est, 'temp')
+
+        abs_zero_temp = 273.15
+        MIN_T = 250
+        MAX_T = 470
+        temp_good_min = np.min(temp_min[:, :], axis=1) > MIN_T
+        temp_good_max = np.max(temp_max[:, :], axis=1) < MAX_T
+        temp_good = np.logical_and(np.min(temp, axis=1) > MIN_T - 273.15, np.max(temp, axis=1) < MAX_T - 273.15)
+        no_bad_els = n_bad_els[:, 0] == 0
+        good_mask = np.logical_and(temp_good_min, temp_good_max)
+        good_mask = np.logical_and(good_mask, no_bad_els)
+        good_mask = np.logical_and(good_mask, temp_good)
+        good_indices = np.arange(0, len(good_mask))[good_mask]
+        bad_indices = np.arange(0, len(good_mask))[~good_mask]
+        mlmc_est.mlmc.subsample_by_indices(good_indices)
+
+
     def process(self):
         """
         Use collected data
@@ -57,29 +87,16 @@ class Process(base_process.Process):
 
         # self.result_text(mlmc)
         # self.plot_density(mlmc)
+        print("N all samples: ", mlmc_est.mlmc.n_samples[0])
+        self.select_samples_with_good_values(mlmc_est)
+        self.plot_temp_power(mlmc_est)
+        print("N good samples: ", mlmc_est.mlmc.n_samples[0])
 
-        # self.plot_temp_power(mlmc_est)
-
-        mlmc_est.mlmc.clean_select()
-        mlmc_est.mlmc.select_values({"n_bad_els": (-10, ">=")}, selected_param="n_bad_els")
[…]

### 2019-07-28 mlmc_random_frac/mlmc_process.py
Fix minor bugs in mlmc plot.

@@ -58,7 +58,7 @@ class Process(base_process.Process):
         # self.result_text(mlmc)
         # self.plot_density(mlmc)
 
-        # self.plot_temp_power(mlmc_est)
+        self.plot_temp_power(mlmc_est)
 
         mlmc_est.mlmc.clean_select()
         mlmc_est.mlmc.select_values({"n_bad_els": (-10, ">=")}, selected_param="n_bad_els")
@@ -114,8 +114,10 @@ class Process(base_process.Process):
                 bad_temp_ele_avg += n_bad_elements[i]
             else:
                 good_temp_ele_avg += n_bad_elements[i]
-        bad_temp_ele_avg /= temp_v_ele_sum[0][0]
-        good_temp_ele_avg /= temp_v_ele_sum[1][0]
+        if temp_v_ele_sum[0][0] != 0:
+            bad_temp_ele_avg /= temp_v_ele_sum[0][0]
+        if temp_v_ele_sum[1][0] != 0:
+            good_temp_ele_avg /= temp_v_ele_sum[1][0]
         print("bad_temp EX: ", bad_temp_ele_avg)
         print("good_temp EX: ", good_temp_ele_avg)
 
@@ -126,8 +128,10 @@ class Process(base_process.Process):
                 bad_temp_ele_var += (n_bad_elements[i]-bad_temp_ele_avg)**2
             else:
                 good_temp_ele_var += (n_bad_elements[i]-good_temp_ele_avg)**2
-        bad_temp_ele_var /= temp_v_ele_sum[0][0]
-        good_temp_ele_var /= temp_v_ele_sum[1][0]
+        if temp_v_ele_sum[0][0] != 0:
+            bad_temp_ele_var /= temp_v_ele_sum[0][0]
+        if temp_v_ele_sum[1][0] != 0:
+            good_temp_ele_var /= temp_v_ele_sum[1][0]
         print("bad_temp varX: ", bad_temp_ele_var)
         print("good_temp varX: ", good_temp_ele_var)
         print("bad_temp s: ", np.sqrt(bad_temp_ele_var))
@@ -277,7 +281,7 @@ class Process(base_process.Process):
         """
         n_moments = 3
         mlmc_est.mlmc.clean_select()
-        mlmc_est.mlmc.select_values({param_name: (0, ">=")})
+        mlmc_est.mlmc.select_values(None, param_name)
         domain = Estimate.estimate_domain(mlmc_est.mlmc)
[…]

### 2019-07-28 mlmc_random_frac/mlmc_process.py
Fix in mlmc

@@ -22,146 +22,6 @@ def load_config_dict():
 
 
 class Process(base_process.Process):
-    def run(self):
-        """
-        Run mlmc
-        :return: None
-        """
-        os.makedirs(self.work_dir, mode=0o775, exist_ok=True)
-        self.n_moments = 10
-
-        mlmc_list = []
-        # Run one level Monte-Carlo method
-        for nl in [1]:  # , 2, 3, 4,5, 7, 9]:
-            mlmc = self.setup_config(nl, clean=True)
-
-            # self.n_sample_estimate(mlmc)
-            self.generate_jobs(mlmc, n_samples=[self.mc_samples])
-            mlmc_list.append(mlmc)
-
-        self.all_collect(mlmc_list)
-
-
-    def get_samples(self, mlmc_est, quantity):
-        mlmc_est.mlmc.select_values(None, selected_param=quantity)
-        q_array = mlmc_est.mlmc.levels[0].sample_values
-        mlmc_est.mlmc.clean_select()
-        return q_array[:,0,:]
-
-
-    def select_samples_with_good_values(self, mlmc_est):
-        # determine samples with correct temperature
-        n_bad_els = self.get_samples(mlmc_est, 'n_bad_els')
-        temp_min = self.get_samples(mlmc_est, 'temp_min')
-        temp_max = self.get_samples(mlmc_est, 'temp_max')
-        temp = self.get_samples(mlmc_est, 'temp')
-
-        abs_zero_temp = 273.15
-        MIN_T = 250
-        MAX_T = 470
-        temp_good_min = np.min(temp_min[:, :], axis=1) > MIN_T
-        temp_good_max = np.max(temp_max[:, :], axis=1) < MAX_T
-        temp_good = np.logical_and(np.min(temp, axis=1) > MIN_T - 273.15, np.max(temp, axis=1) < MAX_T - 273.15)
-        no_bad_els = n_bad_els[:, 0] == 0
-        good_mask = np.logical_and(temp_good_min, temp_good_max)
-        good_mask = np.logical_and(good_mask, no_bad_els)
-        good_mask = np.logical_and(good_mask, temp_good)
-        good_indices = np.arange(0, len(good_mask))[good_mask]
-        bad_indices = np.arange(0, len(good_mask))[~good_mask]
-        mlmc_est.mlmc.subsample_by_indices(good_indices)
-
-
-    def process(self):
-        """
-        Use collected data
-        :return: None
-        """
[…]

### 2019-07-28 mlmc_random_frac/mlmc_process.py
power mask

@@ -255,23 +255,30 @@ class Process(base_process.Process):
         temp_min = self.get_samples(mlmc_est, 'temp_min')
         temp_max = self.get_samples(mlmc_est, 'temp_max')
         temp = self.get_samples(mlmc_est, 'temp')
+        power = self.get_samples(mlmc_est, 'power')
 
         abs_zero_temp = 273.15
         MIN_T = 250
         MAX_T = 470
+        min_power= 1e-2
+        max_power= 10
         temp_good_min = np.min(temp_min[:, :], axis=1) > MIN_T
         temp_good_max = np.max(temp_max[:, :], axis=1) < MAX_T
         temp_good = np.logical_and(np.min(temp, axis=1) > MIN_T - 273.15, np.max(temp, axis=1) < MAX_T - 273.15)
+        power_good = np.logical_and(np.min(power, axis=1) > min_power, np.max(power, axis=1) < max_power)
         no_bad_els = n_bad_els[:, 0] == 0
         good_mask = np.logical_and(temp_good_min, temp_good_max)
         good_mask = np.logical_and(good_mask, no_bad_els)
         good_mask = np.logical_and(good_mask, temp_good)
+        good_mask = np.logical_and(good_mask, power_good)
         good_indices = np.arange(0, len(good_mask))[good_mask]
         bad_indices = np.arange(0, len(good_mask))[~good_mask]
         assert not np.any(np.isnan(temp_max[good_indices]))
         assert not np.any(np.isnan(temp_min[good_indices]))
         assert not np.any(np.isnan(temp[good_indices]))
         mlmc_est.mlmc.subsample_by_indices(good_indices)
+
+        print(np.sort(power[:, 30]))
         self.good_indices = good_indices
 
 
@@ -292,10 +299,18 @@ class Process(base_process.Process):
         print(samples.shape)
         #print(np.any(np.isnan(samples), axis=1))
         domain = Estimate.estimate_domain(mlmc_est.mlmc)
+        domain_diff = domain[1] - domain[0]
+        print(domain)
         moments_fn = Monomial(n_moments, domain, False, ref_domain=domain)
-        means, vars = mlmc_est.estimate_moments(moments_fn)
 
-        return means, vars
+        N = mlmc_est.mlmc.n_samples[0]
+        mom_means, mom_vars = mlmc_est.estimate_moments(moments_fn)
+        q_mean = mom_means[:, 1]
[…]

### 2019-07-28 mlmc_random_frac/mlmc_process.py
Plot second graph for 03 the model.

@@ -168,12 +168,17 @@ class Process(base_process.Process):
         # self.result_text(mlmc)
         # self.plot_density(mlmc)
         print("N all samples: ", mlmc_est.mlmc.n_samples[0])
-        self.select_samples_with_good_values(mlmc_est)
-        print("N good samples: ", mlmc_est.mlmc.n_samples[0])
-        self.plot_temp_power(mlmc_est)
+        th_02_model_params = {"power": "power", "temp": "temp", "temp_min": "temp_min", "temp_max": "temp_max"}
+        self.select_samples_with_good_values(mlmc_est, th_02_model_params)
+        print("N good samples [th_02_model]: ", mlmc_est.mlmc.n_samples[0])
+        self.plot_temp_power(mlmc_est, th_02_model_params)
 
-
-        self.plot_temp_power(mlmc_est)
+        mlmc_est.mlmc.clean_select()
+        mlmc_est.mlmc.clean_subsamples()
+        th_03_model_params = {"power": "power_ref", "temp": "temp_ref", "temp_min": "temp_min_ref", "temp_max": "temp_max_ref"}
+        self.select_samples_with_good_values(mlmc_est, th_03_model_params)
+        print("N good samples [th_03_model]: ", mlmc_est.mlmc.n_samples[0])
+        self.plot_temp_power(mlmc_est, th_03_model_params)
 
         # n_samples = int(mlmc_est.mlmc.n_samples)
         # temp_v_ele = np.zeros((2 * n_samples, 2 * n_samples), dtype=int)
@@ -253,13 +258,19 @@ class Process(base_process.Process):
         return q_array[:,0,:]
 
 
-    def select_samples_with_good_values(self, mlmc_est):
+    def select_samples_with_good_values(self, mlmc_est, result_params):
+        """
+        Selects samples according to the temperature, power, mesh results.
+
+        :param result_params : Dictionary of names of parameters {power :, temp: , temp_min: , temp_max: }
+        :return: None
+        """
         # determine samples with correct temperature
         n_bad_els = self.get_samples(mlmc_est, 'n_bad_els')
-        temp_min = self.get_samples(mlmc_est, 'temp_min')
-        temp_max = self.get_samples(mlmc_est, 'temp_max')
-        temp = self.get_samples(mlmc_est, 'temp')
-        power = self.get_samples(mlmc_est, 'power')
+        temp_min = self.get_samples(mlmc_est, result_params['temp_min'])
[…]

### 2019-07-30 mlmc_random_frac/process.py
Cut max fracture cross section in process.

@@ -367,11 +367,14 @@ def prepare_th_input(config_dict):
     init_fr_K = float(config_dict['hm_params']['fr_conductivity'])
     init_bulk_K = float(config_dict['hm_params']['bulk_conductivity'])
     min_fr_cross_section = float(config_dict['th_params']['min_fr_cross_section'])
+    max_fr_cross_section = float(config_dict['th_params']['max_fr_cross_section'])
 
     time_idx = 1
     time, field_cs = mesh.element_data['cross_section_updated'][time_idx]
 
+    # cut small and large values of cross-section
     cs = np.maximum(np.array([v[0] for v in field_cs.values()]), min_fr_cross_section)
+    cs = np.minimum(cs, max_fr_cross_section)
 
     K = np.where(
         cs == 1.0,      # condition

### 2019-07-30 mlmc_random_frac/rand_frac_simulation.py
add_deep_SKB_data

@@ -268,6 +268,7 @@ class RandomFracSimulation(Simulation):
         eb_file = "energy_balance.yaml"
         heat_file = "Heat_AdvectionDiffusion_region_stat.yaml"
         heal_file = "random_fractures_heal_stats.yaml"
+        fr_param_file = "fr_param_output.yaml"
         extract_series = self._extract_time_series
 
         th_dir = "output_02_th/"
@@ -288,6 +289,7 @@ class RandomFracSimulation(Simulation):
             Quantity("temp_min_ref", ref_dir + heat_file, extract_series, [regions, 'min', 0]),
             Quantity("temp_max_ref", ref_dir + heat_file, extract_series, [regions, 'max', 0]),
 
+            #Quantity("median_cs", fr_param_file
             Quantity("n_bad_els", heal_file, self.get_heal_stat, [])
         ]
         return quantities

### 2019-07-30 mlmc_random_frac/rand_frac_simulation.py
Add extracted quantities

@@ -222,11 +222,18 @@ class RandomFracSimulation(Simulation):
         series = [np.array(region_series, dtype=float) for region_series in reg_series.values()]
         return series
 
-
-    def get_heal_stat(self, f):
+    def extract_yaml_const(self, f, *keys):
         with open(f, "r") as f:
-            stat_doc = yaml.safe_load(f)
-            return  len(stat_doc["flow_stats"]["bad_elements"]) + len(stat_doc["gamma_stats"]["bad_elements"])
+            content = yaml.safe_load(f)
+            for k in keys:
+                content = content[k]
+            assert np.isscalar(content)
+            return content
+
+    def get_heal_stat(self, fname):
+        with open(fname, "r") as f:
+            content = yaml.safe_load(f)
+            return  len(content["flow_stats"]["bad_elements"]) + len(content["gamma_stats"]["bad_elements"])
 
 
     def compute_th_quantities(self, power, temp, flux, t_min, t_max):
@@ -249,6 +256,10 @@ class RandomFracSimulation(Simulation):
         flux.value = sum_flux
         t_min.value = temp_min
         t_max.value = temp_max
+
+    def get_fr_param(self, f, ):
+        with open(f, "r") as f:
+            stat_doc = yaml.safe_load(f)
 
     def manipulate_quantities(self, q_dict):
         # manipulate values
@@ -279,17 +290,32 @@ class RandomFracSimulation(Simulation):
 
             Quantity("power", th_dir + eb_file, extract_series, [bc_regions, 'data', 0]),
             Quantity("temp", th_dir + heat_file, extract_series, [out_regions, 'average', 0]),
-            Quantity("fluxes", th_dir + wb_file, extract_series, [out_regions, 'data', 0]),
             Quantity("temp_min", th_dir + heat_file, extract_series, [regions, 'min', 0]),
             Quantity("temp_max", th_dir + heat_file, extract_series, [regions, 'max', 0]),
+            Quantity("fluxes", th_dir + wb_file, extract_series, [out_regions, 'data', 0]),
+            Quantity("bc_flux_bulk", th_dir + wb_file, extract_series, [['.right_well'], 'data', 0]),
+            Quantity("bc_flux_fr", th_dir + wb_file, extract_series, [['.fr_right_well'], 'data', 0]),
[…]

### 2019-07-30 mlmc_random_frac/rand_frac_simulation.py
Fix extract results.

@@ -295,8 +295,8 @@ class RandomFracSimulation(Simulation):
             Quantity("fluxes", th_dir + wb_file, extract_series, [out_regions, 'data', 0]),
             Quantity("bc_flux_bulk", th_dir + wb_file, extract_series, [['.right_well'], 'data', 0]),
             Quantity("bc_flux_fr", th_dir + wb_file, extract_series, [['.fr_right_well'], 'data', 0]),
-            Quantity("bc_influx_bulk_ref", ref_dir + wb_file, extract_series, [['.left_well'], 'data', 0]),
-            Quantity("bc_influx_fr_ref", ref_dir + wb_file, extract_series, [['.fr_left_well'], 'data', 0]),
+            Quantity("bc_influx_bulk", ref_dir + wb_file, extract_series, [['.left_well'], 'data', 0]),
+            Quantity("bc_influx_fr", ref_dir + wb_file, extract_series, [['.fr_left_well'], 'data', 0]),
 
             Quantity("power_ref", ref_dir + eb_file, extract_series, [bc_regions, 'data', 0]),
             Quantity("temp_ref", ref_dir + heat_file, extract_series, [out_regions, 'average', 0]),

### 2019-07-30 mlmc_random_frac/mlmc_process.py
Improve temp and power histograms.

@@ -187,8 +187,8 @@ class Process(base_process.Process):
         # self.plot_temp_power(mlmc_est, th_02_model_params)
         # self.plot_temp_power(mlmc_est, th_03_model_params)
 
-        self.plot_histogram(mlmc_est, 'temp')
-        self.plot_histogram(mlmc_est, 'power')
+        self.plot_histogram(mlmc_est, 'temp', "temp_histogram", "Temperature [$^\circ$C]", bins=30)
+        self.plot_histogram(mlmc_est, 'power', "power_histogram", "Power [MW]", bins=30)
 
         self.plot_temp_ref_comparison(mlmc_est)
         self.plot_power_ref_comparison(mlmc_est)
@@ -329,10 +329,10 @@ class Process(base_process.Process):
         power = self.get_samples(mlmc_est, result_params['power'])
 
         abs_zero_temp = 273.15
-        MIN_T = 250
+        MIN_T = 273.15
         MAX_T = 470
         min_power= 1e-2
-        max_power= 10
+        max_power= 50
         temp_good_min = np.min(temp_min[:, :], axis=1) > MIN_T
         temp_good_max = np.max(temp_max[:, :], axis=1) < MAX_T
         temp_good = np.logical_and(np.min(temp, axis=1) > MIN_T - 273.15, np.max(temp, axis=1) < MAX_T - 273.15)
@@ -421,7 +421,7 @@ class Process(base_process.Process):
         fig, ax1 = plt.subplots()
         temp_color = 'red'
         ax1.set_xlabel('time [y]')
-        ax1.set_ylabel('Temperature [C deg]', color=temp_color)
+        ax1.set_ylabel('Temperature [$^\circ$C]', color=temp_color)
         ax1.tick_params(axis='y', labelcolor=temp_color)
         self.plot_param(mlmc_est, ax1, times, temp_color, result_params['temp'])
 
@@ -436,16 +436,26 @@ class Process(base_process.Process):
         fig.savefig("temp_power.pdf")
         plt.show()
 
-    def plot_histogram(self, mlmc_est, quantity):
+    def plot_histogram(self, mlmc_est, quantity, name, ylabel, bins):
         fig, ax1 = plt.subplots()
+        ax1.set_xlabel(ylabel)
+        ax1.set_ylabel('Frequency [-]')
+        ax1.tick_params(axis='y', labelcolor='black')
+
[…]

### 2019-07-30 mlmc_random_frac/mlmc_process.py
Fix variance calculation and plot

@@ -178,20 +178,25 @@ class Process(base_process.Process):
         # self.plot_density(mlmc)
 
         mlmc_est.mlmc.clean_select()
-        print("N all samples: ", mlmc_est.mlmc.n_samples[0])
-        th_02_model_params = {"power": "power", "temp": "temp", "temp_min": "temp_min", "temp_max": "temp_max"}
-        th_03_model_params = {"power": "power_ref", "temp": "temp_ref", "temp_min": "temp_min_ref", "temp_max": "temp_max_ref"}
-        self.select_samples_with_good_values(mlmc_est, th_02_model_params)
-        self.select_samples_with_good_values(mlmc_est, th_03_model_params, add=1)
-        print("N good samples: ", mlmc_est.mlmc.n_samples[0])
+
+        th_02_model_params = ["power", "temp", "temp_min", "temp_max"]
+        th_03_model_params = [n + "_ref" for n in th_02_model_params]
+        print("02_th - stimulated EGS")
+        good_mask = self.select_samples_with_good_values(mlmc_est, th_02_model_params)
+        print("03_th - reference EGS")
+        good_mask = self.select_samples_with_good_values(mlmc_est, th_03_model_params, good_mask=good_mask)
+        #good_mask[np.random.choice(len(good_mask), int(len(good_mask)*0.93))] = False
+        #print("N Random choice: ", sum(good_mask))
+        mlmc_est.mlmc.subsample_by_indices(good_mask)
+
         # self.plot_temp_power(mlmc_est, th_02_model_params)
         # self.plot_temp_power(mlmc_est, th_03_model_params)
 
         self.plot_histogram(mlmc_est, 'temp')
-        self.plot_histogram(mlmc_est, 'power')
+        #self.plot_histogram(mlmc_est, 'power')
 
         self.plot_temp_ref_comparison(mlmc_est)
-        self.plot_power_ref_comparison(mlmc_est)
+        #self.plot_power_ref_comparison(mlmc_est)
 
         # n_samples = int(mlmc_est.mlmc.n_samples)
         # print("N samples: ", n_samples)
@@ -314,7 +319,7 @@ class Process(base_process.Process):
         return q_array[:,0,:]
 
 
-    def select_samples_with_good_values(self, mlmc_est, result_params, add=0):
+    def select_samples_with_good_values(self, mlmc_est, result_params, good_mask=None):
         """
[…]

### 2019-07-30 mlmc_random_frac/mlmc_process.py
Add histogram plots.

@@ -187,6 +187,10 @@ class Process(base_process.Process):
         # self.plot_temp_power(mlmc_est, th_02_model_params)
         # self.plot_temp_power(mlmc_est, th_03_model_params)
 
+        self.plot_histogram_const(mlmc_est, 'fr_cs_med', "fr_cs_median", "Median of $\delta_f$ [mm]", bins=30, scale=1000)
+        self.plot_histogram_const(mlmc_est, 'fr_cond_med', "fr_cond_median", "Median of $k_f$ [m.s$^{-1}$]", bins=30, scale=1)
+        self.plot_histogram_flux(mlmc_est, 'bc_flux', "total_flux_histogram", bins=30)
+
         self.plot_histogram(mlmc_est, 'temp', "temp_histogram", "Temperature [$^\circ$C]", bins=30)
         self.plot_histogram(mlmc_est, 'power', "power_histogram", "Power [MW]", bins=30)
 
@@ -449,13 +453,75 @@ class Process(base_process.Process):
         for t in [1, int((t_min + t_max)/2), t_max]:
             print(t)
             label = "t = " + str(t) + " y"
-            ax1.hist(q[:, t], alpha=0.5, bins=bins, label=label)
+            ax1.hist(q[:, t], alpha=0.5, bins=bins, label=label) #edgecolor='white')
 
         ns = "N = " + str(mlmc_est.mlmc.n_samples[0])
         ax1.set_title(ns)
         ax1.legend()
         fig.savefig(os.path.join(self.work_dir, name + ".pdf"))
         fig.savefig(os.path.join(self.work_dir, name + ".png"))
+        plt.show()
+
+    def plot_histogram_const(self, mlmc_est, quantity, name, ylabel, bins=20, scale=1.0):
+        fig, ax1 = plt.subplots()
+        ax1.set_xlabel(ylabel)
+        ax1.set_ylabel('Frequency [-]')
+        ax1.tick_params(axis='y', labelcolor='black')
+
+        q = self.get_samples(mlmc_est, quantity)
+
+        ax1.hist(scale * q[:, 0], alpha=1.0, bins=bins) #edgecolor='white')
+
+        ns = "N = " + str(mlmc_est.mlmc.n_samples[0])
+        ax1.set_title(ns)
+        ax1.legend()
+        fig.savefig(os.path.join(self.work_dir, name + ".pdf"))
[…]

### 2019-07-30 mlmc_random_frac/mlmc_process.py
Output temp and power comparison to file.

@@ -544,6 +544,8 @@ class Process(base_process.Process):
 
         ax1.legend(["MH+TH", "var(T)", "TH only", "var(T)"])
         fig.tight_layout()  # otherwise the right y-label is slightly clipped
+        fig.savefig(os.path.join(self.work_dir, "temp_ref_comparison.pdf"))
+        fig.savefig(os.path.join(self.work_dir, "temp_ref_comparison.png"))
         plt.show()
 
     def plot_power_ref_comparison(self, mlmc_est):
@@ -566,6 +568,8 @@ class Process(base_process.Process):
 
         ax1.legend(["MH+TH", "var(T)", "TH only", "var(T)"])
         fig.tight_layout()  # otherwise the right y-label is slightly clipped
+        fig.savefig(os.path.join(self.work_dir, "power_ref_comparison.pdf"))
+        fig.savefig(os.path.join(self.work_dir, "power_ref_comparison.png"))
         plt.show()
 
 

### 2019-07-30 mlmc_random_frac/mlmc_process.py
update comparison plot legend

@@ -502,8 +502,10 @@ class Process(base_process.Process):
         self.plot_param(mlmc_est, ax1, times, 'red', 'temp')
         self.plot_param(mlmc_est, ax1, times, 'orange', 'temp_ref')
 
-        ax1.legend(["MH+TH", "var(T)", "TH only", "var(T)"])
+        ax1.legend(["stimulated - mean", "stimulated - std", "reference - mean", "reference - std"])
         fig.tight_layout()  # otherwise the right y-label is slightly clipped
+        fig.savefig(os.path.join(self.work_dir, "temp_comparison.pdf"))
+        fig.savefig(os.path.join(self.work_dir, "temp_comparison.png"))
         plt.show()
 
     def plot_power_ref_comparison(self, mlmc_est):
@@ -524,8 +526,10 @@ class Process(base_process.Process):
 
         self.plot_param(mlmc_est, ax1, times, 'forestgreen', 'power_ref')
 
-        ax1.legend(["MH+TH", "var(T)", "TH only", "var(T)"])
+        ax1.legend(["stimulated - mean", "stimulated - std", "reference - mean", "reference - std"])
         fig.tight_layout()  # otherwise the right y-label is slightly clipped
+        fig.savefig(os.path.join(self.work_dir, "power_comparison.pdf"))
+        fig.savefig(os.path.join(self.work_dir, "power_comparison.png"))
         plt.show()
 
 

### 2019-08-01 mlmc_random_frac/output_scenarios/output_por_disp_04/process.py
Add new results.

@@ -1,0 +1,562 @@
+import sys
+import os
+script_dir = os.path.dirname(os.path.realpath(__file__))
+sys.path.append(os.path.join(script_dir, '../../MLMC/src'))
+sys.path.append(os.path.join(script_dir, '../../dfn/src'))
+
+import shutil
+import subprocess
+import yaml
+import attr
+import numpy as np
+import collections
+import gmsh_io
+# import matplotlib.pyplot as plt
+
+import fracture
+
+# TODO:
+# - enforce creation of empty physical groups, or creation of empty regions in the flow input
+# - speedup mechanics
+
+@attr.s(auto_attribs=True)
+class ValueDescription:
+    time: float
+    position: str
+    quantity: str
+    unit: str
+
+
+
+def substitute_placeholders(file_in, file_out, params):
+    """
+    Substitute for placeholders of format '<name>' from the dict 'params'.
+    :param file_in: Template file.
+    :param file_out: Values substituted.
+    :param params: { 'name': value, ...}
+    """
+    used_params = []
+    with open(file_in, 'r') as src:
+        text = src.read()
+    for name, value in params.items():
+        placeholder = '<%s>' % name
+        n_repl = text.count(placeholder)
+        if n_repl > 0:
+            used_params.append(name)
+            text = text.replace(placeholder, str(value))
+    with open(file_out, 'w') as dst:
+        dst.write(text)
+    return used_params
+
+def to_polar(x, y, z):
+    rho = np.sqrt(x ** 2 + y ** 2)
+    phi = np.arctan2(y, x)
+    if z > 0:
+        phi += np.pi
+    return (phi, rho)
+
+def plot_fr_orientation(fractures):
+    family_dict = collections.defaultdict(list)
+    for fr in fractures:
+        x, y, z = fracture.FisherOrientation.rotate(np.array([0,0,1]), axis=fr.rotation_axis, angle=fr.rotation_angle)[0]
+        family_dict[fr.region].append([
+            to_polar(z, y, x),
+            to_polar(z, x, -y),
+            to_polar(y, x, z)
+            ])
+
+    import matplotlib.pyplot as plt
+    fig, axes = plt.subplots(1, 3, subplot_kw=dict(projection='polar'))
+    for name, data in family_dict.items():
+        # data shape = (N, 3, 2)
[…]

### 2019-08-01 paper.tex
Updates from Overleaf

@@
 \documentclass{article}
+\usepackage{graphicx}
 
@@
 \begin{abstract}
-
+Stochastic modeling of the hydraulic stimulation and long term operation of an enhanced geothermal system (EGS) is performed.
+The continuum-fracture approach is applied both for the stimulation model and for the operation model.
+Stochasticaly generated fractures and properties of the continuum are used in a simple hydro-mechanical model of the stimulation process.
+The resulting system of fractures with significant conductivity is used in the heat transfer model of the operation of the ground exchanger.
+The model is used to compute a single realization of key EGS properties: the output power and temperature.
+This complex sampling model is used within the Monte Carlo method predict probabilistic distribution for the EGS properties.
+The conceptual model and input data are loosely based on the geologic properties at the Litoměřice site in the Czech Republic.
 \end{abstract}
@@
 \section{Introduction}
-\todo{
-- problem definition - uncertainty of EGS output due to  fracture network and other properties\\
-- goal - capturing these uncertainties by stochastic modelling\\
-- basic research (find some related work for dealing with modelling EGS with fracture opening and/or stochastic DFN\\
-- paper structure\\
+As the natural hydro thermal resources are sparse in the Czech Republic, the EGS (enhanced geothermal system) concept is subject of recent research focused on usage of the geothermal energy.
+According to the available geological data (\cite{Capova2013}), the Litoměřice site is the most attractive due to hot crystalline rock in relatively shallow depth (about 120 \textdegree C in 5km depth).
+This motivates development of the poroelasticity model within the Flow123d (\cite{flow123d}) simulator for the transport processes in the fractured porous media and its application in the study of uncertainty of the EGS properties caused by the fracture system.
+
+The EGS approach depends on opening and reconnecting of the preexisting fracture network in the low permeable rock by the hydraulic stimulation and the temperature changes.
+Since the fracture network is a-priori unknown as well as spatial distribution of the key parameters, we should consider the operational parameters of the EGS as random variables.
+In order to predict their distribution the Monte Carlo method can be applied provided the forward model and the distribution of the input parameters.
+Considering the forward model, there are three main approaches are used for simulation of the transport processes in the fractured porous media, namely single continuum, discrete fracture networks and continuum-fracture approach.
+The heterogeneous single continuum fully coupled THM model was used  in \cite{Watanabe2009} to study uncertainty in the output temperature.
+Usage of Monte Carlo method is natural in the DFN approach see e.g.
+\cite{Ezzedine2008}.
+We adopt the continuum-fracture approach which is much less common especially for significant technical difficulties.
+In \cite{Doonechaly2016} this approach was used for sensitivity analysis of the EGS operation, but only for the fixed fracture network.
+
+We consider simple two stage forward model.
+First, the hydraulic stimulation is described by the hydro-mechanical poroelasticity model combining discrete fractures and continuum.
+Linear elasticity is used, contact conditions and thermal effects, however they may be significant, are not considered.
+Based on the stimulation model the changed aperture and conductivity fields on the fractures are determined.
+Then the evolution of the temperature and the raw power output is simulated using the combined fracture-continuum approach for the thermo-hydraulic model as well.
+A realistic stochastic model for the distribution of the fractures is considered, other parameters are fixed as we want to study particularly the influence of fracture variability.
+
+\section{Conceptual model}
+According to available geological data for the Litoměřice locality (see \cite{Capova2013}), the crystalline rock possibly consisting of granite, gneiss or mica schist 
+% zula, rula, svor
+is located in the depth
+below 3 km.
+Two geological scenarios predict temperatures 140 C and 146 C respectively in the depth 5 km.
+We consider in our model two vertical wells in distance 200 m with opened part of the length 100 m located in the target depth of 5 km.
+The simulation domain is a box with dimensions $600\times600\times600$ m covering only the close neighbourhood of the stimulated domain.
+The wells are placed in the center.
+In order to reduce the complexity of the computational mesh, we
+cut off cylinders of the radius 10 m at the wells positions and apply boundary conditions on their walls while avoiding very fine meshing at the wells scale.
+This approximation is discussed in detail in Section \ref{}.
+This simplified geometry is consistent with the simplification in the mathematical model as the precision of the model prediction is not of our interest in this work.
+We refer to the contribution paper \cite{Ralek} for the more EGS study for the Litoměřice locality.
+
+\section{Hydro-mechanical model}
+\label{sc:hm_model}
+During the hydraulic stimulation the fluid is injected into both wells in order to open the preexisting fractures.
+This leads to increase in the aperture of fractures, which is then used in the production phase model described in the next section.
+As the HM model of the bulk rock, we use the Biot poroelasticity equations:
+\eq{\label{eq:biot} \partial_t(Sp + \alpha\div\uu) +\div\vc q = 0,\qquad -\div\bbsigma + \tilde\alpha\nabla h = 0. }
+Here $\uu$ is the displacement [m], $p$ is the pressure head [m], $h=p+x_3$ is the piezo-metric head and the flux $\vc q$ [m.s -1] is given by the Darcy law
+\eq{ \vc q = - k\nabla h}
+with the hydraulic conductivity $k$. The coefficient $\tilde\alpha:=\alpha\varrho_l g$ where $\alpha$ is the Biot's effective stress parameter \cite{Biot1941},
+$\rho_l$ is liquid density and $g$ is gravitational acceleration.
+The stress tensor $\bbsigma$ [Pa] is given by the Hook's law
+for the isotropic material:
+\eq{ \bbsigma = \CC : \epsilon 
+= (\lambda_M I\otimes I +2\mu_M I):\epsilon, \quad \epsilon = (\nabla \uu^T + \nabla \uu)/2}
+with the Lamm\' e parameters: $\mu_M=E/(2(1+\nu))$ and $\lambda_M=E\nu/((1+\nu)(1-2\nu))$ [Pa].
+
+The continuum-fracture approach consists in integration of the governing equations across the volume of a thin fracture effectively reducing its dimension to a 2d surface.
+This procedure was introduced by \cite{martin_modeling_2005} for the Darcy flow and extended for general advection-diffusion process in \cite{brezina_analysis_2015}.
+For the purpose of this work, the same approach was applied also to the equations of linear elasticity.
+We introduce the unit normal vector $\vc\nu$ to the fracture plane with arbitrary but fixed orientation.
+For a general quantity $s$ that is defined on both faces of the fracture volume, we define the jump $\jmp{s}:=s^+-s^-$, where $s^\pm$ are the traces of $s$ on the faces. 
+Then, integrating \eqref{eq:biot} across the fracture aperture results in the following system of equations:
+\eq{\label{eq:biot_f}\left.\begin{aligned}
+\delta\left(\partial_t\left(S_fp_f + \alpha_f\left(\div_\tau\uu_f+\tfrac1\delta\jmp{\uu}\cdot\vc\nu\right)\right) - k_f\Delta_\tau h_f\right) &= \jmp{F}, \\
+\delta\left(-\div_\tau\CC_f(\nabla_\tau\uu_f+\tfrac1\delta\vc\nu\otimes\jmp{\uu}) + \tilde\alpha_f\nabla_\tau h_f\right) &= \jmp{\vc Q}.
+\end{aligned}\right\} }
+Here $p_f$ and $\uu_f$ denote the average pressure head and the average displacement on the fractures, respectively.
+The subscript $\tau$ indicates tangential operators on the fracture plane, $\otimes$ is the outer product and $\tilde\alpha_f:=\alpha_f\varrho_l g$.
+The jumps $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are the water sources and the forces, respectively,  that arise from the approximation of derivatives in the normal direction:
+\eq{ 
+\label{eq:fr_coupling}
+\left.\begin{aligned}
+F^\pm &:= \pm\frac2\delta k_f(h^\pm-h_f),\\
+\vc Q^\pm &:= \pm\frac2\delta\CC_f(\nabla_\tau\uu^\pm + \vc\nu\otimes(\uu^\pm-\uu_f))\vc\nu.
+%\pm\frac2\delta\left(\mu_f(\uu^\pm-\uu_f)+(\mu_f+\lambda_f)((\uu^\pm-\uu_f)\cdot\vc\nu)\vc\nu\right) +\mu_f(\nabla_\tau\uu^\pm)^\top\vc\nu + \lambda_f(\div_\tau\uu^\pm)\vc\nu \pm\tilde\alpha_f p_f\vc\nu.
+\end{aligned}\right\} 
+}
+System \eqref{eq:biot} and \eqref{eq:biot_f} is complemented by the boundary conditions for the bulk domain prescribed on the fracture faces:
+\eq{ 
+\label{eq:fr_coupling_bulk}
+k\nabla h^\pm\cdot\vc\nu = F^\pm,\qquad \left(\CC(\nabla\uu^\pm) - \tilde\alpha h^\pm I\right)\vc\nu = \vc Q^\pm. 
 }
 
-\section{Mathematical model}
-According to available geological data of the Litoměřice locality, the rock in the depth of several km is considered crystallinic, possibly consisting of granite, gneiss or mica schist \cite{Capova2013}.
-% zula, rula, svor
-We also refer to the EGS study for this site in  contribution \cite{Ralek}.
+For the sake of simplicity, we stimulate both wells simultaneously which results in very similar change of in the fracture aperture as if the wells were stimulated one by one.
+Zero piezometric head and zero displacement are used as the initial conditions.
+The liquid is injected for 1 day under the stimulation pressure of 50 MPa (piezometric head $h_s=5e3$ m).
+This pressure is not applied directly on the surface of the well's cylinders, but through the Robin boundary condition with the transition parameter derived from the approximate analytical solution at the vicinity of the well.
+The zero piezometric head is prescribed on the box sides, while the no flow condition is applied on the top and bottom face.
+Following boundary conditions were applied for the elasticity equation: zero displacement on the bottom of the box, zero normal displacement on the well's cylinders and zero traction elsewhere.
 
-\todo{short characterization of the locality, rock types,  temperature, reference to Ralek, identification of the model parameters postpone after model introduction.}
+\section{Thermo-hydraulic model}
+The TH model describes the heat transport in the EGS through the operation phase within 30 years after the hydraulic stimulation performed according to Section \ref{sc:hm_model}.
+In this scenario, we consider a steady Darcy flow:
+\eq{ \label{th_darcy}
+    \div \vc q = 0, \quad \div  \vc q_f =  \jmp F,\quad
+    \vc q_f = -\tilde\delta \tilde k_f \nabla h_f
+}
+for the bulk domain and fractures, respectively. Where $\tilde \delta$ is the fracture aperture updated according to the HM model.
+The conductivity of the fractures is also adjusted according to the cubic law $\tilde k_f = k_f\tilde\delta_f^2 / \delta_f^2$
+The communication between the fractures and the bulk rock is defined by the conditions \eqref{eq:fr_coupling} and \eqref{eq:fr_coupling_bulk}.
 
-\todo{Conceptual model:
+The heat transport model is sequentially coupled to the hydraulic model using computed Darcy flux $\vc q$ and $\vc q_f$.
+It assumes thermodynamic equilibrium between the liquid and the rock, the advection-diffusion equation for the bulk and fracture domain reads:
+\eq{
+\left.\begin{aligned}
+&\partial_t\left(\tilde s T \right) + \div(\varrho_l c_l T \vc q) - \div(\lambda\nabla T) = 0,\\
+&\partial_t\left(\delta \tilde s_f T_f \right) + \div(\varrho_l c_l T_f \vc q_f) - \div(\delta\lambda_f\nabla T_f) = \jmp{G}.
+\end{aligned}\right\}
+}
+Here $T$ [$\circ$C] is the temperature, $\varrho_l$, $\varrho_s$ is the density, $c_l$, $c_s$ is the heat capacity.
+\eq{
+\tilde s = \vartheta \varrho_l c_l + (1-\vartheta)\varrho_s c_s,\quad 
+\lambda= \vartheta \lambda_l + (1-\vartheta)\lambda_s
+}
+where $\theta$ is the porosity and $\lambda_l,\lambda_s$ is the thermal conductivity.
 
-\begin{itemize}
-    \item modeling just small box around the heat exchanger
-    \item fracking considered as just opening existing fractures;simple HM model
-    \item deformation considered permanent
-    \item longterm heat transfer as separated problem;
-          TH model sequentialy coupled
-\end{itemize}
+Similarly, to \eqref{eq:fr_coupling} we obtain the relation for the normal heat fluxes on the fracture faces:
+\eq{ 
+G^\pm &:= \pm\frac2\delta \lambda_f(T^\pm-T_f) + 
+\varrho_l c_l F^\pm T_q,
+}
+where $T_q$ is equal to the bulk of fracture temperature as the normal Darcy flux $F^\pm$ is positive or negative, respectively. The system is completed by the bulk boundary conditions:
+\eq{ 
+- (\rho_l c_l \vc q T_q - \lambda\nabla T^\pm)\cdot\vc\nu = G^\pm.
 }
 
-\subsection{Hydro-mechanical model}
-\label{sc:hm_model}
+The liquid is continuously injected into the left well under the injection pressure 0.5 MPa ($h_i = 50$ m) and pumped out to the level of the EGS 
+($h_o = -5000$ m). No flux condition is used on the outer boundary. 
+The injecting temperature is set to $15\ ^\circ$C, the natural convection 
+heat flux is prescribed on the production well. The initial temperature field is considered on the outer boundary, which is given by 
+ $10\ ^\circ$C at the Earth surface and the geothermal gradient $30\ ^\circ$C.km${}^{-1}$.
 
-In the first model scenario the fluid is injected to both wells in order to open the preexisting fractures.
-This leads to increase in the cross-section of fractures, which is then used in the computation of long-term operation.
-As the HM model of 3d rock we use the Biot poroelasticity equations:
-\eq{\label{eq:biot} \partial_t(Sp + \alpha\div\uu) +\div\vc q = 0,\qquad -\div\bbsigma + \tilde\alpha\nabla p = 0. }
-Here $p=h-x_3$ is the pressure head [m], $\uu$ is the displacement [m], $\tilde\alpha:=\alpha\varrho_l g$ and $\alpha$ is the Biot's effective stress parameter \cite{Biot1941}.
-The flux $\vc q$ [m.s${}^{-1}$] and the stress tensor $\bbsigma$ [Pa] are given by the Darcy and Hooke laws:
-\eq{ \vc q = - k\nabla p, \qquad \bbsigma = \CC(\nabla\uu) := \mu(\nabla\uu+\nabla\uu^\top) + \lambda(\II:\nabla\uu)\II, }
-where $\mu=E/(2(1+\nu))$ and $\lambda=E\nu/((1+\nu)(1-2\nu))$ are the Lamé parameters [Pa], $\II$ is the identity matrix and ``:'' denotes the scalar product of matrices.
+%We use the following system of heat equation and steady-state Darcy flow with coupling terms connecting the rock to the fractures:
 
-For the 2d fractures we apply a dimension reduction procedure, which was originally developed in \cite{martin_modeling_2005} for the Darcy flow.
-We introduce the unit normal vector $\vc\nu$ to the fracture plane with arbitrary but fixed orientation.
-For a quantity, say $p$, which is defined on both sides of the fracture, we define the jump $\jmp{p}:=p^+-p^-$, where $p^\pm$ is the trace of $p$ on the positive/negative side of the fracture, determined by the orientation of $\vc\nu$.
-Then, integrating \eqref{eq:biot} across fracture aperture results in the following system of equations:
-\eq{\label{eq:biot_f}\left.\begin{aligned}
-\delta\left(\partial_t\left(S_fp_f + \alpha_f\left(\div_\tau\uu_f+\tfrac1\delta\jmp{\uu}\cdot\vc\nu\right)\right) - k_f\Delta_\tau p_f\right) &= \jmp{F}, \\
-\delta\left(-\div_\tau\CC_f(\nabla_\tau\uu_f+\tfrac1\delta\vc\nu\otimes\jmp{\uu}) + \tilde\alpha_f\nabla_\tau p_f\right) &= \jmp{\vc Q}.
-\end{aligned}\right\} }
-Here $p_f$ and $\uu_f$ is the average pressure head and displacement in fractures, the subscript $\tau$ indicates tangential operators in the fracture plane, ``$\otimes$'' denotes the outer product and $\tilde\alpha_f:=\alpha_f\varrho_l g$.
-The expressions $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are sources and forces that arise from the approximation of derivatives in the direction orthogonal to fracture plane:
-\eq{ \left.\begin{aligned}
-F^\pm &:= \pm\frac2\delta k_f(p^\pm-p_f),\\
-\vc Q^\pm &:= \pm\frac2\delta\CC_f(\nabla_\tau\uu^\pm + \vc\nu\otimes(\uu^\pm-\uu_f))\vc\nu.
-%\pm\frac2\delta\left(\mu_f(\uu^\pm-\uu_f)+(\mu_f+\lambda_f)((\uu^\pm-\uu_f)\cdot\vc\nu)\vc\nu\right) +\mu_f(\nabla_\tau\uu^\pm)^\top\vc\nu + \lambda_f(\div_\tau\uu^\pm)\vc\nu \pm\tilde\alpha_f p_f\vc\nu.
-\end{aligned}\right\} }
-Systems \eqref{eq:biot} and \eqref{eq:biot_f} are complemented by the following interface conditions on the fractures:
-\eq{ k\nabla p^\pm\cdot\vc\nu = F^\pm,\qquad \left(\CC(\nabla\uu^\pm) - \tilde\alpha p^\pm\II\right)\vc\nu = \vc Q^\pm. }
+% \subsection{Conductivity dependent on displacement}
+% \textcolor{red}{PE: If not longer, move it directly to the description of flow above.}
+% We consider that after releasing the fracking pressure the displacement of larger fractures persists. So the change in new porosity is
+% \[
+% \phi = \phi_0 + \alpha\div \vc u
+% \]
+% where $\phi_0$ is the porosity before deformation, c.f. time term in \eqref{eq:biot}. 
+% Using the Carman-Kozeny relation \cite{Kaviany1999}, 
+% % \[
+% %     k = \tilde{k}\frac{\phi^3}{(1-\phi)^2}
+% % \]
+% we can derive formula for the updated conductivity:
+% \[
+%     K = K_0\left(\frac{\phi}{\phi_0}\right)^3 \left(\frac{1-\phi_0}{1-\phi}\right)^2.
+% \]
 
-For simplicity we stimulate both wells simultaneously which produces very similar result in the change of fracture cross-section as if the wells were stimulated one by one.
-The liquid is injected for 1 day under the pressure 10 MPa (which corresponds to piezometric head $10^3$ m).
+\section{Geology and Model Parameters}
 
-Initial condition: zero piezometric head
+Most of the parameters used for computation are gathered in Table \ref{tab:params}.
+The mechanical properties of granitic rocks have been investigated e.g.
+in \cite{Ljunggren1985}, where the measured Young modulus $E$ varies around 50 GPa and Poisson's ratio $\nu$ is around 0.25.
+The Biot coefficient for different rocks is hard to obtain, we use the value $\alpha=0.6$ inspired by \cite{Selvadurai2019}, where $\alpha$ for Grimsel granite varies between 0.46 to 0.72.
+The reported values of the drained compressibility $\beta$ of several types of granite and orthogneiss are approx.
+$20\times 10^{-13}$ cm${}^2$.dyne${}^{-1}$ = $2\times10^{-11}$ Pa${}^{-1}$ \cite{Zisman1933}.
+As far as porosity $\vartheta$ is concerned, \cite{intera} reports values between 0.001 and 0.01 for metamorphic crystalline rocks, we shall consider the value 0.005.
+We assume the standard value of gravitational acceleration $g=9.81$ m.s${}^{-2}$.
+%Here and in what follows, subscript ``l'' indicates quantities associated to the liquid.
+%Similarly, subscript ``f'' will denote quantities for the fractures.
+Storativity $S$ is calculated from water compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$ and porosity following (\cite{Brace1968}, \cite{intera}):
+\[ S = \varrho_l g(\beta + \vartheta\beta_l) \approx 2\times 10^{-7} \mbox{ m}^{-1}. \]
 
-Boundary conditions:
-\begin{itemize}
-\item piezometric head $10^3$ m on surface of wells (including the intersections with fractures)
-\item piezometric head 0 m on lateral sides of the box
-\item no fluid flow through bottom and top part of the box and fracture tips
-\item zero displacement on bottom part of the box, zero displacement in the direction orthogonal to box sides
-\item zero traction elsewhere
-\end{itemize}
+The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
+The aperture of the fractures is considered constant in the input of the HM model according to \cite{Honza-clanek o laborratornich mikropuklinach??}.
+After the process of the hydraulic fractioning, the aperture and hydraulic conductivity are adjusted according to the displacement results in TH model in \eqref{th_darcy}.
+The hydraulic conductivity $k$ in the bulk rock is set similarly to \cite{Sperl2008} and \cite{Doonechaly2016}, where they also consider DFN and surrounding rock.
+Additionally we assume that the hydraulic fractioning impacts also the smaller fractures not included in our DFN,
+so we increase $k=10^{-8}$ of the bulk rock.
 
-\subsection{Thermo-hydraulic model}
+The hydraulic conductivity of fractures dominates than the one of intact rock by several orders of magnitude, we consider $k_f=10^{-3}$ m.s${}^{-1}$
+On the other hand, the elastic modulus of fracture is assumed much smaller, we take $E_f=50$ Pa.
+The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$.
+We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
 
-The second scenario corresponds to operation of the EGS within 30 years after hydraulic stimulation performed according to Sec. \ref{sc:hm_model}.
-The fluid is continuously injected into one well and pumped out of the second one under the pressure difference 2 MPa and with inlet temperature $15\ ^\circ$C.
+\begin{table}
+\centering
+\begin{tabular}{|l|c|c|c|}
+\hline
+ & rock & fractures & liquid\\\hline
+Young modulus $E$ [Pa] & $50\times10^9$ & 50 & --- \\
+Poisson's ratio $\nu$ [--] & 0.25 & 0.25 & --- \\
+Biot's coefficient $\alpha$ [--] & 0.6 & 0.6 & --- \\
+\hline
+Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
+Porosity $\vartheta$ [--] & $5\times10^{-3}$ & $0.8$ & --- \\
+Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ & $2\times10^{-11}$ & $5\times10^{-10}$ \\
+Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
+Aperture $\delta$ [m] & --- & *$5\times10^{-5}$ & --- \\
+Hydraulic conductivity $k$ [m.s${}^{-1}$] & *$10^{-9}$ & *$2.5\times10^{-3}$ & ---\\
+\hline
+Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
+Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
+\hline
+\end{tabular}
+\caption{Physical parameters of rock, fractures and water, used in the model. \\ *) these values hold for HM model.}
+\label{tab:params}
+\end{table}
 
-Initial condition:
-temperature corresponding to $10\ ^\circ$C at the Earth surface and geothermal gradient $30\ ^\circ$C.km${}^{-1}$
+%Papers about effective properties of fractured porous media: \cite{Snow1969}, \cite{Bonnet2001}, \cite{Olson2003}
 
-Boundary conditions:
-\begin{itemize}
-\item piezometric head $\pm10^2$ m on surface of wells ($+$ on the injection well, $-$ on the production well)
-\item no flux on the external boundary of the box and on fracture tips
-\item temperature $15\ ^\circ$C on surface of injection well
-\item initial temperature on lateral sides of the box where fluid flows into the box
-\item no heat flux through bottom and top part of the box and fracture tips
-\end{itemize}
+%\subsection{Parameters of EGS site and model geometry}
 
-We use the following system of heat equation and steady-state Darcy flow with coupling terms connecting the rock to the fractures:
-\[ ... \]
+%Number of wells: 2 \\
+%Depth: 5 km \\
+%Distance: 200 m \\
+%Well radius: $\rho = 15$ cm \\
+%Length of open part of wells: $L=60$ m \\
 
-\subsection{Conductivity dependent on displacement}
-We consider that after releasing the fracking pressure the displacement of larger fractures persists. So the change in new porosity is
-\[
-\phi = \phi_0 + \alpha\div \vc u
-\]
-where $\phi_0$ is the porosity before deformation, c.f. time term in \eqref{eq:biot}. 
-Using the Carman-Kozeny relation \cite{Kaviany1999}, 
-% \[
-%     k = \tilde{k}\frac{\phi^3}{(1-\phi)^2}
-% \]
-we can derive formula for the updated conductivity:
-\[
-    K = K_0\left(\frac{\phi}{\phi_0}\right)^3 \left(\frac{1-\phi_0}{1-\phi}\right)^2.
-\]
+% Pressure difference / water flux: 60 l/s
+
+%In our model, we consider pair of vertical wells with the radius $\rho=0.15$ m in the distance $D=200$ m.
+The open part of the length $L=60$ m is in the depth $5000$ m.
+The cube $600 \times 600 \times 600$ m around the open parts is considered as the computational domain.
+
+%{\bf The test geometry:} The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
 
 \subsection{Analytical models in the vicinity of the well}
+We assume the well radii be $\rho = 15$ cm which is difficult to capture in the computational mesh.
 In order to keep the mesh size treatable, we cut out of the computational domain two cylinders around the open well parts with a larger radius $R=10$ m.
@@
 \]
-where $r$ is the distance from the well and $\alpha$, $\beta$ are arbitrary constants such that $\hat h(\rho)=H_w$:
+where $r$ is the distance from the well and $\alpha$, $\beta$ are arbitrary constants such that $\hat h(\rho)=H_w$ holds.
 First, the flux from the well into the domain through the surface $C_\rho$ of the cylinder with radius $\rho$ and length $L$ is
@@
 
-\subsection{Geological properties}
+\subsection{Stochastic description of the fractures}
+The fracture stochastic model consists of the Fisher's distribution for the fracture orientation, the power law for the fracture size, Poisson process for the number of fractures and uniform distribution for the fracture centers.
+The square fractures are considers with the side $r/2$.
+The von Mises-Fischer distribution \cite{Fisher1993} is used to describe orientation of single family of fractures with given mean normal vector $\vc a$ and the concentration parameter $\kappa$.
+The distribution is radially symmetric around $\vc a$ with the density function for the angle $\theta$ between $\vc a$ and fracture normal:
+\[
+ f(\theta) = c_\theta \exp(\kappa \cos(\theta)) \sin(\theta),
+ \quad
+ c_\theta = \frac{\kappa}{2\pi(\exp(\kappa) - \exp(-\kappa)}.
+\] 
+The power law distribution is considered for the fracture size $r$.
+This kind of distribution is scale invariant, which is the key property of the natural fracture networks, see \cite{Bonnet2001} for detailed discussion.
+In practice the scale invariance holds only on some range of scales, leading to the truncated power law with the density:
+\begin{equation}
+    \label{eq:pow_law}
+    f(r) = c_r r^{-(\gamma+1)},\quad 
+    \frac{1}{c_r}=\int_{r_0}^{r_1} r^{-(\gamma+1)} dr = 
+    \frac{r_0^{-\gamma} - r_1^{-\gamma}}{\gamma},
+\end{equation}
+for $r \in [r_0, r_1]$. 
+For the given scale range, the number of fractures in unit volume is described by the Poison process with the intensity $\lambda=P_{30}[r_0, r_2]$ describing the mean number of fractures in the unit volume.
+The measurable parameter is the related mean total area per unit volume
+\[
+P_{32} =
+   \int_{r_0}^{r_1} r^2 f(r) dr =
+    P_{30} \frac{\gamma}{2-\gamma}    \frac{r_0^{2-\gamma} - r_1^{2-\gamma}}
+    {r_0^{-\gamma} - r_1^{-\gamma}}.
+\]
+In order to obtain fracture sets of the size tractable by our simulation tools, we represent explicitly only the fractures in the scale range $[r_2, r_3]$ while the scales $[r_0, r_1]$ are treated as equivalent porous media.
+The upper bound $r_3$ is set to the diameter of the simulation domain and upper scales are omitted.
+The lower bound is determined by desired mean intensity $P_{30}[r_2, r_3]$ through the relation:
+\eq{
+    \label{eq:pl_rescale}
+    \frac{P_{30}[r_2, r_3]}{P_{30}[r_0, r_1]} 
+    =
+    \frac{r_2^{-\gamma} - r_3^{-\gamma}}
+    {r_0^{-\gamma} - r_1^{-\gamma}}.
+}
+Positions of the fracture centers are sampled uniformly over the smaller box around the wells with dimensions $(300, 150, 150)$.
 
-The mechanical properties of granitic rocks have been investigated e.g. in \cite{Ljunggren1985}, where the measured Young modulus $E$ varies around 50 GPa and Poisson's ratio $\nu$ is around 0.25.
-The reported values of the drained compressibility $\beta$ of several types of granite and orthogneiss are approx.
-$20\times 10^{-13}$ cm${}^2$.dyne${}^{-1}$ = $2\times10^{-11}$ Pa${}^{-1}$ \cite{Zisman1933}.
-As far as porosity $\vartheta$ is concerned, \cite{intera} report values between 0.001 and 0.01 for metamorphic crystalline rocks, we shall consider the value 0.005.
-For water we use the density $\varrho_l=10^3$ kg.m${}^{-3}$, compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$.
-Here and in what follows, subscript ``l'' indicates quantities associated to the liquid.
-Similarly, subscript ``f'' will denote quantities for the fractures.
-Storativity $S$ is calculated from compressibility and porosity as follows (\cite{Brace1968}, see also \cite{intera}):
-\[ S = \varrho_l g(\beta + \vartheta\beta_l). \]
-Here $g=9.81$ m.s${}^{-2}$ is the standard gravity.
-In our case we obtain $S\approx 2\times 10^{-7} \mbox{ m}^{-1}$.
+Since no fracture statistics data for the target locality are available, we use data for the Forsmark site in Sweden
+(\cite{Follin}) with a crystalline bedrock. Parameters of the distributions for five families of the fractures are summarized in Table.
+\todo{JB: copy table from other word report}.
 
-The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
-The hydraulic conductivity of fractures dominates than the one of intact rock by several orders of magnitude, we consider $k_f=10^{-3}$ m.s${}^{-1}$
-On the other hand, the elastic modulus of fracture is assumed much smaller, we take $E_f=50$ Pa.
-The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$.
-We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
+caption: Used stochastic parameters of the five fracture families. Parameters of the distributions of the fracture orientation, size and intensity.
 
-All values of physical parameters used in our model are listed in Table \ref{tab:params}.
+% \subsection{Effective properties of the fractured porous media}
+% Assuming the cubic law for the permeability of fracture and fracture independency (over simplification) the effective permeability tensor of the grid of parallel infinite fractures of the aperture $b$ and distance $D$
+% is (\cite{Snow1969}):
+% \[
+%     K(b) = \frac{1}{12} \frac{b^3}{D} (I - \vc{n} \otimes \vc{n}) 
+% \]
+% Assuming superposition of fracture grids of different orientations, we can integrate over the $\vc n$ obtaining the factor $2/3$.
+Further, assuming simplest linear relationship between the aperture and fracture size (\cite{Klimczak2010}) with a factor $\alpha$, we obtain:
+% \begin{equation}
+%     \label{eq:iso_perm}
+%     k(b) = \frac{2}{3}\frac{1}{12} \frac{\alpha^3r^3}{D}
+% \end{equation}
+% for the isotropic permeability $k$. According to the power law \eqref{eq:power_law} distribution of the fracture sizes, there are
+% \begin{equation}
+%     \label{eq:n_frac}
+%     N \sim f(r)dr
+% \end{equation}
+% fractures in the unit volume, with $r\in [r, r+dr]$.
+We can imagine to arrange these randomly spread fractures into a grid of unit squares with distance $D$, there are $1/Dr^2$ these tiles per unit volume.
+Comparing to \eqref{eq:n_frac}, we obtain
+% \[
+%     \frac{1}{D} \sim c_r r^{(1-\gamma)}
+% \]
+% substituting into \eqref{eq:iso_perm} and integrating over $[r_0, r_2]$,
+% we obtain
+% \[
+%     k[r_0, r_2] \sim \frac{c_r}{18} \alpha^3 \frac{r_2^{(5-\gamma)} - r_0^{(5-\gamma)}}{5-\gamma}
+% \]
 
-\begin{table}
-\centering
-\begin{tabular}{|l|c|c|c|}
-\hline
- & rock & fractures & liquid\\\hline
-Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
-Hydraulic conductivity $k$ [m.s${}^{-1}$] & $10^{-9}$ \cite{Sperl2008} & $10^{-3}$ & ---\\
-Porosity $\vartheta$ [--] & $5\times10^{-3}$ & $5\times10^{-2}$ & --- \\
-Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ & $2\times10^{-11}$ & $5\times10^{-10}$ \\
-Storativity $S$ [m${}^{-1}$] & $2\times10^{-7}$ & $2\times10^{-7}$ & --- \\
-Cross-section $\delta$ [m] & --- & $10^{-3}$ & --- \\
-\hline
-Young modulus $E$ [Pa] & $50\times10^9$ & 50 & --- \\
-Poisson's ratio $\nu$ [--] & 0.25 & 0.25 & --- \\
-Biot's coefficient $\alpha$ [--] & 1 & 1 & --- \\
-\hline
-Heat capacity $c$ [J.kg${}^{-1}$.K${}^{-1}$] & 790 & 790 & 4000 \\
-Thermal conductivity $\lambda$ [W.m${}^{-1}$.K${}^{-1}$] & 2.5 & 2.5 & 0.5 \\
-\hline
-\end{tabular}
-\caption{Physical parameters of rock, fractures and water, used in the model.}
-\label{tab:params}
-\end{table}
+\section{Numerical results}
+The model described above was used to sample the production temperature and the heat power for the years 1 to 30. Single execution of the forward model consists of:
+\begin{enumerate}
+    \item drawing a fracture set sample according to Section 5.2
+    \item create the geometry with fracture intersections and generate the mesh using own code and \cite{GMSH} meshing tool
+    \item perform custom mesh healing algorithm to improve mesh quality
+    \item execute Flow123d for the the HM model
+    \item compute the modified aperture and conductivity for the fractures
+    \item execute Flow123d for the TH model with opened fractures
+    \item execute Flow123d for the TH model with unmodified fractures
+\end{enumerate}
+An example of a single realization of the fracture network is in Figure ...
+confinement of the fracture centers to a box around the wells is clearly visible, we obtain pretty good intersection with the well cylinders but the fracture network typically does not create a connect path between the wells.
+Due to technical reasons, we used the expected number of fractures 100 and set $r_2$ accordingly using \eqref{eq:pl_rescale}, the domain diameter was used for the upper fracture size limit $r_3=600$.
+We do not expect significantly better connectivity of the network for larger number of fractures because due to the nature of the power law this only adds more smaller fractures.
+Temperature distribution on the fractures and on the vertical cut of the domain for the year 5 can be seen in Figures X and Y.
+The influence of the fractures on the heat transfer in the bulk domain is clearly visible.
 
-Papers about effective properties of fractured porous media: \cite{Snow1969}, \cite{Bonnet2001}, \cite{Olson2003}
+\includegraphics[width=\textwidth]{sample_6_bulk_t5_fr_final.pdf}
+caption:
+A fracture set sample with distribution of the temperature in the year 5 of the operation with the stimulated fractures.
+\includegraphics[width=\textwidth]{sample_6_bulk_t5_final.pdf}
+caption:
+Distribution of the temperature on the bulk domain in the year 5 for a single sample.
 
-\subsection{Parameters of EGS site and model geometry}
+Total number of 1000 executions of the complete forward model was performed using a parallel cluster.
+These models have used a precomputed set of good meshes that were selected from about 2000 independent fracture sets.
+Further about 100 samples were rejected because of the extremely bad temperatures caused by the instability of the used numerical method for some extremely heterogeneous cases of the modified conductivity of the fractures.
+Remaining samples were used to estimate the mean and the standard deviance for the output temperature and the output power. Evolution of the mean and standard deviance is plotted in Figure ...
+The histograms for the selected times 1, 15 and 30 is in the Figure ...
+Absolute values especially for the power may not be realistic, but the important observation is the relation between the standard deviation and the gap between the stimulated and the reference case.
+As can bee seen from the histograms the distribution  for the power is close normal just slightly skewed for the initial times.
+Thus the standard variance corresponds to about 70\% of samples, say otherwise doing the stimulation
+there is a 15\% chance you get the power below the plotted std band, which is not very small probability and the value is already pretty close to the reference non-stimulated case.
 
-Number of wells: 2 \\
-Depth: 5 km \\
-Distance: 200 m \\
-Well radius: $\rho = 15$ cm \\
-Length of open part of wells: $L=60$ m \\
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{mlmc_random_frac/output_scenarios/ power_comparison.pdf}
+\includegraphics[width=0.5\textwidth]{temp_comparison.pdf}
+\caption{Evolution of the estimated mean output temperature and the output power for the stimulated and unmodified fracture network.
+The band width is the error of estimation the wide transparent band is the estimate of the standard variance due to the fracture dispersion again with outer band marking the error of the estimation.}
+\end{figure}
 
-% Pressure difference / water flux: 60 l/s
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{power_comparison.pdf}
+\includegraphics[width=0.5\textwidth]{temp_comparison.pdf}
+\caption{Evolution of the estimated mean output temperature and the output power for the stimulated and unmodified fracture network.
+The band width is the error of estimation the wide transparent band is the estimate of the standard variance due to the fracture dispersion again with outer band marking the error of the estimation.}
+\end{figure}
 
-In our model, we consider pair of vertical wells with the radius $\rho=0.15$ m in the distance $D=200$ m.
-The open part of the length $L=60$ m is in the depth $5000$ m.
-The cube $600 \times 600 \times 600$ m around the open parts is considered as the computational domain.
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{power_histogram.pdf}
+\includegraphics[width=0.5\textwidth]{temp_histogram.pdf}
+\caption{Histogram of the output temperature and power in 1,15, and 30 years of operation with the stimulated fractures.}
+\end{figure}
 
-{\bf The test geometry:} The wells are connected by three fractures, two of them intersecting the cylinders and one intersecting the first two fractures.
+\section{Conclusions}
+We have developed the HM model for the hydraulic stimulation and the TH model for the heat transfer using consistently the continuum-fracture approach.
+Although the model is conceptually simple neglecting important thermal effect and contact conditions during the stimulation yet it is capable to reproduce significant effect of the fracture stimulation on the output power and temperature compared to the reference case without stimulation.
 
-\section{Fractured media}
-\subsection{Stochastic description of the fractures}
-\subsection{Effective properties of the fractured porous media}
+We have shown, that the uncertainty of the power due to fracture dispersion is relatively high in the comparison to the power increase caused by the stimulation.
+It may indicate relatively high probability
+that the stimulation may not result in sufficiently high power for the long term production phase. Further research may investigate if the multiple stimulations can deal with such cases.
 

### 2019-08-09 mlmc_random_frac/process.py
First version of very slow algo for finding ngh of fractures in several levels.

@@ -1,5 +1,6 @@
 import sys
 import os
+import itertools
 script_dir = os.path.dirname(os.path.realpath(__file__))
 sys.path.append(os.path.join(script_dir, '../../MLMC/src'))
 sys.path.append(os.path.join(script_dir, '../../dfn/src'))
@@ -341,6 +342,83 @@ def call_flow(config_dict, param_key, result_files):
     return status # and conv_check
 
 
+def find_fracture_neigh(mesh, n_levels=3):
+    """
+    Find neighboring elements in the bulk rock in the vicinity of the fractures.
+    Creates several levels of neighbors.
+    :param mesh: GmshIO mesh object
+    :return: 
+    """
+
+    # {(int)level: [list of elements]}
+    fracture_neighbors = [[]]
+
+    # make node -> element map
+    node_els = collections.defaultdict(set)
+    for eid, e in mesh.elements.items():
+        type, tags, node_ids = e
+        for n in node_ids:
+            node_els[n].add(eid)
+
+    # for n in node_els:
+    #     if len(node_els[n]) > 1:
+    #         print(node_els[n])
+
+    # ele type: 1 - line, 2-triangle, 4-tetrahedron, 15-node
+    for eid, e in mesh.elements.items():
+        type, tags, node_ids = e
+        if type == 2: # fracture elements
+            ngh_elements = common_elements(node_ids, mesh, node_els)
+            # for ngh_eid, ngh_ele in [mesh.elements[i] for i in ngh_elements]:
+            for ngh_eid in ngh_elements:
+                ngh_ele = mesh.elements[ngh_eid]
+                ngh_type, ngh_tags, ngh_node_ids = ngh_ele
+                if ngh_type == 4:                                   # if they are bulk elements and not already added
+                    if ngh_eid not in fracture_neighbors[0]:
+                        fracture_neighbors[0].append(ngh_eid)     # add them to the first level
+
+    for lev in range(1, n_levels):                                # for all levels
+        fracture_neighbors.append([])
+        print("Finding fracture neighbors - level: {}".format(lev))
+        i = 0
+        for bulk_eid in fracture_neighbors[lev-1]:                  # go through already found fracture neighbors
+            i += 1
+            if i % 500 == 0:
+                n = len(fracture_neighbors[lev])
+                print("Bulk elements inspected: {0} Found neighbors: {1}".format(i, n))
[…]

### 2021-01-05 paper.tex
Improve paper.

@@
 
+\newcommand{\degC}{$^\circ\mathrm{C}$}
+
 \title{Stochastic modeling of EGS using continuum-fracture approach.}
@@
 \author{
-  Jan Březina \\
+  Jan B{\v r}ezina \\
   \And
@@
   \And
-  Martin Špetlík \\
+  Martin {\v S}petl{\' i}k \\
   %% \AND
@@
 \begin{abstract}
-Stochastic modeling of the hydraulic stimulation and long term operation of an enhanced geothermal system (EGS) is performed.
-The continuum-fracture approach is applied both for the stimulation model and for the operation model.
-Stochasticaly generated fractures and properties of the continuum are used in a simple hydro-mechanical model of the stimulation process.
-The resulting system of fractures with significant conductivity is used in the heat transfer model of the operation of the ground exchanger.
-The model is used to compute a single realization of key EGS properties: the output power and temperature.
-This complex sampling model is used within the Monte Carlo method predict probabilistic distribution for the EGS properties.
-The conceptual model and input data are loosely based on the geologic properties at the Litoměřice site in the Czech Republic.
+Stochastic modeling of the hydraulic stimulation and long-term operation of an enhanced geothermal system (EGS) is performed.
+Both the stimulation model and the operation model consistently use the continuum-fracture approach. Stochastically generated
+fractures are used in a simple hydro-mechanical model of the stimulation process. The resulting system of fractures with significantly
+higher conductivity is used in the heat transfer model of the operation regime of the underground exchanger. This model is used to
+compute a single realization of the key EGS properties: the output power and temperature. Monte Carlo method is then applied to
+predict probabilistic distribution of the EGS properties. The conceptual model and input data are loosely based on the geologic
+properties at the Litoměřice site, possible location of the first EGS in the Czech Republic.
 \end{abstract}
@@
 % keywords can be removed
-\keywords{First keyword \and Second keyword \and More}
+\keywords{EGS \and DFN \and stochastic modelling \and hydraulic stimulation}
 
 \section{Introduction}
-As the natural hydro thermal resources are sparse in the Czech Republic, the EGS (enhanced geothermal system) concept is subject of recent research focused on usage of the geothermal energy.
-According to the available geological data (\cite{Capova2013}), the Litoměřice site is the most attractive due to hot crystalline rock in relatively shallow depth (about 120 \textdegree C in 5km depth).
-This motivates development of the poroelasticity model within the Flow123d (\cite{flow123d}) simulator for the transport processes in the fractured porous media and its application in the study of uncertainty of the EGS properties caused by the fracture system.
+As the natural hydrothermal resources are sparse in the Czech Republic, the EGS (enhanced geothermal system) concept is subject
+of the recent research focused on the usage of the geothermal energy. According to the available geological data (\cite{Capova2013}), the
+Litoměřice site is the most attractive due to presence of a hot crystalline rock in relatively shallow depth. This motivates the
+implementation of the poroelastic model within the software Flow123d (\cite{flow123d}), the simulator
+of the transport processes in fractured porous media. The study exploits this development to investigate uncertainty in the operational
+properties of EGS caused by the stochastic nature of the fracture system.
 
-The EGS approach depends on opening and reconnecting of the preexisting fracture network in the low permeable rock by the hydraulic stimulation and the temperature changes.
-Since the fracture network is a-priori unknown as well as spatial distribution of the key parameters, we should consider the operational parameters of the EGS as random variables.
-In order to predict their distribution the Monte Carlo method can be applied provided the forward model and the distribution of the input parameters.
-Considering the forward model, there are three main approaches are used for simulation of the transport processes in the fractured porous media, namely single continuum, discrete fracture networks and continuum-fracture approach.
-The heterogeneous single continuum fully coupled THM model was used  in \cite{Watanabe2009} to study uncertainty in the output temperature.
-Usage of Monte Carlo method is natural in the DFN approach see e.g.
-\cite{Ezzedine2008}.
-We adopt the continuum-fracture approach which is much less common especially for significant technical difficulties.
-In \cite{Doonechaly2016} this approach was used for sensitivity analysis of the EGS operation, but only for the fixed fracture network.
+The EGS approach depends on opening and reconnecting of the preexisting fracture network in the low permeable rock induced by
+the hydraulic stimulation and the temperature changes. Since the fracture network is a-priori unknown as well as the spatial
+distribution of the essential parameters, we consider operational properties of the EGS as random variables. In order to predict their
+distribution, the Monte Carlo method can be applied, providing the forward model and the distribution of input parameters.
+Considering the forward model, three main approaches to the simulation of processes in a fractured porous media, namely single
+continuum, discrete fracture networks and continuum-fracture approach. The heterogeneous single continuum fully coupled THM
+model was used in \cite{Watanabe2009} to study uncertainty in the output temperature caused by spatially random input fields.
+The DFN approach essentially uses random fractures and thus is often combined with the Monte Carlo method, see e.g.
+\cite{ezzedine2008}.
+We adopt the continuum-fracture approach that is much less common especially for significantly more demanding
+implementation. In \cite{Doonechaly2016} this approach was used for sensitivity analysis of the EGS operation, but
+only within a fixed fracture network.
 
-We consider simple two stage forward model.
-First, the hydraulic stimulation is described by the hydro-mechanical poroelasticity model combining discrete fractures and continuum.
-Linear elasticity is used, contact conditions and thermal effects, however they may be significant, are not considered.
-Based on the stimulation model the changed aperture and conductivity fields on the fractures are determined.
-Then the evolution of the temperature and the raw power output is simulated using the combined fracture-continuum approach for the thermo-hydraulic model as well.
-A realistic stochastic model for the distribution of the fractures is considered, other parameters are fixed as we want to study particularly the influence of fracture variability.
+We consider a simple two stage forward model. At first the hydraulic stimulation is described by the hydro-mechanical (HM)
+poroelastic model combining the discrete fractures and the continuum. Linear elasticity is used, contact conditions and thermal effects
+are not considered, although these may have significant impact. Based on the stimulation model, the changed aperture and
+conductivity fields on the fractures are determined. Then the evolution of the temperature and the raw power output is simulated
+using the combined fracture-continuum approach for the thermo-hydraulic (TH) model as well. A realistic stochastic model for the
+distribution of the fractures is considered, while the other parameters are fixed as we want to study the influence of variability in the
+fracture network in particular.
 
@@
 below 3 km.
-Two geological scenarios predict temperatures 140 C and 146 C respectively in the depth 5 km.
+Two geological scenarios predict temperatures 140 \degC{}  and 146 \degC{} respectively in the depth 5 km.
 We consider in our model two vertical wells in distance 200 m with opened part of the length 100 m located in the target depth of 5 km.
@@
 cut off cylinders of the radius 10 m at the wells positions and apply boundary conditions on their walls while avoiding very fine meshing at the wells scale.
-This approximation is discussed in detail in Section \ref{}.
+This approximation is discussed in detail in Section \todo{ref}. %\ref{}.
 This simplified geometry is consistent with the simplification in the mathematical model as the precision of the model prediction is not of our interest in this work.
 We refer to the contribution paper \cite{Ralek} for the more EGS study for the Litoměřice locality.
+
+According to available geological data for the Litom{\v e}{\v r}ice site, see \cite{Capova2013}, the crystalline bedrock possibly consisting of granite, gneiss or mica schist is located in the depth below 3 km.
+Two geological scenarios lead to predicted temperatures in the depth of 5 km: 140 and 146 \degC{} respectively.
+In this depth, we consider two vertical wells in distance 200 m with opened part of the length 100 m.
+The simulation domain is a box with dimensions   m covering only the close neighborhood of the fractured volume with the two wells placed in its center.
+In order to reduce the complexity of the computational mesh, we cut off cylinders of the radius 10 m at the wells positions and apply boundary conditions on their walls while avoiding very fine meshing at the wells scale.
+This approximation is discussed in detail in Section 6.
+This simplified geometry is consistent with the simplification in the mathematical model as the precision of the model prediction is not of our interest in this work.
+We refer to the contribution paper \cite{Ralek} for the more thorough EGS study of the Litoměřice site.
 
@@
 During the hydraulic stimulation the fluid is injected into both wells in order to open the preexisting fractures.
-This leads to increase in the aperture of fractures, which is then used in the production phase model described in the next section.
-As the HM model of the bulk rock, we use the Biot poroelasticity equations:
+This leads to increase of aperture of the fractures, which is then used in the production phase model described in the next section.
+As the HM model of 3d rock we use the Biot poroelasticity equations:
 \eq{\label{eq:biot} \partial_t(Sp + \alpha\div\uu) +\div\vc q = 0,\qquad -\div\bbsigma + \tilde\alpha\nabla h = 0. }
@@
 = (\lambda_M I\otimes I +2\mu_M I):\epsilon, \quad \epsilon = (\nabla \uu^T + \nabla \uu)/2}
-with the Lamm\' e parameters: $\mu_M=E/(2(1+\nu))$ and $\lambda_M=E\nu/((1+\nu)(1-2\nu))$ [Pa].
+where $\mu_M=E/(2(1+\nu))$ and $\lambda_M=E\nu/((1+\nu)(1-2\nu))$
+are the Lammé parameters [Pa].
 
@@
 \end{aligned}\right\} }
-Here $p_f$ and $\uu_f$ denote the average pressure head and the average displacement on the fractures, respectively.
+Here $p_f$ and $\uu_f$ is the average pressure head and the average displacement on the fractures, respectively.
 The subscript $\tau$ indicates tangential operators on the fracture plane, $\otimes$ is the outer product and $\tilde\alpha_f:=\alpha_f\varrho_l g$.
-The jumps $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are the water sources and the forces, respectively,  that arise from the approximation of derivatives in the normal direction:
+The jumps $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are the water sources and the forces, respectively that arise from the approximation of derivatives in the normal direction:
 \eq{ 
@@
 
-For the sake of simplicity, we stimulate both wells simultaneously which results in very similar change of in the fracture aperture as if the wells were stimulated one by one.
+For the sake of simplicity we stimulate both wells simultaneously which results in very similar change of the fracture aperture as if the wells were stimulated one by one.
 Zero piezometric head and zero displacement are used as the initial conditions.
-The liquid is injected for 1 day under the stimulation pressure of 50 MPa (piezometric head $h_s=5e3$ m).
+The liquid is injected for 1 day under the stimulation pressure of 50 MPa (piezometric head $h_s=5\times10^3$ m).
 This pressure is not applied directly on the surface of the well's cylinders, but through the Robin boundary condition with the transition parameter derived from the approximate analytical solution at the vicinity of the well.
@@
 }
-for the bulk domain and fractures, respectively. Where $\tilde \delta$ is the fracture aperture updated according to the HM model.
-The conductivity of the fractures is also adjusted according to the cubic law $\tilde k_f = k_f\tilde\delta_f^2 / \delta_f^2$
+for the bulk domain and fractures, respectively. The fracture aperture $\tilde \delta$ is updated according to the HM model.
+The conductivity of the fractures is also adjusted according to the cubic law $\tilde k_f = k_f\tilde\delta_f^2 / \delta_f^2$.
 The communication between the fractures and the bulk rock is defined by the conditions \eqref{eq:fr_coupling} and \eqref{eq:fr_coupling_bulk}.
 
-The heat transport model is sequentially coupled to the hydraulic model using computed Darcy flux $\vc q$ and $\vc q_f$.
+The heat transport model is sequentially coupled with the hydraulic model using the computed Darcy flux $\vc q$ and $\vc q_f$.
 It assumes thermodynamic equilibrium between the liquid and the rock, the advection-diffusion equation for the bulk and fracture domain reads:
@@
 }
-Here $T$ [$\circ$C] is the temperature, $\varrho_l$, $\varrho_s$ is the density, $c_l$, $c_s$ is the heat capacity.
+Here $T$ [\degC] is the temperature, $\varrho_l$, $\varrho_s$ is the density, $c_l$, $c_s$ is the heat capacity.
 \eq{
@@
 
-Similarly, to \eqref{eq:fr_coupling} we obtain the relation for the normal heat fluxes on the fracture faces:
+Similarly, to \eqref{eq:fr_coupling} we obtain the relationship for the normal heat fluxes on the fracture faces:
 \eq{ 
-G^\pm &:= \pm\frac2\delta \lambda_f(T^\pm-T_f) + 
+G^\pm := \pm\frac2\delta \lambda_f(T^\pm-T_f) + 
 \varrho_l c_l F^\pm T_q,
 }
-where $T_q$ is equal to the bulk of fracture temperature as the normal Darcy flux $F^\pm$ is positive or negative, respectively. The system is completed by the bulk boundary conditions:
+where $T_q$ is equal to the temperature in the bulk or fractures as the normal Darcy flux $F^\pm$ is positive or negative, respectively. The system is completed by the bulk boundary conditions:
 \eq{ 
@@
 
-The liquid is continuously injected into the left well under the injection pressure 0.5 MPa ($h_i = 50$ m) and pumped out to the level of the EGS 
-($h_o = -5000$ m). No flux condition is used on the outer boundary. 
-The injecting temperature is set to $15\ ^\circ$C, the natural convection 
-heat flux is prescribed on the production well. The initial temperature field is considered on the outer boundary, which is given by 
- $10\ ^\circ$C at the Earth surface and the geothermal gradient $30\ ^\circ$C.km${}^{-1}$.
+The liquid is continuously injected into the left well under the pressure 0.5 MPa ($h_i = 50$ m) and pumped out to the level of the EGS ($h_o = -5000$ m).
+No flux condition is used on the outer boundary.
+The injecting temperature is set to 15 \degC, the natural convection heat flux is prescribed on the production well.
+The initial temperature field is considered on the outer boundary, which is given by 10 \degC{} at the Earth surface and the geothermal gradient 30 \degC.km${}^{-1}$.
 
@@
 in \cite{Ljunggren1985}, where the measured Young modulus $E$ varies around 50 GPa and Poisson's ratio $\nu$ is around 0.25.
-The Biot coefficient for different rocks is hard to obtain, we use the value $\alpha=0.6$ inspired by \cite{Selvadurai2019}, where $\alpha$ for Grimsel granite varies between 0.46 to 0.72.
+The Biot's coefficient for different rocks is hard to obtain, we use the value $\alpha=0.6$ inspired by \cite{Selvadurai2019}, where $\alpha$ for Grimsel granite varies between 0.46 to 0.72.
 The reported values of the drained compressibility $\beta$ of several types of granite and orthogneiss are approx.
@@
 %Similarly, subscript ``f'' will denote quantities for the fractures.
-Storativity $S$ is calculated from water compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$ and porosity following (\cite{Brace1968}, \cite{intera}):
+The storativity $S$ is calculated from water compressibility $\beta_l=5\times 10^{-10}$ Pa${}^{-1}$ and porosity following (\cite{Brace1968}, \cite{intera}):
 \[ S = \varrho_l g(\beta + \vartheta\beta_l) \approx 2\times 10^{-7} \mbox{ m}^{-1}. \]
@@
 The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
-The aperture of the fractures is considered constant in the input of the HM model according to \cite{Honza-clanek o laborratornich mikropuklinach??}.
+The aperture of the fractures is considered constant in the input of the HM model.
+The elastic modulus of fracture is assumed much smaller than in the bulk rock, $E_f=50$ Pa.
+After the process of the hydraulic fractioning, the aperture and hydraulic conductivity are adjusted according to the displacement results, see TH model in \eqref{th_darcy}.
+The porosity is taken much higher than in the rock due to the increased $k_f$, i.e.
+we consider $\vartheta=0.8$.
+We assume the same compressibility $\beta_f=\beta$ , which gives approximately the same value of storativity $S_f\approx S$.
+
+The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
+The aperture of the fractures is considered constant in the input of the HM model according to \todo{Honza-clanek o laborratornich mikropuklinach??}. %\cite{XXX}
 After the process of the hydraulic fractioning, the aperture and hydraulic conductivity are adjusted according to the displacement results in TH model in \eqref{th_darcy}.
@@
 We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
+
+The hydraulic conductivity $k$ in the bulk rock is set similarly to \cite{Sperl2008} and \cite{Doonechaly2016}, where they also consider DFN and surrounding rock.
+Additionally we assume that the hydraulic fractioning impacts also the smaller fractures not included in our DFN, so we increase $\tilde k=10^{-8}$ m.s${}^{-1}$ of the bulk rock.
 
@@
 Density $\varrho$ [kg.m${}^{-3}$] & 2700 & 2700 & 1000 \\
-Porosity $\vartheta$ [--] & $5\times10^{-3}$ & $0.8$ & --- \\
+Porosity $\vartheta$ [--] & 0.005 & 0.8 & --- \\
 Compressibility $\beta$ [Pa${}^{-1}$] & $2\times10^{-11}$ & $2\times10^{-11}$ & $5\times10^{-10}$ \\
@@
 \end{tabular}
-\caption{Physical parameters of rock, fractures and water, used in the model. \\ *) these values hold for HM model.}
+\caption{Physical parameters of rock, fractures and water, used in the model.}
 \label{tab:params}
@@
 \subsection{Analytical models in the vicinity of the well}
-We assume the well radii be $\rho = 15$ cm which is difficult to capture in the computational mesh.
 In order to keep the mesh size treatable, we cut out of the computational domain two cylinders around the open well parts with a larger radius $R=10$ m.
@@
 \subsection{Stochastic description of the fractures}
+\label{sec:stochastic_dfn}
 The fracture stochastic model consists of the Fisher's distribution for the fracture orientation, the power law for the fracture size, Poisson process for the number of fractures and uniform distribution for the fracture centers.
@@
 \section{Numerical results}
-The model described above was used to sample the production temperature and the heat power for the years 1 to 30. Single execution of the forward model consists of:
+The model described above is used to sample the production temperature and the heat power in the period of 30 years with 1 year time step.
+A single execution of the forward model consists of the following steps:
 \begin{enumerate}
-    \item drawing a fracture set sample according to Section 5.2
-    \item create the geometry with fracture intersections and generate the mesh using own code and \cite{GMSH} meshing tool
+    \item draw a fracture set sample according to Section \ref{sec:stochastic_dfn}
+    \item create the geometry with fracture intersections and generate a mesh using own Python code and GMSH meshing tool \cite{GMSH}
     \item perform custom mesh healing algorithm to improve mesh quality
-    \item execute Flow123d for the the HM model
-    \item compute the modified aperture and conductivity for the fractures
-    \item execute Flow123d for the TH model with opened fractures
-    \item execute Flow123d for the TH model with unmodified fractures
+    \item execute the HM model in Flow123d
+    \item compute the modified aperture and hydraulic conductivity in the fractures
+    \item execute the TH model in Flow123d with opened fractures
+    \item execute the TH model in Flow123d with unmodified fractures (reference sample)
 \end{enumerate}
-An example of a single realization of the fracture network is in Figure ...
-confinement of the fracture centers to a box around the wells is clearly visible, we obtain pretty good intersection with the well cylinders but the fracture network typically does not create a connect path between the wells.
-Due to technical reasons, we used the expected number of fractures 100 and set $r_2$ accordingly using \eqref{eq:pl_rescale}, the domain diameter was used for the upper fracture size limit $r_3=600$.
-We do not expect significantly better connectivity of the network for larger number of fractures because due to the nature of the power law this only adds more smaller fractures.
-Temperature distribution on the fractures and on the vertical cut of the domain for the year 5 can be seen in Figures X and Y.
+An example of a single realization of the fracture network can be seen in Figure \ref{fig:fracture_temp}.
+The confinement of the fracture centers to a box around the wells is clearly visible, we obtain pretty good intersection with the well cylinders but the fracture network typically does not create a connect path between the wells.
+Due to technical reasons, we use the expected number of fractures 100 and set $r_2$ accordingly using \eqref{eq:pl_rescale}, approximately 30 m.
+The domain diameter is used for the upper fracture size limit $r_3=600$ m.
+We do not expect significantly better connectivity of the fracture network for larger number of fractures due to the nature of the power law which permits adding only smaller fractures.
+The temperature distribution on the fractures and on the vertical cut of the domain after 5 years can be seen in Figure \ref{fig:fracture_temp} and Figure \ref{fig:bulk_temp}.
 The influence of the fractures on the heat transfer in the bulk domain is clearly visible.
 
+\begin{figure}
 \includegraphics[width=\textwidth]{sample_6_bulk_t5_fr_final.pdf}
-caption:
-A fracture set sample with distribution of the temperature in the year 5 of the operation with the stimulated fractures.
-\includegraphics[width=\textwidth]{sample_6_bulk_t5_final.pdf}
-caption:
-Distribution of the temperature on the bulk domain in the year 5 for a single sample.
-
-Total number of 1000 executions of the complete forward model was performed using a parallel cluster.
-These models have used a precomputed set of good meshes that were selected from about 2000 independent fracture sets.
-Further about 100 samples were rejected because of the extremely bad temperatures caused by the instability of the used numerical method for some extremely heterogeneous cases of the modified conductivity of the fractures.
-Remaining samples were used to estimate the mean and the standard deviance for the output temperature and the output power. Evolution of the mean and standard deviance is plotted in Figure ...
-The histograms for the selected times 1, 15 and 30 is in the Figure ...
-Absolute values especially for the power may not be realistic, but the important observation is the relation between the standard deviation and the gap between the stimulated and the reference case.
-As can bee seen from the histograms the distribution  for the power is close normal just slightly skewed for the initial times.
-Thus the standard variance corresponds to about 70\% of samples, say otherwise doing the stimulation
-there is a 15\% chance you get the power below the plotted std band, which is not very small probability and the value is already pretty close to the reference non-stimulated case.
-
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{mlmc_random_frac/output_scenarios/ power_comparison.pdf}
-\includegraphics[width=0.5\textwidth]{temp_comparison.pdf}
-\caption{Evolution of the estimated mean output temperature and the output power for the stimulated and unmodified fracture network.
-The band width is the error of estimation the wide transparent band is the estimate of the standard variance due to the fracture dispersion again with outer band marking the error of the estimation.}
+\label{fig:fracture_temp}
+\caption{A fracture set sample with distribution of the temperature in the year 5 of the operation with the stimulated fractures.}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{power_comparison.pdf}
-\includegraphics[width=0.5\textwidth]{temp_comparison.pdf}
+\includegraphics[width=\textwidth]{sample_6_bulk_t5_final.pdf}
+\label{fig:bulk_temp}
+\caption{
+Distribution of the temperature on the bulk domain in the year 5 for a single sample.}
+\end{figure}
+
+Total number of 1000 executions of the complete forward model was performed using a parallel cluster.
+These models used a precomputed set of good quality meshes that were selected from about 2000 independent fracture sets.
+Further, around 100 samples were rejected because of the unrealistic temperatures caused by the instability of the used numerical method for some extremely heterogeneous cases of the modified conductivity of the fractures.
+Remaining samples were used to estimate the mean and the standard deviance for the output temperature and the output power.
+Evolution of the mean and standard deviance is plotted in Figure \ref{fig:evolution_02}.
+The histograms for the selected times 1, 15 and 30 years are in Figure \ref{fig:histogrms}.
+Absolute values especially for the power may not be realistic, but the important observation is the relation between the standard deviation and the gap between the stimulated and the reference case.
+As can be seen from the histograms, the distribution for power is close to normal just slightly skewed for the initial times.
+Thus the standard variance corresponds to about 70 \% of samples, say otherwise: doing the stimulation there is a 15 \% chance you get the power below the plotted standard deviation band, which is not very small probability and the value is already pretty close to the reference non-stimulated case.
+
+%\begin{figure}[!htb]
+%\includegraphics[width=0.5\textwidth]{mlmc_random_frac/output_scenarios/output_por_disp_04/temp_comparison.pdf}
+%\includegraphics[width=0.5\textwidth]{mlmc_random_frac/output_scenarios/output_por_disp_04/power_comparison.pdf}
+%\caption{Evolution of the estimated mean output temperature and the output power for the stimulated and unmodified fracture network.
+The band width is the error of estimation, the wide transparent band is the estimate of the standard variance due to the fracture dispersion again with outer band marking the error of the estimation.}
+%\label{fig:evolution_01}
+%\end{figure}
+
+\begin{figure}[!htb]
+\includegraphics[width=0.49\textwidth]{mlmc_random_frac/output_scenarios/output_por_disp_02/temp_comparison.pdf}
+\includegraphics[width=0.49\textwidth]{mlmc_random_frac/output_scenarios/output_por_disp_02/power_comparison.pdf}
 \caption{Evolution of the estimated mean output temperature and the output power for the stimulated and unmodified fracture network.
 The band width is the error of estimation the wide transparent band is the estimate of the standard variance due to the fracture dispersion again with outer band marking the error of the estimation.}
+\label{fig:evolution_02}
 \end{figure}
 
+%We consider three scenarios:
+
+%A) no stimulation, no HM model, TH model of the EGS in its initial state
+
+%B) TH model of the EGS with increased conductivities of the bulk rock ($10^{-8}$ m.s$^{-1}$) and higher transfer coefficient of the Robin b.c. by 1 order
+
+%C) stimulation by HM model, TH model of the EGS with parameters dependent on the HM model
+
+%The first two scenarios are meant as a reference ones, so we can compare the influence of the DFN on the EGS properties.
+%In Figure \ref{fig:evolution_01} we see that the output temperature of the reference solution A (orange) is not decreasing at all, due to the fact that there is very small flux between the two wells.
+The heat power (green) is very low due to the small pumping flux.
+The fractured EGS (C) is doing much better as expected.
+The generated heat power is 10 times higher, however  it decreases during time following the temperature decrease below 100 \degC.
+
+%In Figure \ref{fig:evolution_02} we see that the output temperature of the reference solution B is now moderately decreasing, the rock cools down faster due the higher bulk rock conductivity which also causes higher flux exchange with the fractures.
+The heat power (green) is now larger since the flux increases with the higher conductivity as well as the pumping flux.
+Still the fractured EGS generates more power thanks to the increased flow by hydraulic fracturing.
+However we also see that there is significant variance due to the randomness of
+
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{power_histogram.pdf}
-\includegraphics[width=0.5\textwidth]{temp_histogram.pdf}
-\caption{Histogram of the output temperature and power in 1,15, and 30 years of operation with the stimulated fractures.}
+\includegraphics[width=0.48\textwidth]{mlmc_random_frac/output_scenarios/output_por_disp_02/temp_histogram.pdf}
+\includegraphics[width=0.48\textwidth]{mlmc_random_frac/output_scenarios/output_por_disp_02/power_histogram.pdf}
+\caption{Histogram of the output temperature and power in 1, 15 and 30 years of operation with the stimulated fractures.}
+\label{fig:histogrms}
 \end{figure}
@@
 We have developed the HM model for the hydraulic stimulation and the TH model for the heat transfer using consistently the continuum-fracture approach.
-Although the model is conceptually simple neglecting important thermal effect and contact conditions during the stimulation yet it is capable to reproduce significant effect of the fracture stimulation on the output power and temperature compared to the reference case without stimulation.
+Although the model is conceptually simple, it neglects important thermal effects and contact conditions during the stimulation, it is yet capable to reproduce a significant effect of the fracture stimulation on the output power and temperature compared to the reference case without stimulation.
 
 We have shown, that the uncertainty of the power due to fracture dispersion is relatively high in the comparison to the power increase caused by the stimulation.
-It may indicate relatively high probability
-that the stimulation may not result in sufficiently high power for the long term production phase. Further research may investigate if the multiple stimulations can deal with such cases.
+It may indicate relatively high probability that the stimulation may not result in sufficiently high power for the long term production phase.
+Further research may investigate if the multiple stimulation can deal with such cases.
 
@@
     \item[b)] RINGEN+:
-    The resrarch was supported by the project No. CZ.02.1.01/0.0/0.0/16\_013/0001792, co-funded by the EU Operational Programme ``Research, Development and Education''.
+    The research was supported by the project No. CZ.02.1.01/0.0/0.0/16\_013/0001792, co-funded by the EU Operational Programme ``Research, Development and Education''.
 \end{itemize}
@@
 
-% \begin{thebibliography}{9}
-
-% \bibitem{brace-et-al} W. F. Brace,  J. B. Walsh, W. T. Grango.  Permeability of Granite Under High Pressure, J. Geophys. Res. 73(6):2225--2236, 1968.
-
-% \bibitem{capova} L. Čápová. Specification of the geothermic model in the environs of several selected boreholes. Diploma thesis, Charles University in Prague, 2013.
-
-% \bibitem{intera} INTERA Environmental Consultants, Inc. Porosity, Permeability,  and Their  Relationship  in  
-% Granite,  Basalt, and Tuff. Accession  DE83-011519, NTIS, Springfield, Virginia, 1983.
-
-% \bibitem{martin-jaffre-roberts} V.
-Martin, J.
-Jaffré, J.
-E.
-Roberts.
-Modeling fractures and barriers as interfaces for flow in porous media.
-SIAM Journal on Scientific Computing 26(5):1667--1691, 2005.
-
-% \bibitem{sperl-trckova} J. Šperl, J. Trčková. Permeability and porosity of rocks and their relationship based on laboratory testing. Acta Geodyn. Geomater. 5(149):41--47, 2008.
-
-% \bibitem{ljunggren} C. Ljunggren, O. Stephansson, O. Alm, H. Hakami, U. Mattila. Mechanical properties of granitic rocks from Gide\aa, Sweden. Technical Report 85-06, SKB, 1985.
-
-% \bibitem{zisman} W.
-A.
-Zisman.
-Compressibility and anisotropy of rocks at and near the Earth's surface.
-Proceedings of the National Academy of Sciences of the United States of America, 19(7):666--679, 1933.
-
-% \end{thebibliography}
-
 \end{document}

### 2021-02-25 wgc2020_model/flow_mc_new.py
Add fracture level factor.

@@ -949,6 +949,7 @@ class Flow123d_WGC2020(Simulation):
 
         # find bulk elements neighboring to the fractures
         fr_n_levels = config_dict[base_variant]["increased_bulk_cond_levels"]
+        fr_lev_factor = config_dict[base_variant]["increase_level_factor"]
         fracture_neighbors = Flow123d_WGC2020.find_fracture_neigh(mesh, fr_regs, n_levels=fr_n_levels)
 
         # IMPORTANT - we suppose mesh nodes continuous, so we can find neighboring elements
@@ -966,7 +967,7 @@ class Flow123d_WGC2020(Simulation):
                                            n_contact_fr_elements)
 
         # increase the bulk conductivity in the vicinity of the fractures
-        level_factor = [10**(fr_n_levels - i) for i in range(fr_n_levels)]
+        level_factor = [fr_lev_factor**(fr_n_levels - i) for i in range(fr_n_levels)]
         for eid, lev in fracture_neighbors:
             assert lev < len(level_factor)
             # if eid >= len(ele_ids_map):

### 2021-02-25 wgc2020_model/flow_mc_new.py
Update config and time parameters.

@@ -243,7 +243,7 @@ class Flow123d_WGC2020(Simulation):
 
         # TODO: define times according to output times of Flow123d
         # TODO: how should be units defined (and other members)?
-        step = 10
+        step = 1
         end_time = 31
         times = list(range(0, end_time, step))
         spec = []

### 2021-02-25 wgc2020_model/run_process_pbs.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 #PBS -S /bin/bash
 #PBS -l select=1:ncpus=1:mem=2gb:cgroups=cpuacct
-#PBS -l walltime=24:00:00
+#PBS -l walltime=16:00:00
 #PBS -q charon
 #PBS -N WGC2020_main
 #PBS -j oe
@@ -9,4 +9,4 @@
 cd /auto/liberec3-tul/home/pavel_exner/WGC2020-THM-MC/wgc2020_model
 source load_modules.sh
 source env/bin/activate
-python process.py run /auto/liberec3-tul/home/pavel_exner/WGC2020-THM-MC/wgc2020_model/samples3
+python process.py run /auto/liberec3-tul/home/pavel_exner/WGC2020-THM-MC/wgc2020_model/samples

### 2021-02-25 paper.tex

@@
 \usepackage[square]{natbib}
+
+\usepackage{calc}
+\usepackage{makecell} % makecell for tables
+\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
+\newcolumntype{N}{@{}m{0pt}@{}}
 
@@
 predict probabilistic distribution of the EGS properties. The conceptual model and input data are loosely based on the geologic
-properties at the Litoměřice site, possible location of the first EGS in the Czech Republic.
+properties at the Litom{\v e}{\v r}ice site, possible location of the first EGS in the Czech Republic.
 \end{abstract}
@@
 of the recent research focused on the usage of the geothermal energy. According to the available geological data (\cite{Capova2013}), the
-Litoměřice site is the most attractive due to presence of a hot crystalline rock in relatively shallow depth. This motivates the
+Litom{\v e}{\v r}ice site is the most attractive due to presence of a hot crystalline rock in relatively shallow depth. This motivates the
 implementation of the poroelastic model within the software Flow123d (\cite{flow123d}), the simulator
@@
 
-The EGS approach depends on opening and reconnecting of the preexisting fracture network in the low permeable rock induced by
+The EGS approach depends on opening and reconnecting the preexisting fracture network in the low permeable rock induced by
 the hydraulic stimulation and the temperature changes. Since the fracture network is a-priori unknown as well as the spatial
@@
 Considering the forward model, three main approaches to the simulation of processes in a fractured porous media, namely single
-continuum, discrete fracture networks and continuum-fracture approach. The heterogeneous single continuum fully coupled THM
+continuum, discrete fracture networks and continuum-fracture approach. The heterogeneous single continuum fully coupled THM \todo{PE: abbreviation not defined}
 model was used in \cite{Watanabe2009} to study uncertainty in the output temperature caused by spatially random input fields.
@@
 \section{Conceptual model}
-According to available geological data for the Litoměřice locality (see \cite{Capova2013}), the crystalline rock possibly consisting of granite, gneiss or mica schist 
+According to available geological data for the Litom{\v e}{\v r}ice site, see \cite{Capova2013}, the crystalline bedrock possibly consisting of granite, gneiss or mica schist is located in the depth below 3 km.
 % zula, rula, svor
-is located in the depth
-below 3 km.
-Two geological scenarios predict temperatures 140 \degC{}  and 146 \degC{} respectively in the depth 5 km.
-We consider in our model two vertical wells in distance 200 m with opened part of the length 100 m located in the target depth of 5 km.
-The simulation domain is a box with dimensions $600\times600\times600$ m covering only the close neighbourhood of the stimulated domain.
-The wells are placed in the center.
-In order to reduce the complexity of the computational mesh, we
-cut off cylinders of the radius 10 m at the wells positions and apply boundary conditions on their walls while avoiding very fine meshing at the wells scale.
-This approximation is discussed in detail in Section \todo{ref}. %\ref{}.
+Two geological scenarios lead to predicted temperatures in the depth of 5 km: 140 \degC{} and 146 \degC{} respectively.
+In this depth, we consider two vertical wells in the distance 200 m with opened part of the length 100 m.
+The simulation domain is a box with dimensions $600\times600\times600$ m covering only the close neighborhood of the fractured volume with the two wells placed in its center.
+In order to reduce the complexity of the computational mesh, we cut off cylinders of the radius 10 m at the wells positions and apply boundary conditions on their walls while avoiding very fine meshing at the wells scale.
+This approximation is discussed in detail in Section \ref{sec:analytic_wells}.
 This simplified geometry is consistent with the simplification in the mathematical model as the precision of the model prediction is not of our interest in this work.
-We refer to the contribution paper \cite{Ralek} for the more EGS study for the Litoměřice locality.
-
-According to available geological data for the Litom{\v e}{\v r}ice site, see \cite{Capova2013}, the crystalline bedrock possibly consisting of granite, gneiss or mica schist is located in the depth below 3 km.
-Two geological scenarios lead to predicted temperatures in the depth of 5 km: 140 and 146 \degC{} respectively.
-In this depth, we consider two vertical wells in distance 200 m with opened part of the length 100 m.
-The simulation domain is a box with dimensions   m covering only the close neighborhood of the fractured volume with the two wells placed in its center.
-In order to reduce the complexity of the computational mesh, we cut off cylinders of the radius 10 m at the wells positions and apply boundary conditions on their walls while avoiding very fine meshing at the wells scale.
-This approximation is discussed in detail in Section 6.
-This simplified geometry is consistent with the simplification in the mathematical model as the precision of the model prediction is not of our interest in this work.
-We refer to the contribution paper \cite{Ralek} for the more thorough EGS study of the Litoměřice site.
+We refer to the contribution paper \cite{Ralek} for the more thorough EGS study of the Litom{\v e}{\v r}ice site.
 
@@
 are the Lammé parameters [Pa].
+\todo{define delta}
 
@@
 The subscript $\tau$ indicates tangential operators on the fracture plane, $\otimes$ is the outer product and $\tilde\alpha_f:=\alpha_f\varrho_l g$.
-The jumps $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are the water sources and the forces, respectively that arise from the approximation of derivatives in the normal direction:
+The jumps $\jmp{F}=F^+-F^-$ and $\jmp{\vc Q}=\vc Q^+-\vc Q^-$ are the water sources and the forces, respectively, that arise from the approximation of derivatives in the normal direction:
 \eq{ 
@@
 \section{Thermo-hydraulic model}
-The TH model describes the heat transport in the EGS through the operation phase within 30 years after the hydraulic stimulation performed according to Section \ref{sc:hm_model}.
+The TH model describes the heat transport in the EGS during the operation phase within 30 years after the hydraulic stimulation performed according to Section \ref{sc:hm_model}.
 In this scenario, we consider a steady Darcy flow:
@@
 for the bulk domain and fractures, respectively. The fracture aperture $\tilde \delta$ is updated according to the HM model.
-The conductivity of the fractures is also adjusted according to the cubic law $\tilde k_f = k_f\tilde\delta_f^2 / \delta_f^2$.
+The conductivity of the fractures is also adjusted according to the cubic law 
+\eq{ \label{eqn:cubic_law}
+\tilde k_f = k_f\tilde\delta_f^2 / \delta_f^2.
+}
 The communication between the fractures and the bulk rock is defined by the conditions \eqref{eq:fr_coupling} and \eqref{eq:fr_coupling_bulk}.
@@
 }
-where $\theta$ is the porosity and $\lambda_l,\lambda_s$ is the thermal conductivity.
+where $\vartheta$ is the porosity and $\lambda_l,\lambda_s$ is the thermal conductivity.
 
@@
 The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
-The aperture of the fractures is considered constant in the input of the HM model.
+The aperture of the fractures is considered constant in the input of the HM model according to \todo{JB-clanek o laborratornich mikropuklinach??}. %\cite{XXX}
 The elastic modulus of fracture is assumed much smaller than in the bulk rock, $E_f=50$ Pa.
-After the process of the hydraulic fractioning, the aperture and hydraulic conductivity are adjusted according to the displacement results, see TH model in \eqref{th_darcy}.
+After the process of the hydraulic fractioning, the aperture and hydraulic conductivity are adjusted according to the displacement results, see TH model in \eqref{eqn:cubic_law}.
 The porosity is taken much higher than in the rock due to the increased $k_f$, i.e.
 we consider $\vartheta=0.8$.
-We assume the same compressibility $\beta_f=\beta$ , which gives approximately the same value of storativity $S_f\approx S$.
+We assume the same compressibility $\beta_f=\beta$ , which gives approximately the same value of storativity $S_f\approx S$ \todo{check whether it is true - different porosity in fractures}.
 
-The values of physical parameters in the fractured zone are mostly unknown, hence the following values are only approximate.
-The aperture of the fractures is considered constant in the input of the HM model according to \todo{Honza-clanek o laborratornich mikropuklinach??}. %\cite{XXX}
-After the process of the hydraulic fractioning, the aperture and hydraulic conductivity are adjusted according to the displacement results in TH model in \eqref{th_darcy}.
 The hydraulic conductivity $k$ in the bulk rock is set similarly to \cite{Sperl2008} and \cite{Doonechaly2016}, where they also consider DFN and surrounding rock.
-Additionally we assume that the hydraulic fractioning impacts also the smaller fractures not included in our DFN,
-so we increase $k=10^{-8}$ of the bulk rock.
+Additionally we assume that the hydraulic fractioning impacts also the smaller fractures not included in our DFN, so we increase $k=10^{-8}$ of the bulk rock.
+\todo{PE: describe increased $k$ in the vicinity of fracture network.}
 
@@
 On the other hand, the elastic modulus of fracture is assumed much smaller, we take $E_f=50$ Pa.
-The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$.
+The porosity is taken one order higher that in the rock, i.e. we consider $\vartheta_f=0.05$ \todo{PE:nonsense-0.05 is one order lower,
+we have actually $\vartheta=0.005$ and $\vartheta_f=0.8$ in the model}.
 We assume the same compressibility $\beta_f=\beta$, which gives approximately the same value of storativity $S_f\approx S$.
-
-The hydraulic conductivity $k$ in the bulk rock is set similarly to \cite{Sperl2008} and \cite{Doonechaly2016}, where they also consider DFN and surrounding rock.
-Additionally we assume that the hydraulic fractioning impacts also the smaller fractures not included in our DFN, so we increase $\tilde k=10^{-8}$ m.s${}^{-1}$ of the bulk rock.
 
@@
 \subsection{Analytical models in the vicinity of the well}
+\label{sec:analytic_wells}
 In order to keep the mesh size treatable, we cut out of the computational domain two cylinders around the open well parts with a larger radius $R=10$ m.
@@
 Since no fracture statistics data for the target locality are available, we use data for the Forsmark site in Sweden
-(\cite{Follin}) with a crystalline bedrock. Parameters of the distributions for five families of the fractures are summarized in Table.
-\todo{JB: copy table from other word report}.
+(\cite{Follin}) with a crystalline bedrock. Parameters of the distributions for five families of the fractures are summarized in Table \ref{tab:stochastic_params}.
 
-caption: Used stochastic parameters of the five fracture families. Parameters of the distributions of the fracture orientation, size and intensity.
+\begin{table}[!htb]
+\centering
+\begin{tabular}{|M{15pt}|M{20pt}|M{25pt}|M{30pt}|M{22pt}|M{22pt}|M{22pt}|M{20pt}|M{54pt}|N}
+\hline
+\multicolumn{9}{|M{275pt+10\tabcolsep}|}{SKB data - FFM01 and FFM06 (depth 0 to -200 m)} &\\[10pt]
+\hline
+\multicolumn{2}{|M{35pt+3\tabcolsep}|}{FFM01a} &
+\multicolumn{3}{M{77pt+4\tabcolsep}|}{Fisher orientation} &
+\multicolumn{4}{M{118pt+5\tabcolsep}|}{Power law}&\\[10pt]
+\hline
+Set & Name &
+\makecell{Trend \\ $[^\circ]$} &
+\makecell{Plunge \\ $[^\circ]$} &
+\makecell{Conc. \\ $[-]$} &
+\makecell{$r_0$ \\ $[\mathrm{m}]$} &
+\makecell{$k_r$ \\ $[-]$ } &
+\makecell{$r_{\mathrm{max}}$ \\ $[\mathrm{m}]$} &
+\makecell{$P_{32}[r_0,r_{\mathrm{max}}]$ \\ $[\mathrm{m}^2\cdot\mathrm{m}^{-3}]$} & \\[10pt]\hline
+1 & NS & 292 & 1  & 17.8 & 0.038 & 2.50 & 564 & 0.073 &\\[8pt]\hline
+2 & NE & 326 & 2  & 14.3 & 0.038 & 2.70 & 564 & 0.319 &\\[8pt]\hline
+3 & NW & 60  & 6  & 12.9 & 0.038 & 3.10 & 564 & 0.107 &\\[8pt]\hline
+4 & EW & 15  & 2  & 14.0 & 0.038 & 3.10 & 564 & 0.088 &\\[8pt]\hline
+5 & HZ & 5   & 86 & 15.2 & 0.038 & 2.38 & 564 & 0.543 &\\[8pt]
+\hline
+\end{tabular}
+\caption{Used stochastic parameters of the five fracture families. Parameters of the distributions of the fracture orientation, size and intensity.}
+\label{tab:stochastic_params}
+\end{table}
 
@@
 An example of a single realization of the fracture network can be seen in Figure \ref{fig:fracture_temp}.
-The confinement of the fracture centers to a box around the wells is clearly visible, we obtain pretty good intersection with the well cylinders but the fracture network typically does not create a connect path between the wells.
-Due to technical reasons, we use the expected number of fractures 100 and set $r_2$ accordingly using \eqref{eq:pl_rescale}, approximately 30 m.
+The confinement of the fracture centers to a box around the wells is clearly visible, we obtain pretty good intersection with the well cylinders but the fracture network typically does not create a connected path between the wells.
+Due to technical reasons, we use the expected number of fractures 300 and set $r_2$ accordingly using \eqref{eq:pl_rescale}, approximately 30 m.
 The domain diameter is used for the upper fracture size limit $r_3=600$ m.
@@
 \begin{itemize}
-    \item[a)] RINGEN:
-    The research was supported by the Czech Ministry of Education, Youth and Sports under the project No. LM2015084.
+    \item[a)] RINGEN+:
+    The research was supported by the project No. CZ.02.1.01/0.0/0.0/16\_013/0001792, co-funded by the EU Operational Program ``Research, Development and Education''.
 
-    \item[b)] RINGEN+:
-    The research was supported by the project No. CZ.02.1.01/0.0/0.0/16\_013/0001792, co-funded by the EU Operational Programme ``Research, Development and Education''.
+    \item[b)] Metacentrum:
+    Computational resources were supplied by the project ``e-Infrastruktura CZ'' (e-INFRA LM2018140) provided within the program Projects of Large Research, Development and Innovations Infrastructures.
 \end{itemize}

### 2021-02-26 wgc2020_model/flow_mc_new.py
Saves some additional stats on fracture generation.

@@ -454,8 +454,11 @@ class Flow123d_WGC2020(Simulation):
         # possibly read fractures from file
         if os.path.isfile('fractures.yaml'):
             with open('fractures.yaml', 'r') as outfile:
-                deserialized_fracs = yaml.load(outfile, yaml.CSafeLoader)
-                fractures = [fracture.FractureShape.load(df) for df in deserialized_fracs]
+                loaded_dict = yaml.load(outfile, yaml.CSafeLoader)
+                fractures = [fracture.FractureShape.load(df) for df in loaded_dict["fractures"]]
+                print("final sample range: ", loaded_dict["sample_range"])
+                print("final fracture count: ", len(fractures))
+                print("total mean size: ", loaded_dict["mean_size"])
                 return fractures
 
         geom = config_dict["geometry"]
@@ -470,8 +473,9 @@ class Flow123d_WGC2020(Simulation):
         volume = np.product(fracture_box)
         pop = fracture.Population(volume)
         pop.initialize(geom["fracture_stats"])
-        pop.set_sample_range([1, well_dist], max_sample_size=geom["n_frac_limit"])
-        print("total mean size: ", pop.mean_size())
+        sample_range = pop.set_sample_range(sample_range=[geom["fracture_rmin"], geom["fracture_rmax"]],
+                                            max_sample_size=geom["n_frac_limit"])
+
         connected_position = geom.get('connected_position_distr', False)
         if connected_position:
             eps = well_r / 2
@@ -485,14 +489,21 @@ class Flow123d_WGC2020(Simulation):
         fractures = pop.sample(pos_distr=pos_gen, keep_nonempty=True)
         # fracture.fr_intersect(fractures)
 
+        print("final sample range: ", sample_range)
+        print("final fracture count: ", len(fractures))
+        print("total mean size: ", pop.mean_size())
+
         for i, fr in enumerate(fractures):
             fr_name = "fr_" + str(i)
             fr.region = fr_name
 
         # write fracture data into a file for possible future deserialization
-        fractures_list = [f.yaml() for f in fractures]
+        fracture_dict = {"sample_range": [float(sample_range[0]), float(sample_range[1])],
+                         "mean_size": float(pop.mean_size()),
+                         "box": fracture_box,
+                         "fractures": [f.yaml() for f in fractures]}
         with open('fractures.yaml', 'w') as outfile:
[…]

### 2021-02-26 wgc2020_model/fracture.py

@@ -645,6 +645,8 @@ class Population:
                     family_intensity = size / total_size * max_sample_size / self.volume
                     f.shape.set_range_by_intensity(family_intensity)
 
+        return f.shape.sample_range
+
     def sample(self, pos_distr = None, keep_nonempty=False):
         """
         Provide a single fracture set  sample from the population.

### 2021-02-26 wgc2020_model/flow_mc_new.py
Minor cleanup and fixes.

@@ -398,52 +398,52 @@ class Flow123d_WGC2020(Simulation):
         fractures = []
         f1 = fracture.FractureShape(
             r=140,
-            centre=np.array([-100, -5, 15]),
-            rotation_axis=np.array([-0.17, -0.98, 0]),
-            rotation_angle=0.3,
-            shape_angle=4.2,
-            aspect=1,
-            region='fr_1')
+            c=np.array([-100, -5, 15]),
+            rax=np.array([-0.17, -0.98, 0]),
+            rag=0.3,
+            sg=4.2,
+            asp=1,
+            reg='fr_1')
         fractures.append(f1)
 
         f2 = fracture.FractureShape(
             r=140,
-            centre=np.array([100, 5, -5]),
-            rotation_axis=np.array([0.07, -0.99, 0]),
-            rotation_angle=2.48,
-            shape_angle=0.6,
-            aspect=1,
-            region='fr_2')
+            c=np.array([100, 5, -5]),
+            rax=np.array([0.07, -0.99, 0]),
+            rag=2.48,
+            sg=0.6,
+            asp=1,
+            reg='fr_2')
         fractures.append(f2)
 
         f3 = fracture.FractureShape(
             r=65,
-            centre=np.array([0, 0, 120]),
-            rotation_axis=np.array([0.5, 0.5, 0]),
-            rotation_angle=1.31,
-            shape_angle=1.56,
-            aspect=1,
-            region='fr_3')
+            c=np.array([0, 0, 120]),
+            rax=np.array([0.5, 0.5, 0]),
+            rag=1.31,
+            sg=1.56,
+            asp=1,
+            reg='fr_3')
         fractures.append(f3)
 
         f4 = fracture.FractureShape(
             r=65,
-            centre=np.array([0, 0, -95]),
-            rotation_axis=np.array([0.5, -0.5, 0]),
-            rotation_angle=1.31,
-            shape_angle=1.56,
-            aspect=1,
-            region='fr_4')
+            c=np.array([0, 0, -95]),
+            rax=np.array([0.5, -0.5, 0]),
+            rag=1.31,
+            sg=1.56,
+            asp=1,
+            reg='fr_4')
         fractures.append(f4)
 
         f5 = fracture.FractureShape(
             r=40,
-            centre=np.array([0, 0, 10]),
-            rotation_axis=np.array([0.5, -0.5, 0]),
-            rotation_angle=0.1,
[…]
